### 2021-03-08 comment.tex
Initial Overleaf Import

\documentclass{article}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{amssymb}
\usepackage{subcaption}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{marginnote}
\usepackage[colorlinks=true,urlcolor=blue,citecolor=red,linkcolor=red,bookmarks=true]{hyperref}
%strikethrough
\usepackage{soul}
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usepackage{float}
\usepackage[margin=0.5in]{geometry}

\newcommand{\fp}{P_{\text{fp}}}
\newcommand{\fn}{P_{\text{fn}}}
\newcommand{\Se}{S_e}
\newcommand{\Sp}{S_p}
\newcommand{\mi}{P_{\text{sfn}}}
\newcommand{\rel}{E_{\text{rel}}}
\renewcommand\Affilfont{\itshape\small}
\renewcommand{\Pr}{\mathbb{P}}

\begin{document}

\title{Inflated false-negative rates in pooled RT-PCR tests of SARS-CoV-2}

\author[1,2]{Yair Daon, PhD}
\author[2,3]{Amit Huppert, PhD}
\author[1,2]{Uri Obolski, PhD}

\affil[1]{Porter School of the Environment and Earth Sciences, The Faculty of Exact Sciences, Tel Aviv
  University\\ Tel Aviv, Israel}
\affil[2]{School of Public Health, The Faculty of Medicine, Tel Aviv University\\ Tel Aviv, Israel}
\affil[3]{The Gertner Institute for Epidemiology and Health Policy Research, Tel Hashomer, Israel}
%\email{yair.daon@gmail.com}
\date{\today}

\maketitle
%%% MAKE TITLE PAGE %%%
% \bigskip
% \bigskip
% \bigskip
% \bigskip
% \begin{itemize}
%     \item Article type: Brief report.
%     \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
%     \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
%     \item Conflict of interests: None.
%     \item Financial support: Yair Daon was supported by a
%     post-doctoral fellowship from the Tel Aviv University Center for
%     Combating Pandemics and the Raymond and Beverly Sackler Dean's
%     Post-Doctoral fellowship.
%     \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
% \end{itemize}
% \end{document}

\begin{abstract}
\textbf{Background}: Pooling is a popular strategy for increasing SARS-CoV-2 testing throughput. One popular pooling scheme is Dorfman pooling: test $N$ individuals simultaneously. If the test is positive --- retest each individual separately. However, requiring more than one positive test may lead to increased false-negative rates.

\textbf{Methods}: We analyze the false-negative rate (i.e., the probability of a negative result for an infected individual) of Dorfman pooling via a new probabilistic model. We demonstrate that different, previously made probabilistic assumptions regarding pooling are unlikely in light of empiric data. Our model is conservative in that it ignores sample dilution effects, which can only worsen pooling performance.

\textbf{Results}: We show that one can expect a 60-80\% increase in false-negative rates under Dorfman pooling, for reasonable parameter values. Moreover, we show that the false-negative rates under Dorfman pooling increase when the prevalence of infection decreases.

\textbf{Discussion}: In most pooling schemes, identifying an infected individual requires positive results in multiple tests and hence substantially increases false-negative rates. Furthermore, this phenomenon is more pronounced when infection prevalence is low --- exactly when pooling is most efficient. Thus, pooling presents an inherent trade-off: it is most efficient when it is least accurate. The deterioration of false-negative rates and the aforementioned trade-off are inherent problems of pooling schemes and should be kept in mind by practitioners and policy makers.
\end{abstract}

\section{Introduction}

RT-PCR testing is a key component in breaking transmission chains and mitigating the COVID-19 pandemic. As such, the need for large-scale testing has resulted in the development of pooling schemes of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman, MatrixPooling, LionDorfman}. One such popular scheme is Dorfman pooling \cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$ individuals and perform a single RT-PCR test on their combined (``pooled'') samples. If the pooled test yields a positive result --- test each individual separately. The throughput efficiency of Dorfman pooling has been demonstrated empirically \cite{DorfmanYuvalDor}. However, when test error rates are taken into consideration, a sharp increase in false-negative rates can be expected.

It is important to distinguish three types of false-negative events when performing pooling. For convenience, we follow a single \emph{infected} individual, henceforth referred to as "Donald". A \emph{single test's} false-negative is the event of a negative result upon testing Donald separately, i.e., in an RT-PCR test without pooling. We denote the test sensitivity $\Se$, so the probability of a single test false-negative is $1-\Se$. A \emph{pooled} false-negative occurs when a pooled test containing Donald's sample (and other samples) yields a negative result, i.e., the pooling fails to detect at least one positive result. Lastly, a \emph{scheme} false-negative occurs when an entire pooling scheme fails to identify Donald as infected. Our goal is to calculate Dorfman's scheme false-negative rate. Rephrasing, we wish to answer the following question: what is the probability of not identifying Donald as infected under a Dorfman pooling scheme?

\section{Methods}
\subsection{Probabilistic Assumptions}\label{subsec:assumptions}
We assume two pathways for a positive pooled test result: Viral RNA from an infected individual is correctly detected; or, some erroneous detection occurs (e.g. contaminant viral RNA is introduced). We ignore cross-reactivity with other Coronaviruses, which is negligible \cite{KitComparison}. We assume a homogeneous and disconnected population (each individual is infected independently and with equal probability). For simplicity, we do not take into account sample dilution, since it can only further increase false-negative rates \cite{DorfmanYuvalDor}.

Our assumptions, although natural, stand in contrast to assumptions commonly made in the literature \cite{Kim, Simplistic1, Simplistic2, OptimalDorfmanPool}. It is commonly assumed that the probability of a positive pooled test is not increased by having more than one infected individual in the pool. We refute the common assumptions with experimental data summarized in table \ref{table}, collected from \cite{Salazar}. There, the authors investigate Dorfman pooling and, regardless of the pooled test result, follow up and test each pool member separately. Here, we focus on 128 pools for which at least one subsequent separate test was positive --- of which 29 pooled tests were negative and 99 positive. In the data cited in \cite{Salazar}, of the 29 negative pools, subsequent separate testing yielded a single positive result in 24. In contrast, of the 99 positive pools, 42 yielded a single positive test upon subsequent separate testing.

The data in table \ref{table} allows us to test the following null hypothesis $H_0$: The probability of a pooled false-negative is equal for pools with one subsequent positively tested member and pools with two or more such members. We apply Fisher's exact test for the presence of more than one positive individual in correctly identified pools. Fisher's test yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a p-value $\approx 10^{-4}$. Thus we reject $H_0$, refuting the independence of a pooled result from one or more individuals infected within the pool, as assumed in \cite{Simplistic1, Simplistic2, OptimalDorfmanPool, Kim}

\begin{table}[h]
\centering
\begin{tabular}{ c c c }
                                & Negative pool  & Positive pool \\%& Row total \\ 
\# subsequent positives $=1$    & 24             & 42            \\%& 66        \\  
\# subsequent positives $\geq1$ & 5              & 57            \\%& 62        \\
 Total                          & 29             & 99            \\%& 128 
\end{tabular}
\caption{Contingency table of data from \cite{Salazar}}\label{table}
\end{table}

\subsection{calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
Denote the prevalence of infection in the (tested) population $q$. As before, $\Se$ is the test's sensitivity, so $1-\Se$ is the single RT-PCR test's false-negative rate. We also denote $\Sp$ the test specificity --- the true-negative probability. This encompasses events of an erroneous RNA detection (via any possible pathway), which may cause a false-positive. By our assumptions, a pool containing Donald's sample and $N-1$ other samples will yield a \emph{negative} result if all of the following occur: 
\begin{itemize}
\item No erroneous detection occurred (i.e. no false-positive). This happens with probability $\Sp$.
\item The detection process fails for Donald's sample. A false-negative occurs for Donald, with probability $1-\Se$.
\item No detection for any of the other $N-1$ samples. For a single sample, the probability of being detected is the prevalence of SARS-CoV-2 in the tested population $q$, multiplied by the sensitivity $\Se$, hence the probability of detection is $q\Se$. For $N-1$ such samples, the probability of \emph{not} being identified is $(1-q\Se)^{N-1}$. 
\end{itemize}
The pooled false-negative probability for Donald is simply the product of the terms above. Hence:

\begin{align}
    \begin{split}
        \Pr(\text{pool is positive}) &= 1 - \Pr(\text{pool false-negative}) \\
        &= 1 - \Sp(1-\Se)(1-q\Se)^{N-1}.
    \end{split}
\end{align}

If the pooled test yields a positive result, Donald is tested separately. We assume such a simple procedure poses no risk of introducing contaminant RNA. Therefore, the separate test yields a positive result with probability $\Se$.

We calculate the probability that Donald is mistakenly identified as not infected --- the scheme's false-negative rate --- denoted $\mi$ below. To \emph{correctly} identify Donald as infected, both pooled and separate tests have to yield a positive result. Thus, the scheme's false-negative rate $\mi$ is the complement of the product of the two previous terms:
\begin{align}\label{eq:sfn}
    \begin{split}
        \mi :&= 1-\Pr(\text{correctly identify Donald as infected})\\
        %
        &= 1 - \Se\left [1 - \Sp(1-\Se)(1-q\Se))^{N-1}\right].
    \end{split}
\end{align}

\subsection{Comparison metric}
The single test false-negative rate $1-\Se$ and scheme false-negative rate $\mi$ are compared via:
\begin{equation}\label{eq:erel}
\rel := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
\end{equation}
$\rel$ is the percentage increase in the pooling scheme false-negative rate, relative to the single test false-negative rate.

For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (section \ref{section:ass} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}). A short calculation shows that this approximation implies the percentage increase in scheme false-negative rate is just $100\cdot \Se$.

\section{Results}\label{section:results}

We plot $\rel$ for varying prevalence $q$ and sensitivity $\Se$ values. As recommended by \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for different prevalence values. We observe that for a false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable sensitivity and prevalence values \cite{KitComparison, InterpretingCOVID19Test, EstimatingRatesLourenco, FalsePositiveEstimate}, an increase of at least $60\%$ in $\rel$ can be expected (Figure \ref{figy}). Interestingly, an increase in infection prevalence monotonically decreases the scheme false-negative rate, as can also be easily seen from equation \eqref{eq:sfn}. For the chosen parameter ranges, the increase in the single test false-negative rates increases the relative error $\rel$. These effects can be seen in Figure \ref{figy} (left panel), upon conditioning on pool size. Extending the range for $\Sp$ yields no qualitative differences. We further compare $\rel$ to the commonly used approximation, showing the discrepancy changes as a function of both prevalence and the single test sensitivity (Figure \ref{figy}, right panel)   
\begin{figure}[H]
    \centering
    \includegraphics[width=18cm]{heatmap.jpg}
    \caption{Relative increase in Dorfman pooling false-negative rates $\rel$. Left: Colors represent $\rel$, the relative percentage increase in the scheme false-negative rates relative to the single test false-negative rates (eq \eqref{eq:erel}). Right: colors represent the difference between the approximation assumed in other studies and $\rel$ as calculated by our model. The disease prevalence $q$, is varied on the x-axis, while the test sensitivity is varied on the y-axis. Pool size $N$, was chosen according to $q$ as in \cite{DorfmanYuvalDor}.}\label{figy}
 \end{figure}

\section{Discussion}
Although Dorfman pooling improves testing throughput, we have shown that it can increase $\rel$ --- the false-negative rates relative to individual testing. Furthermore, low values of infection prevalence, or low values of single test false-negative rates, increase $\rel$. These results remain qualitatively similar under varying parameter values, in the observed ranges \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure \ref{figy}).

Our results lead to a fact almost disregarded in \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient when prevalence is low, such circumstances are exactly those leading to a substantial increase in false-negative rates.

An approximation to our results has been previously considered: $\mi \approx 1-\Se^2 $\cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}. Such an approximation arises when one assumes that the probability of a false-negative is identical for the pooled and single tests. Although simplistic in nature, this approximation does capture the intuition behind our results: For Donald to be considered negative under Dorfman pooling, he has to test negative twice. In addition to this approximation only being an upper bound \cite{Simplistic2}, it does not account for the effect of infection prevalence on the false-negative rates under Dorfman pooling. Supporting the importance of this effect, such an association between prevalence and the scheme false-negative rate under Dorfman pooling has been empirically noted  \cite{DorfmanYuvalDor}.

Although we have shown the inherent risk of Dorfman pooling, this shortcoming applies to other pooling schemes. Pooling schemes (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of positive pooled results to correctly identify Donald as infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}: If the first pool yields a positive result, it is split in two. Then the splitting is repeated until resulting pools are negative or individuals are tested separately. With an initial pool size of 32, Donald will necessarily have to test positive in pools of size 32, 16, 8, 4 and 2, as well as in a single test, for the scheme to correctly identify him as infected. Compare this to the Dorfman scheme that requires a positive test in a pool of size $N=8$, and an additional single positive test to identify Donald as infected. The hierarchical pooling scheme of \cite{Lion, Kim} will necessarily yield  more false-negatives than Dorfman pooling --- there are additional places for it to fail.

As mentioned in \cite{DorfmanYuvalDor}, introducing a positive dependence within a pool decreases the false-positive rate. In the extreme case, consider a fully connected pool, where one infection implies the entire pool is infected. In this case, a calculation analogous to the one conducted above recovers the initial false-negative rate $1-\Se$. Interestingly, pooling was also noted to have increased throughput when infection probabilities are dependent between the pooled individuals \cite{DorfmanYuvalDor}, providing  another advantage to sampling dependent individuals in pooling schemes.

To conclude, pooling is an important technique which can improve testing throughput in a cost-effective manner. Nevertheless, a substantial increase in pooling schemes' false-negative rates can be expected. Such a increase has crucial implications for controlling the spread of COVID-19.

\bibliographystyle{amsplain}
\bibliography{pooling, fpfn}

\end{document}

### 2021-03-09 comment.tex
ama bst
many more refs

@@
 \usepackage[margin=0.5in]{geometry}
+\hypersetup{linkcolor=black,citecolor=black,urlcolor=black}
 
@@
 %\email{yair.daon@gmail.com}
-\date{\today}
+\date{}
 
@@
 %%% MAKE TITLE PAGE %%%
-% \bigskip
-% \bigskip
-% \bigskip
-% \bigskip
-% \begin{itemize}
-%     \item Article type: Brief report.
-%     \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
-%     \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
-%     \item Conflict of interests: None.
-%     \item Financial support: Yair Daon was supported by a
-%     post-doctoral fellowship from the Tel Aviv University Center for
-%     Combating Pandemics and the Raymond and Beverly Sackler Dean's
-%     Post-Doctoral fellowship.
-%     \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
-% \end{itemize}
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\begin{itemize}
+    \item Article type: Brief report.
+    \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
+    \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
+    \item Conflict of interests: None.
+    \item Financial support: Yair Daon was supported by a
+    post-doctoral fellowship from the Tel Aviv University Center for
+    Combating Pandemics and the Raymond and Beverly Sackler Dean's
+    Post-Doctoral fellowship.
+    \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
+\end{itemize}
+\newpage
 % \end{document}
@@
 
-For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (section \ref{section:ass} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}).
+For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (Section \ref{section:assumptions} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}).
 A short calculation shows that this approximation implies the percentage increase in scheme false-negative rate is just $100\cdot \Se$.
@@
 
-An approximation to our results has been previously considered: $\mi \approx 1-\Se^2 $\cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}.
+An approximation to our results has been previously considered: $\mi \approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}.
 Such an approximation arises when one assumes that the probability of a false-negative is identical for the pooled and single tests.
@@
 
-\bibliographystyle{amsplain}
-\bibliography{pooling, fpfn}
+\bibliographystyle{ama}
+\bibliography{refs}
 

### 2021-03-09 comment.tex
final jama

@@
 %%% MAKE TITLE PAGE %%%
-\bigskip
-\bigskip
-\bigskip
-\bigskip
-\begin{itemize}
-    \item Article type: Brief report.
-    \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
-    \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
-    \item Conflict of interests: None.
-    \item Financial support: Yair Daon was supported by a
-    post-doctoral fellowship from the Tel Aviv University Center for
-    Combating Pandemics and the Raymond and Beverly Sackler Dean's
-    Post-Doctoral fellowship.
-    \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
-\end{itemize}
-\newpage
-% \end{document}
+%% \bigskip
+%% \bigskip
+%% \bigskip
+%% \bigskip
+%% \begin{itemize}
+%%     \item Article type: Brief report.
+%%     \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
+%%     \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
+%%     \item Conflict of interests: None.
+%%     \item Financial support: Yair Daon was supported by a
+%%     post-doctoral fellowship from the Tel Aviv University Center for
+%%     Combating Pandemics and the Raymond and Beverly Sackler Dean's
+%%     Post-Doctoral fellowship.
+%%     \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
+%% \end{itemize}
+%% \newpage
 
@@
 
-For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (Section \ref{section:assumptions} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}).
+For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (Section \ref{subsec:assumptions} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}).
 A short calculation shows that this approximation implies the percentage increase in scheme false-negative rate is just $100\cdot \Se$.
@@
 Nevertheless, a substantial increase in pooling schemes' false-negative rates can be expected.
-Such a increase has crucial implications for controlling the spread of COVID-19.
+Such an increase has crucial implications for controlling the spread of COVID-19.
 

### 2021-03-16 comment.tex
New story, new structure

@@
 \documentclass{article}
-\usepackage{authblk}
-\usepackage{amsmath}
-\usepackage{todonotes}
-\usepackage{amssymb}
-\usepackage{subcaption}
-\usepackage[normalem]{ulem}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{marginnote}
-\usepackage[colorlinks=true,urlcolor=blue,citecolor=red,linkcolor=red,bookmarks=true]{hyperref}
-%strikethrough
-\usepackage{soul}
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.14}
-\usepackage{float}
-\usepackage[margin=0.5in]{geometry}
-\hypersetup{linkcolor=black,citecolor=black,urlcolor=black}
-
-\newcommand{\fp}{P_{\text{fp}}}
-\newcommand{\fn}{P_{\text{fn}}}
-\newcommand{\Se}{S_e}
-\newcommand{\Sp}{S_p}
-\newcommand{\mi}{P_{\text{sfn}}}
-\newcommand{\rel}{E_{\text{rel}}}
-\renewcommand\Affilfont{\itshape\small}
-\renewcommand{\Pr}{\mathbb{P}}
+\input{definitions}
 
 \begin{document}
-
-\title{Inflated false-negative rates in pooled RT-PCR tests of SARS-CoV-2}
+\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Tests Errors}
 
@@
 \begin{abstract}
-\textbf{Background}: Pooling is a popular strategy for increasing SARS-CoV-2 testing throughput.
-One popular pooling scheme is Dorfman pooling: test $N$ individuals simultaneously.
-If the test is positive --- retest each individual separately.
-However, requiring more than one positive test may lead to increased false-negative rates.
+\textbf{Background}: Pooling is a popular strategy for increasing
+SARS-CoV-2 testing throughput. One popular pooling scheme is Dorfman
+pooling: test $N$ individuals simultaneously. If the test is positive
+--- retest each individual separately. However, requiring more than
+one positive test may lead to increased false-negative rates.
 
-\textbf{Methods}: We analyze the false-negative rate (i.e., the probability of a negative result for an infected individual) of Dorfman pooling via a new probabilistic model.
-We demonstrate that different, previously made probabilistic assumptions regarding pooling are unlikely in light of empiric data.
-Our model is conservative in that it ignores sample dilution effects, which can only worsen pooling performance.
+\textbf{Methods}: We analyze the false-negative rate (i.e., the
+probability of a negative result for an infected individual) of
+Dorfman pooling via a new probabilistic model. We demonstrate that
+different, previously made probabilistic assumptions regarding pooling
+are unlikely in light of empiric data. Our model is conservative in
+that it ignores sample dilution effects, which can only worsen pooling
+performance.
 
-\textbf{Results}: We show that one can expect a 60-80\% increase in false-negative rates under Dorfman pooling, for reasonable parameter values.
-Moreover, we show that the false-negative rates under Dorfman pooling increase when the prevalence of infection decreases.
+\textbf{Results}: We show that one can expect a 60-80\% increase in
+false-negative rates under Dorfman pooling, for reasonable parameter
+values. Moreover, we show that the false-negative rates under Dorfman
+pooling increase when the prevalence of infection decreases.
 
-\textbf{Discussion}: In most pooling schemes, identifying an infected individual requires positive results in multiple tests and hence substantially increases false-negative rates.
-Furthermore, this phenomenon is more pronounced when infection prevalence is low --- exactly when pooling is most efficient.
-Thus, pooling presents an inherent trade-off: it is most efficient when it is least accurate.
-The deterioration of false-negative rates and the aforementioned trade-off are inherent problems of pooling schemes and should be kept in mind by practitioners and policy makers.
+\textbf{Discussion}: In most pooling schemes, identifying an infected
+individual requires positive results in multiple tests and hence
+substantially increases false-negative rates. Furthermore, this
+phenomenon is more pronounced when infection prevalence is low ---
+exactly when pooling is most efficient. Thus, pooling presents an
+inherent trade-off: it is most efficient when it is least
+accurate. The deterioration of false-negative rates and the
+aforementioned trade-off are inherent problems of pooling schemes and
+should be kept in mind by practitioners and policy makers.
 \end{abstract}
@@
 \section{Introduction}
+Reverse transcription polymerase chain reaction (RT-PCR) testing is a
+key component in breaking transmission chains and mitigating the
+COVID-19 pandemic. As such, the need for large-scale testing has
+resulted in the development techniques that of pooling schemes of RT-PCR tests
+\cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman, MatrixPooling,
+  LionDorfman}. One such popular scheme is Dorfman pooling
+\cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$ individuals and
+perform a single RT-PCR test on their combined (\emph{pooled})
+samples. If the pooled test yields a positive result --- test each
+individual separately. The throughput efficiency of Dorfman pooling
+has been demonstrated empirically \cite{DorfmanYuvalDor}.
 
-RT-PCR testing is a key component in breaking transmission chains and mitigating the COVID-19 pandemic.
-As such, the need for large-scale testing has resulted in the development of pooling schemes of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman, MatrixPooling, LionDorfman}.
-One such popular scheme is Dorfman pooling \cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$ individuals and perform a single RT-PCR test on their combined (``pooled'') samples.
-If the pooled test yields a positive result --- test each individual separately.
-The throughput efficiency of Dorfman pooling has been demonstrated empirically \cite{DorfmanYuvalDor}.
-However, when test error rates are taken into consideration, a sharp increase in false-negative rates can be expected.
+Pooling (or RT-PCR tests, for that matter) is not perfect, and there
+is always some nonzero probability of error. Tyipically, authors
+assume the probability of a true-positive (the test's
+\emph{sensitivity}) does not depend on the number of infected samples
+in the pool, but rather on the existence of at least one such
+sample. Thus, the probability of a positive result in a pooled test is
+(assumed) identical for a pool with one sample from an infected
+individual and (e.g.) five such samples. This assumption is common in
+the group testing literature \cite{Kim, OptimalDorfmanPool}, as well
+as in more specific, COVID-19 focused studies \cite{Simplistic1,
+  Simplistic2}. In this study we refute these commonly made
+assumptions and show how they affect estimation of false-negative
+rates for Dorfman pooling.
 
-It is important to distinguish three types of false-negative events when performing pooling.
-For convenience, we follow a single \emph{infected} individual, henceforth referred to as "Donald".
-A \emph{single test's} false-negative is the event of a negative result upon testing Donald separately, i.e., in an RT-PCR test without pooling.
-We denote the test sensitivity $\Se$, so the probability of a single test false-negative is $1-\Se$.
-A \emph{pooled} false-negative occurs when a pooled test containing Donald's sample (and other samples) yields a negative result, i.e., the pooling fails to detect at least one positive result.
-Lastly, a \emph{scheme} false-negative occurs when an entire pooling scheme fails to identify Donald as infected.
-Our goal is to calculate Dorfman's scheme false-negative rate.
-Rephrasing, we wish to answer the following question: what is the probability of not identifying Donald as infected under a Dorfman pooling scheme?
+%% Before we proceed, it is important to distinguish three types of
+%% false-negative events when performing pooling. For convenience, we
+%% follow a single \emph{infected} individual, henceforth referred to as
+%% "Donald". A \emph{single test's} false-negative is the event of a
+%% negative result upon testing Donald separately, i.e., in an RT-PCR
+%% test without pooling. We denote the test sensitivity $\Se$, so the
+%% probability of a single test false-negative is $1-\Se$. A
+%% \emph{pooled} false-negative occurs when a pooled test containing
+%% Donald's sample (and other samples) yields a negative result, i.e.,
+%% the pooling fails to detect at least one positive result. Lastly, a
+%% \emph{scheme} false-negative occurs when an entire pooling scheme
+%% fails to identify Donald as infected. Our goal is to calculate
+%% Dorfman's scheme false-negative rate. Rephrasing, we wish to answer
+%% the following question: what is the probability of not identifying
+%% Donald as infected under a Dorfman pooling scheme?
 
-\section{Methods}
-\subsection{Probabilistic Assumptions}\label{subsec:assumptions}
-We assume two pathways for a positive pooled test result: Viral RNA from an infected individual is correctly detected; or, some erroneous detection occurs (e.g.
-contaminant viral RNA is introduced).
-We ignore cross-reactivity with other Coronaviruses, which is negligible \cite{KitComparison}.
-We assume a homogeneous and disconnected population (each individual is infected independently and with equal probability).
-For simplicity, we do not take into account sample dilution, since it can only further increase false-negative rates \cite{DorfmanYuvalDor}.
+\subsection{RT-PCR}
+Before delving into our probabilistic assumptions, we briefly explain
+the process of testing for SARS-CoV-2 by RT-PCR. Our exposition
+intentionally avoids many details and a comprehensive review of RT-PCR
+can be found in \cite{COVID-RTPCR, PCRBook}.
 
-Our assumptions, although natural, stand in contrast to assumptions commonly made in the literature \cite{Kim, Simplistic1, Simplistic2, OptimalDorfmanPool}.
-It is commonly assumed that the probability of a positive pooled test is not increased by having more than one infected individual in the pool.
-We refute the common assumptions with experimental data summarized in table \ref{table}, collected from \cite{Salazar}.
-There, the authors investigate Dorfman pooling and, regardless of the pooled test result, follow up and test each pool member separately.
-Here, we focus on 128 pools for which at least one subsequent separate test was positive --- of which 29 pooled tests were negative and 99 positive.
-In the data cited in \cite{Salazar}, of the 29 negative pools, subsequent separate testing yielded a single positive result in 24.
-In contrast, of the 99 positive pools, 42 yielded a single positive test upon subsequent separate testing.
+PCR is a process in which a targeted DNA strand's frequency is
+amplified to create billions of copies in a reaction mixture. At the
+end of this \emph{amplification} process, the presence of targeted DNA
+molecules in a reaction mixture can be confidently
+determined. Arguably, the best way to understand PCR is as a chain
+reaction: Each targeted DNA molecule is copied into two, that are
+copied into four and so forth. This cascade gives PCR its name:
+polymerase chain reaction.
 
-The data in table \ref{table} allows us to test the following null hypothesis $H_0$: The probability of a pooled false-negative is equal for pools with one subsequent positively tested member and pools with two or more such members.
-We apply Fisher's exact test for the presence of more than one positive individual in correctly identified pools.
-Fisher's test yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a p-value $\approx 10^{-4}$.
-Thus we reject $H_0$, refuting the independence of a pooled result from one or more individuals infected within the pool, as assumed in \cite{Simplistic1, Simplistic2, OptimalDorfmanPool, Kim}
+Specifically, the PCR process is comprised of cycles, with each cycle
+doubling the abundance of the targeted DNA. In each cycle, each
+double-strand DNA molecule is broken into two separate strands. An
+enzyme called DNA polymerase generates a new double-strand DNA replica
+from each separate strand. Replication cannot start without a specific
+short DNA sequence, called a primer, that has to be introduced into
+the reaction mixture. Introducing the right primer into the reaction
+mixture ensures (almost) only the targeted DNA sequence is copied. A
+protein is added to the reaction mixture, which emits light upon
+successfully binding the targeted DNA. Once enough light is emitted
+the tested reaction mixture is declared positive and a
+\emph{detection} is said to occur. If, on the other hand, a predefined
+number of cycles pass without a detection event then the reaction
+mixture is declared negative.
+
+Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
+extra preprocessing step is required. An enzyme called
+reverse transcription (hence the RT in RT-PCR) replicates existing RNA
+in the reaction mixture into DNA. Once DNA replicas are made, the PCR
+process proceeds as described above, targeting SARS-CoV-2 DNA
+replicas.
+
+\section{RT-PCR errors}
+Failed detection of SARS-CoV-2 RNA in pooled RT-PCR testing is
+referred to below as a false-negative. One possible source of
+false-negatives in pooling is sample dilution. When pooling, several
+samples are mixed, so the concentration of viral RNA is reduced. This
+effect may cause a delay in amplification \cite{DorfmanYuvalDor}, no
+detection and, consequently, a false-negative. However, \cite{Lion}
+showed that this effect can be safely ignored when mixing up to 32
+samples, which we correspondingly assume.
+
+\subsection{Common assumptions}
+Previous studies of group testing strategies assumed that the
+false-negative probability does not depend on the number of infected
+samples, but merely on the existence of at least one such sample in a
+pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
+context of SARS-CoV-2 also employ similar assumptions
+\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
+that the probability of a negative result is the same for a pool with
+a single sample coming from an infected individual and (e.g.)
+five. We refute these assumptions and suggest more plausible ones in
+Section~\ref{subsec:ours} below.
+
+\subsection{Our model}\label{subsec:ours}
+We assume two pathways for a positive pooled test result: Viral RNA
+from an infected individual is correctly detected; or, some erroneous
+detection occurs (e.g. contaminant viral RNA is introduced). We assume
+a homogeneous and disconnected population (each individual is infected
+independently and with equal probability).
+
+Our assumptions, although natural, stand in contrast to assumptions
+commonly made in the literature \cite{Kim, Simplistic1, Simplistic2,
+  OptimalDorfmanPool}. It is commonly assumed that the probability of
+a positive pooled test is not increased by having more than one
+infected individual in the pool. We refute the common assumptions with
+experimental data summarized in Table \ref{table}, collected from
+\cite{Salazar}. There, the authors investigate Dorfman pooling and,
+regardless of the pooled test result, follow up and test each pool
+member separately. Here, we focus on 128 pools for which at least one
+subsequent separate test was positive --- of which 29 pooled tests
+were negative and 99 positive. In the data cited in \cite{Salazar}, of
+the 29 negative pools, subsequent separate testing yielded a single
+positive result in 24. In contrast, of the 99 positive pools, 42
+yielded a single positive test upon subsequent separate testing.
+
+The data in table \ref{table} allows us to test the following null
+hypothesis $H_0$: The probability of a pooled false-negative is equal
+for pools with one subsequent positively tested member and pools with
+two or more such members. We apply Fisher's exact test for the
+presence of more than one positive individual in correctly identified
+pools. Fisher's test yields an increased odds ratio of 6.4, 95\% CI
+(2.2,23.4), with a p-value $\approx 10^{-4}$. Thus we reject $H_0$,
+refuting the independence of a pooled result from one or more
+individuals infected within the pool, as assumed in \cite{Simplistic1,
+  Simplistic2, OptimalDorfmanPool, Kim}
 
@@
 
-\subsection{calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
-Denote the prevalence of infection in the (tested) population $q$.
-As before, $\Se$ is the test's sensitivity, so $1-\Se$ is the single RT-PCR test's false-negative rate.
-We also denote $\Sp$ the test specificity --- the true-negative probability.
-This encompasses events of an erroneous RNA detection (via any possible pathway), which may cause a false-positive.
-By our assumptions, a pool containing Donald's sample and $N-1$ other samples will yield a \emph{negative} result if all of the following occur:
+Erroneous detection of SARS-CoV-2 RNA (a false-positive) in pooled
+RT-PCR testing can also occur. A common assumption \cite{Simplistic1,
+  Simplistic2, Kim, OptimalDorfmanPool} is that the false-positive
+probability does not depend on the number of negative samples in a
+pool. We incorporate this assumption in our likelihood model with a
+small modification. We assume that an erroneous amplification can
+occur in \emph{any} pool. Specifically, it is possible that correct
+amplifications fail and an erroneous one occurs simultaneously. This
+assumption is relatively specific for the current application of
+screening for SARS-CoV-2 via RT-PCR.  For example, cross-reactivity
+with other coronaviruses would have violated this assumption, but it
+was ruled out in \cite{KitComparison}.
+
+\begin{figure}[H]
+  \centering
+  \begin{tikzpicture}
+    \begin{scope}[every node/.style={circle,thick,draw}]
+      \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
+      \node[shape=circle,draw=black] (B) at (2,0) {$\theta_2=0$};
+      \node[shape=circle,draw=black] (D) at (4,0) {$\theta_3=0$};
+      \node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};
+
+      \node[shape=ellipse,draw=black]
+      (F) at (9,0) {False-positive};
+
+      \node[shape=rectangle,draw=black] (G) at (4,7)
+           {$\Pr(\data=0|\theta)=\fn^2(1-\fp)$};
+    \end{scope}
+
+    \begin{scope}[>={Stealth[black]},
+        every node/.style={fill=white},
+        every edge/.style={draw=red,very thick}]
+      \path [->] (A) edge node {$\fn$} (4.5,6.6);
+      %\path [->] (B) edge node {$0$} (G);
+      %\path [->] (D) edge node {$0$} (G);
+      \path [->] (E) edge node {$\fn$} (4.5,6.6);
+      \path [->] (F) edge node {$1-\fp$} (5.5,6.6); 
+    \end{scope}
+  \end{tikzpicture}
+  \caption{Illustration of the likelihood model of a pooled test
+    result. A pool contains individuals $\{1,2,3,4\}$ with state
+    $\theta=(1,0,0,1)$ (i.e. individuals 1 and 4 are infected). A
+    negative pooled test implies that three detection paths failed. A
+    false negative occurred for $1$ and $4$, each with probability
+    $\fn$. Additionally, no erroneous detection of SARS-CoV-2
+    occurred, with probability $1-\fp$. Individuals $2$ and $3$ are
+    not infected and do not contribute to the probability of the
+    pooled test result. A positive pooled test arises if any one of
+    the above mentioned paths results in a
+    detection.}\label{fig:likelihood}
+\end{figure}
+
+To summarize, we assume that for a single pool, a positive test result
+is generated in one of two paths. Either, SARS-CoV-2 RNA from an
+infected individual's sample is correctly amplified and detected, and
+this can happen for each positive sample in a pool. Or, some erroneous
+amplification occurs (e.g. contaminant viral RNA is introduced), an
+event that occurs at most once per pool. Our model is illustrated in
+Figure~\ref{fig:likelihood}, and explicitly defined below.
+
+Consider a pool containing $N$ individuals $\{1,\dots,N\}$. We denote
+the true infection states $\theta \in \{0,1\}^N$, so individual $i$ is
+infected iff $\theta_i=1$. Data are denoted $\data$. The tests
+false-negative rate is denoted $\fn$ and the false-positve rate is
+denoted $\fp$. The probability of a negative pooled test result is
+presented in \eqref{eq:likelihood a} (along with its complement
+\eqref{eq:likelihood b}), and explained below.
+\begin{subequations}
+\begin{align}%\label{eq:single likelihood}
+  %\begin{split}
+    \Pr(\data=0 | \theta) &= (1-\fp)\prod_{i=1}^N
+    \fn^{\theta_i}, \label{eq:likelihood a}\\
+    %
+    %
+    %
+    \Pr(\data=1 | \theta) &= 1 - (1-\fp)\prod_{i=1}^N \fn^{\theta_i}. \label{eq:likelihood b}
+  %\end{split}
+\end{align}
+\end{subequations}
+A negative pooled test result occurs when all detection paths (both
+correct and erroneous) fail. The probability of no false-detection
+accounts for the $1-\fp$ term. The probability of no correct detection
+is $\fn$ per infected individual. The probability that all such paths
+fail is the product of the above mentioned terms, displayed in
+\eqref{eq:likelihood a}. The probability of a positive result,
+presented in \eqref{eq:likelihood b}, is simply the complement.
+Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
+recalling that $\data \in \{0,1\}$ yields:
+
+\begin{align}\label{eq:combined likelihood}
+  \begin{split}
+    \Pr(\data| \theta) &= \left [(1-\fp)\prod_{i=1}^N
+      \fn^{\theta_i} \right ]^{1-\data} \left [1 - (1-\fp)\prod_{i=1}^N \fn^{\theta_i} \right ]^{\data}.
+  \end{split}
+\end{align}
+
+\section{Application: Calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
+As an application of our model, we present the differnece in scheme
+false-negative rates (...) Denote the prevalence of infection in the
+(tested) population $q$. As before, $\Se$ is the test's sensitivity,
+so $1-\Se$ is the single RT-PCR test's false-negative rate. We also
+denote $\Sp$ the test specificity --- the true-negative
+probability. This encompasses events of an erroneous RNA detection
+(via any possible pathway), which may cause a false-positive. By our
+assumptions, a pool containing Donald's sample and $N-1$ other samples
+will yield a \emph{negative} result if all of the following occur:
 \begin{itemize}
-\item No erroneous detection occurred (i.e. no false-positive). This happens with probability $\Sp$.
-\item The detection process fails for Donald's sample. A false-negative occurs for Donald, with probability $1-\Se$.
-\item No detection for any of the other $N-1$ samples.
-For a single sample, the probability of being detected is the prevalence of SARS-CoV-2 in the tested population $q$, multiplied by the sensitivity $\Se$, hence the probability of detection is $q\Se$.
-For $N-1$ such samples, the probability of \emph{not} being identified is $(1-q\Se)^{N-1}$.
+\item No erroneous detection occurred (i.e. no false-positive). This
+  happens with probability $\Sp$.
+\item The detection process fails for Donald's sample. A
+  false-negative occurs for Donald, with probability $1-\Se$.
+\item No detection for any of the other $N-1$ samples. For a single
+  sample, the probability of being detected is the prevalence of
+  SARS-CoV-2 in the tested population $q$, multiplied by the
+  sensitivity $\Se$, hence the probability of detection is $q\Se$. For
+  $N-1$ such samples, the probability of \emph{not} being identified
+  is $(1-q\Se)^{N-1}$.
 \end{itemize}
@@
     \begin{split}
-        \Pr(\text{pool is positive}) &= 1 - \Pr(\text{pool false-negative}) \\
+        \Pr(\data=1|\theta_1=1) &= 1 - \Pr(\data=0|\theta_1=1) \\
         &= 1 - \Sp(1-\Se)(1-q\Se)^{N-1}.
@@
 
-If the pooled test yields a positive result, Donald is tested separately.
-We assume such a simple procedure poses no risk of introducing contaminant RNA.
-Therefore, the separate test yields a positive result with probability $\Se$.
+If the pooled test yields a positive result, Donald is tested
+separately. We assume such a simple procedure poses no risk of
+introducing contaminant RNA. Therefore, the separate test yields a
+positive result with probability $\Se$.
 
-We calculate the probability that Donald is mistakenly identified as not infected --- the scheme's false-negative rate --- denoted $\mi$ below.
-To \emph{correctly} identify Donald as infected, both pooled and separate tests have to yield a positive result.
-Thus, the scheme's false-negative rate $\mi$ is the complement of the product of the two previous terms:
+We calculate the probability that Donald is mistakenly identified as
+not infected --- the scheme's false-negative rate --- denoted $\mi$
+below. To \emph{correctly} identify Donald as infected, both pooled
+and separate tests have to yield a positive result. Thus, the scheme's
+false-negative rate $\mi$ is the complement of the product of the two
+previous terms:
 \begin{align}\label{eq:sfn}
@@
 \subsection{Comparison metric}
-The single test false-negative rate $1-\Se$ and scheme false-negative rate $\mi$ are compared via:
+The single test false-negative rate $1-\Se$ and scheme false-negative
+rate $\mi$ are compared via:
 \begin{equation}\label{eq:erel}
@@
 \end{equation}
-$\rel$ is the percentage increase in the pooling scheme false-negative rate, relative to the single test false-negative rate.
+$\rel$ is the percentage increase in the pooling scheme false-negative
+rate, relative to the single test false-negative rate.
 
-For the approximation commonly used in the literature estimates the scheme false-negative rate as $1-\Se^2$ (Section \ref{subsec:assumptions} and \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}).
-A short calculation shows that this approximation implies the percentage increase in scheme false-negative rate is just $100\cdot \Se$.
+For the approximation commonly used in the literature estimates the
+scheme false-negative rate as $1-\Se^2$ (Section
+\ref{subsec:assumptions} and \cite{Simplistic1, Simplistic2, Kim,
+  OptimalDorfmanPool}). A short calculation shows that this
+approximation implies the percentage increase in scheme false-negative
+rate is just $100\cdot \Se$.
 
-\section{Results}\label{section:results}
+\subsection{Results}\label{section:results}
 
-We plot $\rel$ for varying prevalence $q$ and sensitivity $\Se$ values.
-As recommended by \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for different prevalence values.
-We observe that for a false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable sensitivity and prevalence values \cite{KitComparison, InterpretingCOVID19Test, EstimatingRatesLourenco, FalsePositiveEstimate}, an increase of at least $60\%$ in $\rel$ can be expected (Figure \ref{figy}).
-Interestingly, an increase in infection prevalence monotonically decreases the scheme false-negative rate, as can also be easily seen from equation \eqref{eq:sfn}.
-For the chosen parameter ranges, the increase in the single test false-negative rates increases the relative error $\rel$.
-These effects can be seen in Figure \ref{figy} (left panel), upon conditioning on pool size.
-Extending the range for $\Sp$ yields no qualitative differences.
-We further compare $\rel$ to the commonly used approximation, showing the discrepancy changes as a function of both prevalence and the single test sensitivity (Figure \ref{figy}, right panel)
+We plot $\rel$ for varying prevalence $q$ and sensitivity $\Se$
+values. As recommended by \cite{DorfmanYuvalDor}, we apply different
+pool sizes $N$, for different prevalence values. We observe that for a
+false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of
+reasonable sensitivity and prevalence values \cite{KitComparison,
+  InterpretingCOVID19Test, EstimatingRatesLourenco,
+  FalsePositiveEstimate}, an increase of at least $60\%$ in $\rel$ can
+be expected (Figure \ref{figy}). Interestingly, an increase in
+infection prevalence monotonically decreases the scheme false-negative
+rate, as can also be easily seen from equation \eqref{eq:sfn}. For the
+chosen parameter ranges, the increase in the single test
+false-negative rates increases the relative error $\rel$. These
+effects can be seen in Figure \ref{figy} (left panel), upon
+conditioning on pool size. Extending the range for $\Sp$ yields no
+qualitative differences. We further compare $\rel$ to the commonly
+used approximation, showing the discrepancy changes as a function of
+both prevalence and the single test sensitivity (Figure \ref{figy},
+right panel)
 \begin{figure}[H]
@@
     \includegraphics[width=18cm]{heatmap.jpg}
-    \caption{Relative increase in Dorfman pooling false-negative rates $\rel$.
-Left: Colors represent $\rel$, the relative percentage increase in the scheme false-negative rates relative to the single test false-negative rates (eq \eqref{eq:erel}).
-Right: colors represent the difference between the approximation assumed in other studies and $\rel$ as calculated by our model.
-The disease prevalence $q$, is varied on the x-axis, while the test sensitivity is varied on the y-axis.
-Pool size $N$, was chosen according to $q$ as in \cite{DorfmanYuvalDor}.}\label{figy}
+    \caption{Relative increase in Dorfman pooling false-negative rates
+      $\rel$. Left: Colors represent $\rel$, the relative percentage
+      increase in the scheme false-negative rates relative to the
+      single test false-negative rates (eq \eqref{eq:erel}). Right:
+      colors represent the difference between the approximation
+      assumed in other studies and $\rel$ as calculated by our
+      model. The disease prevalence $q$, is varied on the x-axis,
+      while the test sensitivity is varied on the y-axis. Pool size
+      $N$, was chosen according to $q$ as in
+      \cite{DorfmanYuvalDor}.}\label{figy}
  \end{figure}
@@
 \section{Discussion}
-Although Dorfman pooling improves testing throughput, we have shown that it can increase $\rel$ --- the false-negative rates relative to individual testing.
-Furthermore, low values of infection prevalence, or low values of single test false-negative rates, increase $\rel$.
-These results remain qualitatively similar under varying parameter values, in the observed ranges \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure \ref{figy}).
+Although Dorfman pooling improves testing throughput, we have shown
+that it can increase $\rel$ --- the false-negative rates relative to
+individual testing. Furthermore, low values of infection prevalence,
+or low values of single test false-negative rates, increase
+$\rel$. These results remain qualitatively similar under varying
+parameter values, in the observed ranges
+\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
+  InterpretingCOVID19Test} (Figure \ref{figy}).
 
-Our results lead to a fact almost disregarded in \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient when prevalence is low, such circumstances are exactly those leading to a substantial increase in false-negative rates.
+Our results lead to a fact almost disregarded in
+\cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
+when prevalence is low, such circumstances are exactly those leading
+to a substantial increase in false-negative rates.
 
-An approximation to our results has been previously considered: $\mi \approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}.
-Such an approximation arises when one assumes that the probability of a false-negative is identical for the pooled and single tests.
-Although simplistic in nature, this approximation does capture the intuition behind our results: For Donald to be considered negative under Dorfman pooling, he has to test negative twice.
-In addition to this approximation only being an upper bound \cite{Simplistic2}, it does not account for the effect of infection prevalence on the false-negative rates under Dorfman pooling.
-Supporting the importance of this effect, such an association between prevalence and the scheme false-negative rate under Dorfman pooling has been empirically noted  \cite{DorfmanYuvalDor}.
+An approximation to our results has been previously considered: $\mi
+\approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
+  OptimalDorfmanPool}. Such an approximation arises when one assumes
+that the probability of a false-negative is identical for the pooled
+and single tests. Although simplistic in nature, this approximation
+does capture the intuition behind our results: For Donald to be
+considered negative under Dorfman pooling, he has to test negative
+twice. In addition to this approximation only being an upper bound
+\cite{Simplistic2}, it does not account for the effect of infection
+prevalence on the false-negative rates under Dorfman
+pooling. Supporting the importance of this effect, such an association
+between prevalence and the scheme false-negative rate under Dorfman
+pooling has been empirically noted \cite{DorfmanYuvalDor}.
 
-Although we have shown the inherent risk of Dorfman pooling, this shortcoming applies to other pooling schemes.
-Pooling schemes (e.g.
-\cite{MatrixPooling,Lion, Kim}), require some sequence of positive pooled results to correctly identify Donald as infected.
-Consider the hierarchical pooling scheme \cite{Lion, Kim}: If the first pool yields a positive result, it is split in two.
-Then the splitting is repeated until resulting pools are negative or individuals are tested separately.
-With an initial pool size of 32, Donald will necessarily have to test positive in pools of size 32, 16, 8, 4 and 2, as well as in a single test, for the scheme to correctly identify him as infected.
-Compare this to the Dorfman scheme that requires a positive test in a pool of size $N=8$, and an additional single positive test to identify Donald as infected.
-The hierarchical pooling scheme of \cite{Lion, Kim} will necessarily yield  more false-negatives than Dorfman pooling --- there are additional places for it to fail.
+Although we have shown the inherent risk of Dorfman pooling, this
+shortcoming applies to other pooling schemes. Pooling schemes
+(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
+positive pooled results to correctly identify Donald as
+infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
+If the first pool yields a positive result, it is split in two. Then
+the splitting is repeated until resulting pools are negative or
+individuals are tested separately. With an initial pool size of 32,
+Donald will necessarily have to test positive in pools of size 32, 16,
+8, 4 and 2, as well as in a single test, for the scheme to correctly
+identify him as infected. Compare this to the Dorfman scheme that
+requires a positive test in a pool of size $N=8$, and an additional
+single positive test to identify Donald as infected. The hierarchical
+pooling scheme of \cite{Lion, Kim} will necessarily yield more
+false-negatives than Dorfman pooling --- there are additional places
+for it to fail.
 
-As mentioned in \cite{DorfmanYuvalDor}, introducing a positive dependence within a pool decreases the false-positive rate.
-In the extreme case, consider a fully connected pool, where one infection implies the entire pool is infected.
-In this case, a calculation analogous to the one conducted above recovers the initial false-negative rate $1-\Se$.
-Interestingly, pooling was also noted to have increased throughput when infection probabilities are dependent between the pooled individuals \cite{DorfmanYuvalDor}, providing  another advantage to sampling dependent individuals in pooling schemes.
+As mentioned in \cite{DorfmanYuvalDor}, introducing a positive
+dependence within a pool decreases the false-positive rate. In the
+extreme case, consider a fully connected pool, where one infection
+implies the entire pool is infected. In this case, a calculation
+analogous to the one conducted above recovers the initial
+false-negative rate $1-\Se$. Interestingly, pooling was also noted to
+have increased throughput when infection probabilities are dependent
+between the pooled individuals \cite{DorfmanYuvalDor}, providing
+another advantage to sampling dependent individuals in pooling
+schemes.
 
-To conclude, pooling is an important technique which can improve testing throughput in a cost-effective manner.
-Nevertheless, a substantial increase in pooling schemes' false-negative rates can be expected.
-Such an increase has crucial implications for controlling the spread of COVID-19.
+To conclude, pooling is an important technique which can improve
+testing throughput in a cost-effective manner. Nevertheless, a
+substantial increase in pooling schemes' false-negative rates can be
+expected. Such an increase has crucial implications for controlling
+the spread of COVID-19.
 

### 2021-03-16 comment.tex
better

@@
 COVID-19 pandemic. As such, the need for large-scale testing has
-resulted in the development techniques that of pooling schemes of RT-PCR tests
-\cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman, MatrixPooling,
-  LionDorfman}. One such popular scheme is Dorfman pooling
-\cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$ individuals and
-perform a single RT-PCR test on their combined (\emph{pooled})
-samples. If the pooled test yields a positive result --- test each
-individual separately. The throughput efficiency of Dorfman pooling
-has been demonstrated empirically \cite{DorfmanYuvalDor}.
+resulted in the development of techniques to increase the the
+throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
+  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques are
+called \emph{pooling} and originated in the seminal work of Dorfman
+pooling \cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$
+individuals and perform a single RT-PCR test on their combined
+(\emph{pooled}) samples. If the pooled test yields a positive result
+--- test each individual separately. The throughput efficiency of
+Dorfman pooling for has been demonstrated empirically
+\cite{DorfmanYuvalDor} and its error rates thoroguhly investigated
+\cite{Kim, Simplistic1, OptimalDorfmanPool}.
 
-Pooling (or RT-PCR tests, for that matter) is not perfect, and there
-is always some nonzero probability of error. Tyipically, authors
-assume the probability of a true-positive (the test's
-\emph{sensitivity}) does not depend on the number of infected samples
-in the pool, but rather on the existence of at least one such
-sample. Thus, the probability of a positive result in a pooled test is
-(assumed) identical for a pool with one sample from an infected
-individual and (e.g.) five such samples. This assumption is common in
-the group testing literature \cite{Kim, OptimalDorfmanPool}, as well
-as in more specific, COVID-19 focused studies \cite{Simplistic1,
-  Simplistic2}. In this study we refute these commonly made
-assumptions and show how they affect estimation of false-negative
-rates for Dorfman pooling.
+Pooling RT-PCR tests is not perfect, and there is always some nonzero
+probability of error. Typically, authors assume the probability of a
+true-positive (the test's \emph{sensitivity}) does not depend on the
+number of infected samples in the pool, but rather on the existence of
+at least one such sample. Thus, the probability of a positive result
+in a pooled test is (assumed) identical for a pool with one sample
+from an infected individual and (e.g.) five such samples. This
+assumption is common in the group testing literature \cite{Kim,
+  OptimalDorfmanPool}, as well as in more specific, COVID-19 focused
+studies \cite{Simplistic1, Simplistic2}. In this study we refute these
+commonly made assumptions and show how they affect estimation of
+false-negative rates for Dorfman pooling.
 
@@
 
+Formally, we consider a pool containing $N$ individuals
+$\{1,\dots,N\}$. We denote the true infection states $\theta \in
+\{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
+test's sensitivity (true-negative rate) is denoted $\sens$, and the
+test's specificity (true-negative rate) is denoted $\spec$. Test results
+(data) are denoted $\data$, where $\data=1$ iff the test returned a
+positive result.
+
 \subsection{Common assumptions}
@@
 a single sample coming from an infected individual and (e.g.)
-five. We refute these assumptions and suggest more plausible ones in
-Section~\ref{subsec:ours} below.
+five. Explicitly:
 
-\subsection{Our model}\label{subsec:ours}
-We assume two pathways for a positive pooled test result: Viral RNA
-from an infected individual is correctly detected; or, some erroneous
-detection occurs (e.g. contaminant viral RNA is introduced). We assume
-a homogeneous and disconnected population (each individual is infected
-independently and with equal probability).
+\begin{align}\label{eq:common}
+\Pr(\data=1|\theta) = \left \{
+\begin{array}{lr}
+        \sens & \exists i\ \text{such that } \theta_i = 1\\
+        1-\spec & \text{otherwise}
+\end{array} 
+\end{align}
+We refute the assumptions underlying \eqref{eq:common} in
+Section~\ref{subsec:ours} and suggest more plausible ones in
+Section~\ref{subsec:ours}.
 
-Our assumptions, although natural, stand in contrast to assumptions
-commonly made in the literature \cite{Kim, Simplistic1, Simplistic2,
-  OptimalDorfmanPool}. It is commonly assumed that the probability of
-a positive pooled test is not increased by having more than one
-infected individual in the pool. We refute the common assumptions with
+\subsection{Refuting common assumptions}\label{subsec:refute}
+We refute the common assumptions of Section~\ref{subsec:common} with
 experimental data summarized in Table \ref{table}, collected from
@@
 
-The data in table \ref{table} allows us to test the following null
+The data in Table \ref{table} allows us to test the following null
 hypothesis $H_0$: The probability of a pooled false-negative is equal
@@
 
-Consider a pool containing $N$ individuals $\{1,\dots,N\}$. We denote
-the true infection states $\theta \in \{0,1\}^N$, so individual $i$ is
-infected iff $\theta_i=1$. Data are denoted $\data$. The tests
-false-negative rate is denoted $\fn$ and the false-positve rate is
-denoted $\fp$. The probability of a negative pooled test result is
-presented in \eqref{eq:likelihood a} (along with its complement
+The probability of a negative pooled test result is presented in
+\eqref{eq:likelihood a} (along with its complement
 \eqref{eq:likelihood b}), and explained below.

### 2021-03-16 comment.tex
Better even more gooder

@@
 
-%% Before we proceed, it is important to distinguish three types of
-%% false-negative events when performing pooling. For convenience, we
-%% follow a single \emph{infected} individual, henceforth referred to as
-%% "Donald". A \emph{single test's} false-negative is the event of a
-%% negative result upon testing Donald separately, i.e., in an RT-PCR
-%% test without pooling. We denote the test sensitivity $\Se$, so the
-%% probability of a single test false-negative is $1-\Se$. A
-%% \emph{pooled} false-negative occurs when a pooled test containing
-%% Donald's sample (and other samples) yields a negative result, i.e.,
-%% the pooling fails to detect at least one positive result. Lastly, a
-%% \emph{scheme} false-negative occurs when an entire pooling scheme
-%% fails to identify Donald as infected. Our goal is to calculate
-%% Dorfman's scheme false-negative rate. Rephrasing, we wish to answer
-%% the following question: what is the probability of not identifying
-%% Donald as infected under a Dorfman pooling scheme?
-
 \subsection{RT-PCR}
@@
 test's specificity (true-negative rate) is denoted $\spec$. Test results
-(data) are denoted $\data$, where $\data=1$ iff the test returned a
+(data) are denoted $\data$, where $\data=0$ iff the test returned a
 positive result.
@@
 \begin{align}\label{eq:common}
-\Pr(\data=1|\theta) = \left \{
+\Pr(\data=0|\theta) = \left \{
 \begin{array}{lr}
-        \sens & \exists i\ \text{such that } \theta_i = 1\\
-        1-\spec & \text{otherwise}
+        1-\sens & \exists i\ \text{such that } \theta_i = 1\\
+        \spec & \text{otherwise}
 \end{array} 
@@
 
-Erroneous detection of SARS-CoV-2 RNA (a false-positive) in pooled
-RT-PCR testing can also occur. A common assumption \cite{Simplistic1,
-  Simplistic2, Kim, OptimalDorfmanPool} is that the false-positive
-probability does not depend on the number of negative samples in a
-pool. We incorporate this assumption in our likelihood model with a
-small modification. We assume that an erroneous amplification can
-occur in \emph{any} pool. Specifically, it is possible that correct
-amplifications fail and an erroneous one occurs simultaneously. This
-assumption is relatively specific for the current application of
-screening for SARS-CoV-2 via RT-PCR.  For example, cross-reactivity
-with other coronaviruses would have violated this assumption, but it
-was ruled out in \cite{KitComparison}.
+\subsection{Our model}
+We assume two pathways for a positive pooled test result: Viral RNA
+from an infected individual is correctly detected; or, some erroneous
+detection occurs (e.g. contaminant viral RNA is introduced). We assume
+a homogeneous and disconnected population (each individual is infected
+independently and with equal probability. Erroneous detection of
+SARS-CoV-2 RNA (a false-positive) in pooled RT-PCR testing can also
+occur. A common assumption \cite{Simplistic1, Simplistic2, Kim,
+  OptimalDorfmanPool} is that the false-positive probability does not
+depend on the number of negative samples in a pool. We incorporate
+this assumption in our likelihood model with a small modification. We
+assume that an erroneous amplification can occur in \emph{any}
+pool. Specifically, it is possible that correct amplifications fail
+and an erroneous one occurs simultaneously. This assumption is
+relatively specific for the current application of screening for
+SARS-CoV-2 via RT-PCR.  For example, cross-reactivity with other
+coronaviruses would have violated this assumption, but it was ruled
+out in \cite{KitComparison}.
 
@@
       \node[shape=rectangle,draw=black] (G) at (4,7)
-           {$\Pr(\data=0|\theta)=\fn^2(1-\fp)$};
+           {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
     \end{scope}
@@
         every edge/.style={draw=red,very thick}]
-      \path [->] (A) edge node {$\fn$} (4.5,6.6);
+      \path [->] (A) edge node {$1-\sens$} (4.5,6.6);
       %\path [->] (B) edge node {$0$} (G);
       %\path [->] (D) edge node {$0$} (G);
-      \path [->] (E) edge node {$\fn$} (4.5,6.6);
-      \path [->] (F) edge node {$1-\fp$} (5.5,6.6); 
+      \path [->] (E) edge node {$1-\sens$} (4.5,6.6);
+      \path [->] (F) edge node {$\spec$} (5.5,6.6); 
     \end{scope}
@@
     false negative occurred for $1$ and $4$, each with probability
-    $\fn$. Additionally, no erroneous detection of SARS-CoV-2
-    occurred, with probability $1-\fp$. Individuals $2$ and $3$ are
+    $(1-\sens)$. Additionally, no erroneous detection of SARS-CoV-2
+    occurred, with probability $\spec$. Individuals $2$ and $3$ are
     not infected and do not contribute to the probability of the
@@
 
-The probability of a negative pooled test result is presented in
-\eqref{eq:likelihood a} (along with its complement
-\eqref{eq:likelihood b}), and explained below.
-\begin{subequations}
+A negative pooled test result occurs when all detection paths (both
+correct and erroneous) fail. The probability of no false-detection
+accounts for the $\spec$ term. The probability of no correct detection
+is $(1-\sens)$ per infected individual. The probability that all such
+paths fail is the product of the above mentioned terms, displayed in
+\eqref{eq:likelihood a}.%%  The probability of a positive result,
+%% presented in \eqref{eq:likelihood b}, is simply the complement.
+%% The probability of a negative pooled test result is presented in
+%% \eqref{eq:likelihood a} (along with its complement
+%% \eqref{eq:likelihood b}), and explained below.
+%\begin{subequations}
 \begin{align}%\label{eq:single likelihood}
   %\begin{split}
-    \Pr(\data=0 | \theta) &= (1-\fp)\prod_{i=1}^N
-    \fn^{\theta_i}, \label{eq:likelihood a}\\
+    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
+    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum\theta_i}, \label{eq:likelihood a}
     %
@@
     %
-    \Pr(\data=1 | \theta) &= 1 - (1-\fp)\prod_{i=1}^N \fn^{\theta_i}. \label{eq:likelihood b}
+    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
+    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
   %\end{split}
 \end{align}
-\end{subequations}
-A negative pooled test result occurs when all detection paths (both
-correct and erroneous) fail. The probability of no false-detection
-accounts for the $1-\fp$ term. The probability of no correct detection
-is $\fn$ per infected individual. The probability that all such paths
-fail is the product of the above mentioned terms, displayed in
-\eqref{eq:likelihood a}. The probability of a positive result,
-presented in \eqref{eq:likelihood b}, is simply the complement.
-Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
-recalling that $\data \in \{0,1\}$ yields:
+%\end{subequations}
+%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
+%% recalling that $\data \in \{0,1\}$ yields:
 
-\begin{align}\label{eq:combined likelihood}
-  \begin{split}
-    \Pr(\data| \theta) &= \left [(1-\fp)\prod_{i=1}^N
-      \fn^{\theta_i} \right ]^{1-\data} \left [1 - (1-\fp)\prod_{i=1}^N \fn^{\theta_i} \right ]^{\data}.
-  \end{split}
-\end{align}
+%% \begin{align}\label{eq:combined likelihood}
+%%   \begin{split}
+%%     \Pr(\data| \theta) &= \left [\spec\prod_{i=1}^N
+%%       (1-\sens)^{\theta_i} \right ]^{1-\data} \left [1 - \spec\prod_{i=1}^N (1-\sens)^{\theta_i} \right ]^{\data}.
+%%   \end{split}
+%% \end{align}
 
 \section{Application: Calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
-As an application of our model, we present the differnece in scheme
-false-negative rates (...) Denote the prevalence of infection in the
-(tested) population $q$. As before, $\Se$ is the test's sensitivity,
-so $1-\Se$ is the single RT-PCR test's false-negative rate. We also
-denote $\Sp$ the test specificity --- the true-negative
-probability. This encompasses events of an erroneous RNA detection
-(via any possible pathway), which may cause a false-positive. By our
-assumptions, a pool containing Donald's sample and $N-1$ other samples
-will yield a \emph{negative} result if all of the following occur:
-\begin{itemize}
-\item No erroneous detection occurred (i.e. no false-positive). This
-  happens with probability $\Sp$.
-\item The detection process fails for Donald's sample. A
-  false-negative occurs for Donald, with probability $1-\Se$.
-\item No detection for any of the other $N-1$ samples. For a single
-  sample, the probability of being detected is the prevalence of
-  SARS-CoV-2 in the tested population $q$, multiplied by the
-  sensitivity $\Se$, hence the probability of detection is $q\Se$. For
-  $N-1$ such samples, the probability of \emph{not} being identified
-  is $(1-q\Se)^{N-1}$.
-\end{itemize}
-The pooled false-negative probability for Donald is simply the product of the terms above. Hence:
+As an application of our model, we present the differnece in
+calculated false-negative rate for a single \emph{infected}
+individual, henceforth referred to as "Donald".  We distinguish three
+types of false-negative events when performing pooling. A \emph{single
+  test's} false-negative is the event of a negative result upon
+testing Donald separately, i.e., in an RT-PCR test without pooling. A
+\emph{pooled} false-negative occurs when a pooled test containing
+Donald's sample (and other samples) yields a negative result, i.e.,
+the pooling fails to detect at least one positive result. Lastly, a
+\emph{scheme} false-negative occurs when an entire pooling scheme
+fails to identify Donald as infected. Our goal is to calculate
+Dorfman's scheme false-negative rate. Rephrasing, we wish to answer
+the following question: what is the probability of not identifying
+Donald as infected under a Dorfman pooling scheme?
+
+Denote the prevalence of infection in the (tested) population $q$. We
+give Donald the respect he deserves and assign him as individual $1$,
+so that $\theta_1=1$. Then:
 
@@
   \begin{split}
-        \Pr(\data=1|\theta_1=1) &= 1 - \Pr(\data=0|\theta_1=1) \\
-        &= 1 - \Sp(1-\Se)(1-q\Se)^{N-1}.
+    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta') \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
+    %
+    %
+    %
+    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)\Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
+    %
+    %
+    %
+    &= \sum_{k=0}^{N-1}\binom{N-1}{k}q^k(1-q)^{N-1-k} \Sp(1-\Se)^{1+k}\\
+    %
+    %
+    %
+    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}\left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
+    %
+    %
+    %
+    &= \Sp(1-\Se)(1-q\Se)^{N-1}, \text{ by Newton's Binomial Theorem.} \\
   \end{split}
@@
 If the pooled test yields a positive result, Donald is tested
-separately. We assume such a simple procedure poses no risk of
-introducing contaminant RNA. Therefore, the separate test yields a
-positive result with probability $\Se$.
+separately. Taking a conservative stand, we assume such a simple
+procedure poses no risk of introducing contaminant RNA. Therefore, the
+separate test yields a positive result with probability $\Se$.
 
@@
     \begin{split}
-        \mi :&= 1-\Pr(\text{correctly identify Donald as infected})\\
+        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
         %
@@
 
-For the approximation commonly used in the literature estimates the
-scheme false-negative rate as $1-\Se^2$ (Section
-\ref{subsec:assumptions} and \cite{Simplistic1, Simplistic2, Kim,
-  OptimalDorfmanPool}). A short calculation shows that this
-approximation implies the percentage increase in scheme false-negative
-rate is just $100\cdot \Se$.
+The approximation commonly used in the literature estimates the scheme
+false-negative rate as $1-\Se^2$ (Section \ref{subsec:assumptions} and
+\cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}). A short
+calculation shows that this approximation implies the percentage
+increase in scheme false-negative rate is just $100\cdot \Se$.
 
@@
 
-\bibliographystyle{ama}
+\bibliographystyle{amsplain}
+%\bibliographystyle{ama}
 \bibliography{refs}

### 2021-03-16 comment.tex
More words

@@
 infected individual's sample is correctly amplified and detected, and
-this can happen for each positive sample in a pool. Or, some erroneous
-amplification occurs (e.g. contaminant viral RNA is introduced), an
-event that occurs at most once per pool. Our model is illustrated in
-Figure~\ref{fig:likelihood}, and explicitly defined below.
-
-A negative pooled test result occurs when all detection paths (both
-correct and erroneous) fail. The probability of no false-detection
-accounts for the $\spec$ term. The probability of no correct detection
-is $(1-\sens)$ per infected individual. The probability that all such
-paths fail is the product of the above mentioned terms, displayed in
+this can happen for each positive sample in a pool, with probability
+$\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
+RNA is introduced), an event that occurs at most once per pool, with
+probability $1-\Sp$. Our model is illustrated in
+Figure~\ref{fig:likelihood}, and summarized in
+%% A negative pooled test result occurs when all detection paths (both
+%% correct and erroneous) fail. The probability of no false-detection
+%% accounts for the $\spec$ term. The probability of no correct detection
+%% is $(1-\sens)$ per infected individual. The probability that all such
+%% paths fail is the product of the above mentioned terms, displayed in
 \eqref{eq:likelihood a}.%%  The probability of a positive result,
@@
     \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
-    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum\theta_i}, \label{eq:likelihood a}
+    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum\theta_i}. \label{eq:likelihood a}
     %

### 2021-03-17 comment.tex
worz

@@
 
-\subsection{RT-PCR}
+\section{Methods}
 Before delving into our probabilistic assumptions, we briefly explain
-the process of testing for SARS-CoV-2 by RT-PCR. Our exposition
-intentionally avoids many details and a comprehensive review of RT-PCR
-can be found in \cite{COVID-RTPCR, PCRBook}.
+the process of testing for SARS-CoV-2 by RT-PCR in
+Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
+details and a comprehensive review of RT-PCR can be found in
+\cite{COVID-RTPCR, PCRBook}. We then present the \emph{common
+  assumption} in Section~\ref{subsec:common} and refute it in
+Section~\ref{subsec:refute}. Our model is presented in
+Section~\ref{subsec:ours}.
 
+\subsection{RT-PCR}\label{subsec:rtpcr}
 PCR is a process in which a targeted DNA strand's frequency is
@@
 Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
-extra preprocessing step is required. An enzyme called
-reverse transcription (hence the RT in RT-PCR) replicates existing RNA
-in the reaction mixture into DNA. Once DNA replicas are made, the PCR
-process proceeds as described above, targeting SARS-CoV-2 DNA
-replicas.
+extra preprocessing step is required. An enzyme called reverse
+transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
+reaction mixture into DNA. Once DNA replicas are made, the PCR process
+proceeds as described above, targeting SARS-CoV-2 DNA replicas.
 
-\section{RT-PCR errors}
-Failed detection of SARS-CoV-2 RNA in pooled RT-PCR testing is
-referred to below as a false-negative. One possible source of
-false-negatives in pooling is sample dilution. When pooling, several
-samples are mixed, so the concentration of viral RNA is reduced. This
-effect may cause a delay in amplification \cite{DorfmanYuvalDor}, no
-detection and, consequently, a false-negative. However, \cite{Lion}
-showed that this effect can be safely ignored when mixing up to 32
-samples, which we correspondingly assume.
+One possible source of failed detection of SARS-CoV-2 RNA in pooled
+RT-PCR testing (a \emph{false-negative}) is sample dilution. When
+pooling, several samples are mixed, so the concentration of viral RNA
+is reduced. This effect may cause a delay in amplification
+\cite{DorfmanYuvalDor}, no detection and, consequently, a
+false-negative. Another potential source for false-negatives is a
+failed sampling process.
+
+%% However, \cite{Lion} showed that this effect can be
+%% safely ignored when mixing up to 32 samples, which we correspondingly
+%% assume.
 
@@
 test's sensitivity (true-negative rate) is denoted $\sens$, and the
-test's specificity (true-negative rate) is denoted $\spec$. Test results
-(data) are denoted $\data$, where $\data=0$ iff the test returned a
-positive result.
+test's specificity (true-positive rate) is denoted $\spec$. Pooled
+test result (data) is denoted $\data$, where $\data=0$ iff the test
+returned a negative result.
 
-\subsection{Common assumptions}
+\subsection{The common assumption}\label{subsec:common}
 Previous studies of group testing strategies assumed that the
@@
 \begin{align}\label{eq:common}
-\Pr(\data=0|\theta) = \left \{
-\begin{array}{lr}
+  \Pr(\data=0|\theta) = 
+  \begin{cases} 
     1-\sens & \exists i\ \text{such that } \theta_i = 1\\
     \spec & \text{otherwise}
-\end{array} 
+  \end{cases} 
 \end{align}
-We refute the assumptions underlying \eqref{eq:common} in
-Section~\ref{subsec:ours} and suggest more plausible ones in
-Section~\ref{subsec:ours}.
 
-\subsection{Refuting common assumptions}\label{subsec:refute}
-We refute the common assumptions of Section~\ref{subsec:common} with
-experimental data summarized in Table \ref{table}, collected from
-\cite{Salazar}. There, the authors investigate Dorfman pooling and,
-regardless of the pooled test result, follow up and test each pool
-member separately. Here, we focus on 128 pools for which at least one
-subsequent separate test was positive --- of which 29 pooled tests
-were negative and 99 positive. In the data cited in \cite{Salazar}, of
-the 29 negative pools, subsequent separate testing yielded a single
-positive result in 24. In contrast, of the 99 positive pools, 42
-yielded a single positive test upon subsequent separate testing.
+Below, we refer to \eqref{eq:common} as the \emph{common
+  assumption}. The essence of the common assumption is that
+amplification takes place for all samples at once.
+%% We refute the assumptions underlying \eqref{eq:common} in
+%% Section~\ref{subsec:ours} and suggest more plausible ones in
+%% Section~\ref{subsec:ours}.
+
+\subsection{Refuting common assumption}\label{subsec:refute}
+We refute the common assumption with experimental data summarized in
+Table \ref{table}, collected from \cite{Salazar}. There, the authors
+investigate Dorfman pooling and, regardless of the pooled test result,
+follow up and test each pool member separately. We focus on 128 pools
+for which at least one subsequent separate test was positive --- of
+which 29 pooled tests were negative and 99 positive. In the data cited
+in \cite{Salazar}, of the 29 negative pools, subsequent separate
+testing yielded a single positive result in 24. In contrast, of the 99
+positive pools, 42 yielded a single positive test upon subsequent
+separate testing.
 
@@
 for pools with one subsequent positively tested member and pools with
-two or more such members. We apply Fisher's exact test for the
-presence of more than one positive individual in correctly identified
-pools. Fisher's test yields an increased odds ratio of 6.4, 95\% CI
-(2.2,23.4), with a p-value $\approx 10^{-4}$. Thus we reject $H_0$,
-refuting the independence of a pooled result from one or more
-individuals infected within the pool, as assumed in \cite{Simplistic1,
-  Simplistic2, OptimalDorfmanPool, Kim}
+two or more such members. $H_0$ is a direct consequence of the
+assumptions, and rejecting $H_0$ implies the common assumption are not
+realistic, at least for SARS-CoV-2.
+
+We apply Fisher's exact test for the presence of more than one
+positive individual in correctly identified pools. Fisher's test
+yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a
+p-value $\approx 10^{-4}$. Thus, we reject $H_0$, refuting the common
+assumption.
 
@@
 \subsection{Our model}
-We assume two pathways for a positive pooled test result: Viral RNA
-from an infected individual is correctly detected; or, some erroneous
-detection occurs (e.g. contaminant viral RNA is introduced). We assume
-a homogeneous and disconnected population (each individual is infected
-independently and with equal probability. Erroneous detection of
-SARS-CoV-2 RNA (a false-positive) in pooled RT-PCR testing can also
-occur. A common assumption \cite{Simplistic1, Simplistic2, Kim,
-  OptimalDorfmanPool} is that the false-positive probability does not
-depend on the number of negative samples in a pool. We incorporate
-this assumption in our likelihood model with a small modification. We
-assume that an erroneous amplification can occur in \emph{any}
-pool. Specifically, it is possible that correct amplifications fail
-and an erroneous one occurs simultaneously. This assumption is
-relatively specific for the current application of screening for
-SARS-CoV-2 via RT-PCR.  For example, cross-reactivity with other
-coronaviruses would have violated this assumption, but it was ruled
-out in \cite{KitComparison}.
+Since the essence of the failed common assumption is that
+amplification of all samples occurs only once, we assume the opposite:
+amplification of viral RNA occurs for each sample independently.
+According to \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool},
+erroneous detection (a false-positive) does not depend on the number
+of negative samples in a pool. We incorporate this assumption in our
+model with a small modification: We assume that an erroneous
+amplification can occur in \emph{any} pool. Specifically, it is
+possible that correct amplifications fail and an erroneous one occurs
+simultaneously. This assumption is somewhat specific for the current
+application of screening for SARS-CoV-2 via RT-PCR. For example,
+cross-reactivity with other coronaviruses would have violated this
+assumption, but it was ruled out in \cite{KitComparison}.
 
@@
   \end{tikzpicture}
-  \caption{Illustration of the likelihood model of a pooled test
-    result. A pool contains individuals $\{1,2,3,4\}$ with state
-    $\theta=(1,0,0,1)$ (i.e. individuals 1 and 4 are infected). A
-    negative pooled test implies that three detection paths failed. A
-    false negative occurred for $1$ and $4$, each with probability
-    $(1-\sens)$. Additionally, no erroneous detection of SARS-CoV-2
-    occurred, with probability $\spec$. Individuals $2$ and $3$ are
-    not infected and do not contribute to the probability of the
-    pooled test result. A positive pooled test arises if any one of
-    the above mentioned paths results in a
-    detection.}\label{fig:likelihood}
+  \caption{Illustration of our model. A pool contains individuals
+    $\{1,2,3,4\}$ with state $\theta=(1,0,0,1)$ (i.e. individuals 1
+    and 4 are infected). A negative pooled test ($\data=0$) implies
+    that three detection paths failed. A false-negative occurred for
+    $1$ and $4$, each with probability $(1-\sens)$ for
+    each. Additionally, no erroneous detection of SARS-CoV-2 occurred,
+    with probability $\spec$. Individuals $2$ and $3$ are not infected
+    and do not contribute to the probability of the pooled test
+    result. A positive pooled test arises if any one of the above
+    mentioned paths results in a detection.}\label{fig:likelihood}
 \end{figure}
@@
 
-\section{Application: Calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
-As an application of our model, we present the differnece in
-calculated false-negative rate for a single \emph{infected}
-individual, henceforth referred to as "Donald".  We distinguish three
-types of false-negative events when performing pooling. A \emph{single
-  test's} false-negative is the event of a negative result upon
-testing Donald separately, i.e., in an RT-PCR test without pooling. A
-\emph{pooled} false-negative occurs when a pooled test containing
-Donald's sample (and other samples) yields a negative result, i.e.,
-the pooling fails to detect at least one positive result. Lastly, a
-\emph{scheme} false-negative occurs when an entire pooling scheme
-fails to identify Donald as infected. Our goal is to calculate
-Dorfman's scheme false-negative rate. Rephrasing, we wish to answer
-the following question: what is the probability of not identifying
-Donald as infected under a Dorfman pooling scheme?
+\subsection{Application}
+As an application of our model, we calculate the differnece in
+false-negative rate for a single \emph{infected} individual,
+henceforth referred to as "Donald".  We distinguish three types of
+false-negative events when performing pooling. A \emph{single test's}
+false-negative is the event of a negative result upon testing Donald
+separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
+false-negative occurs when a pooled test containing Donald's sample
+(and other samples) yields a negative result, i.e., the pooling fails
+to detect at least one positive result. Lastly, a \emph{scheme}
+false-negative occurs when an entire pooling scheme fails to identify
+Donald as infected. Our goal is to calculate Dorfman's scheme
+false-negative rate. Rephrasing, we wish to answer the following
+question: what is the probability of not identifying Donald as
+infected under a Dorfman pooling scheme?
 
+\subsubsection{Scheme false-negative rate}
 Denote the prevalence of infection in the (tested) population $q$. We
@@
   \begin{split}
-    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta') \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
+    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta_2,\dots,\theta_N) \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
     %
@@
 We calculate the probability that Donald is mistakenly identified as
-not infected --- the scheme's false-negative rate --- denoted $\mi$
-below. To \emph{correctly} identify Donald as infected, both pooled
-and separate tests have to yield a positive result. Thus, the scheme's
-false-negative rate $\mi$ is the complement of the product of the two
-previous terms:
+not infected, henceforth referred to as the scheme's false-negative
+rate and denoted $\mi$. In order to correctly identify Donald
+as infected, both pooled and separate tests have to yield a positive
+result. Thus, the scheme's false-negative rate is:
 \begin{align}\label{eq:sfn}
@@
 
-\subsection{Comparison metric}
+\subsubsection{Comparison metric}
 The single test false-negative rate $1-\Se$ and scheme false-negative
@@
 \begin{equation}\label{eq:erel}
-\rel := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
+\rel{ours} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
 \end{equation}
-$\rel$ is the percentage increase in the pooling scheme false-negative
+$\rel{ours}$ is the percentage increase in the pooling scheme false-negative
 rate, relative to the single test false-negative rate.
 
-The approximation commonly used in the literature estimates the scheme
-false-negative rate as $1-\Se^2$ (Section \ref{subsec:assumptions} and
-\cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}). A short
-calculation shows that this approximation implies the percentage
-increase in scheme false-negative rate is just $100\cdot \Se$.
+According to the common assumption, the scheme false-negative rate is
+$1-\Se^2$. A short calculation shows that this approximation implies
+the percentage increase in scheme false-negative rate is just
+$\rel{common} := 100\cdot \Se$.
 
-\subsection{Results}\label{section:results}
-
-We plot $\rel$ for varying prevalence $q$ and sensitivity $\Se$
+\section{Results}\label{section:results}
+We plot $\rel{ours}$ for varying prevalence $q$ and sensitivity $\Se$
 values. As recommended by \cite{DorfmanYuvalDor}, we apply different
@@
   InterpretingCOVID19Test, EstimatingRatesLourenco,
-  FalsePositiveEstimate}, an increase of at least $60\%$ in $\rel$ can
-be expected (Figure \ref{figy}). Interestingly, an increase in
-infection prevalence monotonically decreases the scheme false-negative
-rate, as can also be easily seen from equation \eqref{eq:sfn}. For the
-chosen parameter ranges, the increase in the single test
-false-negative rates increases the relative error $\rel$. These
-effects can be seen in Figure \ref{figy} (left panel), upon
-conditioning on pool size. Extending the range for $\Sp$ yields no
-qualitative differences. We further compare $\rel$ to the commonly
-used approximation, showing the discrepancy changes as a function of
-both prevalence and the single test sensitivity (Figure \ref{figy},
-right panel)
+  FalsePositiveEstimate}, an increase of at least $60\%$ in
+$\rel{ours}$ can be expected (Figure \ref{figy}). Interestingly, an
+increase in infection prevalence monotonically decreases the scheme
+false-negative rate, as can also be easily seen from equation
+\eqref{eq:sfn}. For the chosen parameter ranges, the increase in the
+single test false-negative rates increases the relative error
+$\rel{ours}$. These effects can be seen in Figure \ref{figy} (left panel),
+upon conditioning on pool size. Extending the range for $\Sp$ yields
+no qualitative differences. We further compare $\rel{ours}$ to
+$\rel{common}$, showing the discrepancy changes as a function of both
+prevalence and the single test sensitivity (Figure \ref{figy}, right
+panel).
 \begin{figure}[H]
@@
     \caption{Relative increase in Dorfman pooling false-negative rates
-      $\rel$. Left: Colors represent $\rel$, the relative percentage
-      increase in the scheme false-negative rates relative to the
-      single test false-negative rates (eq \eqref{eq:erel}). Right:
-      colors represent the difference between the approximation
-      assumed in other studies and $\rel$ as calculated by our
-      model. The disease prevalence $q$, is varied on the x-axis,
-      while the test sensitivity is varied on the y-axis. Pool size
-      $N$, was chosen according to $q$ as in
+      $\rel{ours}$. Left: Colors represent $\rel{ours}$, the relative
+      percentage increase in the scheme false-negative rates relative
+      to the single test false-negative rates (eq
+      \eqref{eq:erel}). Right: colors represent the difference between
+      $\rel{common}$ and $\rel{ours}$. The disease prevalence $q$, is
+      varied on the x-axis, while the test sensitivity is varied on
+      the y-axis. Pool size $N$, was chosen according to $q$ as in
       \cite{DorfmanYuvalDor}.}\label{figy}
@@
 Although Dorfman pooling improves testing throughput, we have shown
-that it can increase $\rel$ --- the false-negative rates relative to
+that it can increase $\rel{ours}$ --- the false-negative rates relative to
 individual testing. Furthermore, low values of infection prevalence,
 or low values of single test false-negative rates, increase
-$\rel$. These results remain qualitatively similar under varying
+$\rel{ours}$. These results remain qualitatively similar under varying
 parameter values, in the observed ranges
@@
 
-An approximation to our results has been previously considered: $\mi
-\approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
-  OptimalDorfmanPool}. Such an approximation arises when one assumes
-that the probability of a false-negative is identical for the pooled
-and single tests. Although simplistic in nature, this approximation
-does capture the intuition behind our results: For Donald to be
-considered negative under Dorfman pooling, he has to test negative
-twice. In addition to this approximation only being an upper bound
-\cite{Simplistic2}, it does not account for the effect of infection
-prevalence on the false-negative rates under Dorfman
-pooling. Supporting the importance of this effect, such an association
-between prevalence and the scheme false-negative rate under Dorfman
-pooling has been empirically noted \cite{DorfmanYuvalDor}.
+%% An approximation to our results has been previously considered: $\mi
+%% \approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
+%%   OptimalDorfmanPool}. Such an approximation arises when one assumes
+%% that the probability of a false-negative is identical for the pooled
+%% and single tests. Although simplistic in nature, this approximation
+%% does capture the intuition behind our results: For Donald to be
+%% considered negative under Dorfman pooling, he has to test negative
+%% twice. In addition to this approximation only being an upper bound
+%% \cite{Simplistic2}, it does not account for the effect of infection
+%% prevalence on the false-negative rates under Dorfman
+%% pooling. Supporting the importance of this effect, such an association
+%% between prevalence and the scheme false-negative rate under Dorfman
+%% pooling has been empirically noted \cite{DorfmanYuvalDor}.
 
-Although we have shown the inherent risk of Dorfman pooling, this
-shortcoming applies to other pooling schemes. Pooling schemes
-(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
-positive pooled results to correctly identify Donald as
-infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-If the first pool yields a positive result, it is split in two. Then
-the splitting is repeated until resulting pools are negative or
-individuals are tested separately. With an initial pool size of 32,
-Donald will necessarily have to test positive in pools of size 32, 16,
-8, 4 and 2, as well as in a single test, for the scheme to correctly
-identify him as infected. Compare this to the Dorfman scheme that
-requires a positive test in a pool of size $N=8$, and an additional
-single positive test to identify Donald as infected. The hierarchical
-pooling scheme of \cite{Lion, Kim} will necessarily yield more
-false-negatives than Dorfman pooling --- there are additional places
-for it to fail.
+%% Although we have shown the inherent risk of Dorfman pooling, this
+%% shortcoming applies to other pooling schemes. Pooling schemes
+%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
+%% positive pooled results to correctly identify Donald as
+%% infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
+%% If the first pool yields a positive result, it is split in two. Then
+%% the splitting is repeated until resulting pools are negative or
+%% individuals are tested separately. With an initial pool size of 32,
+%% Donald will necessarily have to test positive in pools of size 32, 16,
+%% 8, 4 and 2, as well as in a single test, for the scheme to correctly
+%% identify him as infected. Compare this to the Dorfman scheme that
+%% requires a positive test in a pool of size $N=8$, and an additional
+%% single positive test to identify Donald as infected. The hierarchical
+%% pooling scheme of \cite{Lion, Kim} will necessarily yield more
+%% false-negatives than Dorfman pooling --- there are additional places
+%% for it to fail.
 

### 2021-03-21 comment.tex
wordz
intro

@@
 throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
-  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques are
-called \emph{pooling} and originated in the seminal work of Dorfman
-pooling \cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$
-individuals and perform a single RT-PCR test on their combined
-(\emph{pooled}) samples. If the pooled test yields a positive result
---- test each individual separately. The throughput efficiency of
+  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques, often termed
+ \emph{pooling}, or group testing,  have originated in the seminal work of Dorfman et al. and their pooling technique
+ \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$
+individuals and performs a single RT-PCR test on their combined
+(\emph{pooled}) samples. If the pooled test yields a positive result, then  each individual is retested separately; otherwise, everyone is declared negative. The throughput efficiency of
 Dorfman pooling for has been demonstrated empirically
@@
 Pooling RT-PCR tests is not perfect, and there is always some nonzero
-probability of error. Typically, authors assume the probability of a
-true-positive (the test's \emph{sensitivity}) does not depend on the
+probability of error. Typically, previous research assumed that the probability of a true-positive (the test's \emph{sensitivity}) does not depend on the
 number of infected samples in the pool, but rather on the existence of
 at least one such sample. Thus, the probability of a positive result
-in a pooled test is (assumed) identical for a pool with one sample
-from an infected individual and (e.g.) five such samples. This
+in a pooled test has been (assumed) identical for a pool with one sample
+from an infected individual and, e.g., five such samples. This
 assumption is common in the group testing literature \cite{Kim,
@@
 studies \cite{Simplistic1, Simplistic2}. In this study we refute these
-commonly made assumptions and show how they affect estimation of
+commonly made assumptions and show how using a more accurate probabilistic model affects estimation of
 false-negative rates for Dorfman pooling.
@@
     \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
-    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum\theta_i}. \label{eq:likelihood a}
+    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
     %
@@
 henceforth referred to as "Donald".  We distinguish three types of
-false-negative events when performing pooling. A \emph{single test's}
+false-negative events when performing pooling. A \emph{single test}'s
 false-negative is the event of a negative result upon testing Donald

### 2021-03-21 comment.tex
a bit of Methods

@@
 \begin{document}
-\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Tests Errors}
+\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors}
 
@@
 Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
-details and a comprehensive review of RT-PCR can be found in
+details, and a comprehensive review of RT-PCR can be found in
 \cite{COVID-RTPCR, PCRBook}. We then present the \emph{common
@@
 that the probability of a negative result is the same for a pool with
-a single sample coming from an infected individual and (e.g.)
-five. Explicitly:
+a single sample coming from an infected individual and several
+ samples from infected individuals. Explicitly:
 

### 2021-03-21 comment.tex
uri commentz

@@
   BayesianDorfman, MatrixPooling, LionDorfman}. These techniques, often termed
- \emph{pooling}, or group testing,  have originated in the seminal work of Dorfman et al. and their pooling technique
+ \emph{pooling}, or group testing, have originated in the seminal work of Dorfman and his pooling technique
  \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$
@@
 
-Pooling RT-PCR tests is not perfect, and there is always some nonzero
-probability of error. Typically, previous research assumed that the probability of a true-positive (the test's \emph{sensitivity}) does not depend on the
+Previous studies assumed that the probability of a true-positive (the test's \emph{sensitivity}) does not depend on the
 number of infected samples in the pool, but rather on the existence of
@@
   OptimalDorfmanPool}, as well as in more specific, COVID-19 focused
-studies \cite{Simplistic1, Simplistic2}. In this study we refute these
+studies \cite{Simplistic1, Simplistic2}. In this study we challenge these
 commonly made assumptions and show how using a more accurate probabilistic model affects estimation of

### 2021-03-22 comment.tex
push ot alale

@@
 
-\affil[1]{Porter School of the Environment and Earth Sciences, The Faculty of Exact Sciences, Tel Aviv
+\affil[1]{Porter School of the Environment and Earth Sciences, The
+  Faculty of Exact Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+
+\affil[2]{School of Public Health, The Faculty of Medicine, Tel Aviv
   University\\ Tel Aviv, Israel}
-\affil[2]{School of Public Health, The Faculty of Medicine, Tel Aviv University\\ Tel Aviv, Israel}
-\affil[3]{The Gertner Institute for Epidemiology and Health Policy Research, Tel Hashomer, Israel}
+
+\affil[3]{The Gertner Institute for Epidemiology and Health Policy
+  Research, Tel Hashomer, Israel}
 %\email{yair.daon@gmail.com}
@@
 \maketitle
-%%% MAKE TITLE PAGE %%%
-%% \bigskip
-%% \bigskip
-%% \bigskip
-%% \bigskip
-%% \begin{itemize}
-%%     \item Article type: Brief report.
-%%     \item Corresponding author: Uri Obolski, email uriobols@tauex.tau.ac.il, Phone +972526309508
-%%     \item Running head: Inflated false-negative rates in pooled SARS-CoV-2 tests.
-%%     \item Conflict of interests: None.
-%%     \item Financial support: Yair Daon was supported by a
-%%     post-doctoral fellowship from the Tel Aviv University Center for
-%%     Combating Pandemics and the Raymond and Beverly Sackler Dean's
-%%     Post-Doctoral fellowship.
-%%     \item Data is presented in the manuscript itself and can be found in a paper we reference to. Calculations are simple and can be easily implemented in any programming language.
-%% \end{itemize}
-%% \newpage
 
@@
 throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
-  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques, often termed
- \emph{pooling}, or group testing, have originated in the seminal work of Dorfman and his pooling technique
- \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$
-individuals and performs a single RT-PCR test on their combined
-(\emph{pooled}) samples. If the pooled test yields a positive result, then each individual is retested separately; otherwise, everyone is declared negative. The throughput efficiency of
-Dorfman pooling for has been demonstrated empirically
-\cite{DorfmanYuvalDor} and its error rates thoroguhly investigated
-\cite{Kim, Simplistic1, OptimalDorfmanPool}.
+  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques,
+often termed \emph{pooling}, or group testing, have originated in the
+seminal work of Dorfman and his pooling technique
+\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
+selects $N$ individuals and performs a single RT-PCR test on their
+combined (\emph{pooled}) samples. If the pooled test yields a positive
+result, then each individual is retested separately; otherwise,
+everyone is declared negative. The throughput efficiency of Dorfman
+pooling for has been demonstrated empirically \cite{DorfmanYuvalDor}
+and its error rates thoroguhly investigated \cite{Kim, Simplistic1,
+  OptimalDorfmanPool}.
 
-Previous studies assumed that the probability of a true-positive (the test's \emph{sensitivity}) does not depend on the
-number of infected samples in the pool, but rather on the existence of
-at least one such sample. Thus, the probability of a positive result
-in a pooled test has been (assumed) identical for a pool with one sample
-from an infected individual and, e.g., five such samples. This
-assumption is common in the group testing literature \cite{Kim,
-  OptimalDorfmanPool}, as well as in more specific, COVID-19 focused
-studies \cite{Simplistic1, Simplistic2}. In this study we challenge these
-commonly made assumptions and show how using a more accurate probabilistic model affects estimation of
-false-negative rates for Dorfman pooling.
+Previous studies assumed that the probability of a true-positive (the
+test's \emph{sensitivity}) does not depend on the number of infected
+samples in the pool, but rather on the existence of at least one such
+sample. Thus, the probability of a positive result in a pooled test
+has been (assumed) identical for a pool with one sample from an
+infected individual and, e.g., five such samples. This assumption is
+common in the group testing literature \cite{Kim, OptimalDorfmanPool},
+as well as in more specific, COVID-19 focused studies
+\cite{Simplistic1, Simplistic2}. In this study we challenge these
+commonly made assumptions and show how using a more accurate
+probabilistic model affects estimation of false-negative rates for
+Dorfman pooling.
 
-\section{Methods}
-Before delving into our probabilistic assumptions, we briefly explain
-the process of testing for SARS-CoV-2 by RT-PCR in
+\subsection{RT-PCR}\label{subsec:rtpcr}
+We briefly explain the process of testing for SARS-CoV-2 by RT-PCR in
 Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
 details, and a comprehensive review of RT-PCR can be found in
-\cite{COVID-RTPCR, PCRBook}. We then present the \emph{common
-  assumption} in Section~\ref{subsec:common} and refute it in
-Section~\ref{subsec:refute}. Our model is presented in
-Section~\ref{subsec:ours}.
+\cite{COVID-RTPCR, PCRBook}.
 
-\subsection{RT-PCR}\label{subsec:rtpcr}
 PCR is a process in which a targeted DNA strand's frequency is
@@
 
+\subsection{Pooling}
 One possible source of failed detection of SARS-CoV-2 RNA in pooled
@@
 %% assume.
+
+\section{Methods}
+We present the \emph{common assumption} in Section~\ref{subsec:common}
+and refute it in Section~\ref{subsec:refute}. Our model is presented
+in Section~\ref{subsec:ours}.
 
@@
   \begin{split}
-    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta_2,\dots,\theta_N) \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
+    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N}
+    \Pr(\theta_2,\dots,\theta_N)
+    \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
     %
@@
     %
-    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)\Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
+    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)
+    \Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
     %
@@
     %
-    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}\left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
+    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}
+    \left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
     %
@@
 We plot $\rel{ours}$ for varying prevalence $q$ and sensitivity $\Se$
-values. As recommended by \cite{DorfmanYuvalDor}, we apply different
-pool sizes $N$, for different prevalence values. We observe that for a
-false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of
-reasonable sensitivity and prevalence values \cite{KitComparison,
+values, and make the comparison with $\rel{common}$. As recommended by
+\cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
+different prevalence values. We observe that for a false-positive rate
+$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
+sensitivity and prevalence values \cite{KitComparison,
   InterpretingCOVID19Test, EstimatingRatesLourenco,
   FalsePositiveEstimate}, an increase of at least $60\%$ in
-$\rel{ours}$ can be expected (Figure \ref{figy}). Interestingly, an
-increase in infection prevalence monotonically decreases the scheme
-false-negative rate, as can also be easily seen from equation
-\eqref{eq:sfn}. For the chosen parameter ranges, the increase in the
-single test false-negative rates increases the relative error
-$\rel{ours}$. These effects can be seen in Figure \ref{figy} (left panel),
-upon conditioning on pool size. Extending the range for $\Sp$ yields
-no qualitative differences. We further compare $\rel{ours}$ to
+$\rel{ours}$ can be expected (Figure \ref{figy}).
+
+Interestingly, an increase in infection prevalence monotonically
+decreases the scheme false-negative rate, as can also be easily seen
+from \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
+the single test false-negative rates increases the relative error
+$\rel{ours}$. These effects can be seen in Figure \ref{figy} (left
+panel), upon conditioning on pool size. Extending the range for $\Sp$
+yields no qualitative differences. We further compare $\rel{ours}$ to
 $\rel{common}$, showing the discrepancy changes as a function of both
@@
 \section{Discussion}
-Although Dorfman pooling improves testing throughput, we have shown
-that it can increase $\rel{ours}$ --- the false-negative rates relative to
-individual testing. Furthermore, low values of infection prevalence,
-or low values of single test false-negative rates, increase
+We have presented a novel model for pooled tests. It is hard to
+assess the validity of our model, since this would require
+considerably more data than available in \cite{Salazar}. However, our
+model is as simple as possible, given that the standard assumptions
+are refuted. We have no doubt that further study of the assumptions
+underlying pooling is due. Furthermore, adapting common group testing
+algorithms \cite{Kim} to novel error models should be further
+studied.
+
+Our revised model has some implications for the practice of
+pooling. For example, although Dorfman pooling improves testing
+throughput, we have shown that low values of infection prevalence, or
+low values of single test false-negative rates, increase
 $\rel{ours}$. These results remain qualitatively similar under varying
@@
 when prevalence is low, such circumstances are exactly those leading
-to a substantial increase in false-negative rates.
+to a substantial increase in false-negative rates. Such behavior of
+pooling is not implied by the common assumption
 
-%% An approximation to our results has been previously considered: $\mi
-%% \approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
-%%   OptimalDorfmanPool}. Such an approximation arises when one assumes
-%% that the probability of a false-negative is identical for the pooled
-%% and single tests. Although simplistic in nature, this approximation
-%% does capture the intuition behind our results: For Donald to be
-%% considered negative under Dorfman pooling, he has to test negative
-%% twice. In addition to this approximation only being an upper bound
-%% \cite{Simplistic2}, it does not account for the effect of infection
-%% prevalence on the false-negative rates under Dorfman
-%% pooling. Supporting the importance of this effect, such an association
-%% between prevalence and the scheme false-negative rate under Dorfman
-%% pooling has been empirically noted \cite{DorfmanYuvalDor}.
+%% The common assumption implies that one assumes that the probability of
+%% a false-negative is identical for the pooled and single
+%% tests. Although simplistic in nature, the common assumption does capture
+%% the intuition behind our results: For Donald to be considered negative
+%% under Dorfman pooling, he has to test negative twice. In addition to
+%% this approximation only being an upper bound \cite{Simplistic2}, it
+%% does not account for the effect of infection prevalence on the
+%% false-negative rates under Dorfman pooling. Supporting the importance
+%% of this effect, such an association between prevalence and the scheme
+%% false-negative rate under Dorfman pooling has been empirically noted
+%% \cite{DorfmanYuvalDor}.
 
-%% Although we have shown the inherent risk of Dorfman pooling, this
-%% shortcoming applies to other pooling schemes. Pooling schemes
-%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
-%% positive pooled results to correctly identify Donald as
-%% infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-%% If the first pool yields a positive result, it is split in two. Then
-%% the splitting is repeated until resulting pools are negative or
-%% individuals are tested separately. With an initial pool size of 32,
-%% Donald will necessarily have to test positive in pools of size 32, 16,
-%% 8, 4 and 2, as well as in a single test, for the scheme to correctly
-%% identify him as infected. Compare this to the Dorfman scheme that
-%% requires a positive test in a pool of size $N=8$, and an additional
-%% single positive test to identify Donald as infected. The hierarchical
-%% pooling scheme of \cite{Lion, Kim} will necessarily yield more
-%% false-negatives than Dorfman pooling --- there are additional places
-%% for it to fail.
+Although our application focused on Dorfman pooling, our analysis
+applies to other pooling schemes. Pooling schemes
+(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
+positive pooled results to correctly identify Donald as
+infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
+If the first pool yields a positive result, it is split in two. Then
+the splitting is repeated until resulting pools are negative or
+individuals are tested separately. With an initial pool size of 32,
+Donald will necessarily have to test positive in pools of size 32, 16,
+8, 4 and 2, as well as in a single test, for the scheme to correctly
+identify him as infected. Compare this to the Dorfman scheme that
+requires a positive test in a pool of size $N=8$, and an additional
+single positive test to identify Donald as infected. The hierarchical
+pooling scheme of \cite{Lion, Kim} will necessarily yield more
+false-negatives than Dorfman pooling --- there are additional places
+for it to fail.
 

### 2021-03-22 comment.tex
Discussion

@@
 \section{Discussion}
+This is my suggestion for a Discussion structure:
+short recap; implications for group testing in different prevalence settings; limitations; ways to reduce errors (*cough* DOPE); conclusion/take home message
+
 We have presented a novel model for pooled tests. It is hard to
@@
 
-Our revised model has some implications for the practice of
-pooling. For example, although Dorfman pooling improves testing
+Our results have practical implications for the use of
+pooling schemes.
+As the COVID-19 pandemic progresses, the infection prevalence in various tested populations undergoes frequent changes.
+Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low.
+Such mass-tested populations often include air travel passengers, potentially importing infections, or asymptomatic individuals, such as children,  and can be crucial for controlling outbreaks(ref)
+
+For example, although Dorfman pooling improves testing
 throughput, we have shown that low values of infection prevalence, or
@@
 \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
-  InterpretingCOVID19Test} (Figure \ref{figy}).
+  InterpretingCOVID19Test} (Figure \ref{figy}). Hence,
 

### 2021-03-22 comment.tex
ttt

@@
 everyone is declared negative. The throughput efficiency of Dorfman
-pooling for has been demonstrated empirically \cite{DorfmanYuvalDor}
+pooling has been demonstrated empirically \cite{DorfmanYuvalDor}
 and its error rates thoroguhly investigated \cite{Kim, Simplistic1,
   OptimalDorfmanPool}.
+
+Studies focused on pooling for SARS-CoV-2 are in a consensus that sample dilution 
+effects \cite{Dilution1, Dilution2} are not a concern \cite{Lion, DorfmanYuvalDor}.
 
@@
 
-\subsection{Refuting common assumption}\label{subsec:refute}
+\subsection{Refuting the common assumption}\label{subsec:refute}
 We refute the common assumption with experimental data summarized in
@@
 probability $1-\Sp$. Our model is illustrated in
-Figure~\ref{fig:likelihood}, and summarized in
+Figure~\ref{fig:likelihood} and summarized in
 %% A negative pooled test result occurs when all detection paths (both
@@
 \subsection{Application}
-As an application of our model, we calculate the differnece in
-false-negative rate for a single \emph{infected} individual,
-henceforth referred to as "Donald". We distinguish three types of
+As an application, we calculate the false-negative rate for a single \emph{infected} individual,
+henceforth referred to as "Donald" under our model and under the common assumption. We distinguish three types of
 false-negative events when performing pooling. A \emph{single test}'s
@@
 false-negative occurs when an entire pooling scheme fails to identify
-Donald as infected. Our goal is to calculate Dorfman's scheme
+Donald as infected. Our first task is to calculate Dorfman's scheme
 false-negative rate. Rephrasing, we wish to answer the following
@@
 Denote the prevalence of infection in the (tested) population $q$. We
-give Donald the respect he deserves and assign him as individual $1$,
-so that $\theta_1=1$. Then:
+assign Donald as individual $1$, so that $\theta_1=1$. Then:
 
@@
 Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low.
-Such mass-tested populations often include air travel passengers, potentially importing infections, or asymptomatic individuals, such as children,  and can be crucial for controlling outbreaks(ref)
+Such mass-tested populations often include air travel passengers \cite{JTM}, potentially importing infections, or asymptomatic individuals, such as children, and can be crucial for controlling outbreaks \cite{MinaScience}.
 

### 2021-03-22 comment.tex
B4 merge
new fig

@@
 everyone is declared negative. The throughput efficiency of Dorfman
-pooling has been demonstrated empirically \cite{DorfmanYuvalDor}
-and its error rates thoroguhly investigated \cite{Kim, Simplistic1,
+pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
+its error rates thoroguhly investigated \cite{Kim, Simplistic1,
   OptimalDorfmanPool}.
 
-Studies focused on pooling for SARS-CoV-2 are in a consensus that sample dilution 
-effects \cite{Dilution1, Dilution2} are not a concern \cite{Lion, DorfmanYuvalDor}.
-
-Previous studies assumed that the probability of a true-positive (the
-test's \emph{sensitivity}) does not depend on the number of infected
-samples in the pool, but rather on the existence of at least one such
-sample. Thus, the probability of a positive result in a pooled test
-has been (assumed) identical for a pool with one sample from an
-infected individual and, e.g., five such samples. This assumption is
-common in the group testing literature \cite{Kim, OptimalDorfmanPool},
-as well as in more specific, COVID-19 focused studies
-\cite{Simplistic1, Simplistic2}. In this study we challenge these
-commonly made assumptions and show how using a more accurate
-probabilistic model affects estimation of false-negative rates for
-Dorfman pooling.
+Studies focused on pooling for SARS-CoV-2 are in a consensus that
+sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
+concern for pools as large as 64 individuals \cite{Lion,
+  DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed that
+the probability of a true-positive (the test's \emph{sensitivity})
+does not depend on the number of infected samples in the pool, but
+rather on the existence of at least one such sample. Thus, the
+probability of a positive result in a pooled test has been (assumed)
+identical for a pool with one sample from an infected individual and,
+e.g., five such samples. This assumption is common in the group
+testing literature \cite{Kim, OptimalDorfmanPool}, as well as in more
+specific, COVID-19 focused studies \cite{Simplistic1, Simplistic2}. In
+this study we challenge these commonly made assumptions and show how
+using a more accurate probabilistic model affects estimation of
+false-negative rates for Dorfman pooling.
 
@@
 
-\subsection{Application}
-As an application, we calculate the false-negative rate for a single \emph{infected} individual,
-henceforth referred to as "Donald" under our model and under the common assumption. We distinguish three types of
-false-negative events when performing pooling. A \emph{single test}'s
-false-negative is the event of a negative result upon testing Donald
-separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
-false-negative occurs when a pooled test containing Donald's sample
-(and other samples) yields a negative result, i.e., the pooling fails
-to detect at least one positive result. Lastly, a \emph{scheme}
-false-negative occurs when an entire pooling scheme fails to identify
-Donald as infected. Our first task is to calculate Dorfman's scheme
+\subsection{Scheme efficiency}
+The efficiency of Dorfman pooling scheme is the expected number of
+tests utilized. Let $T$ the number of tests. Then:
+\begin{align}
+  \begin{split}
+    \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
+    %
+    %
+    %
+    &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
+    %
+    %
+    %
+    &= 1 + N\sum_{k=0}^N \Pr(\sum \theta_i = k) \Pr(\data=1|\sum \theta_i = k) \\
+    %
+    %
+    %
+    &= 1 + N\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Sp(1-\Se)^k) \\
+    %
+    %
+    %
+    &= 1+N - N\Sp\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Se)^k \\
+    %
+    %
+    %
+    %&= 1+N - N\Sp \sum_{k=0}^N \binom{N}{k}(q-q\Se)^k(1-q)^k \\
+    %
+    %
+    %
+    &= 1+N - N\Sp (1-q\Se)^N\\    
+  \end{split}
+\end{align}
+
+Compare to the common assumption:
+
+\begin{align}
+  \begin{split}
+    \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
+    %
+    %
+    %
+    &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
+    %
+    %
+    %
+    &= 1 + N\Big[\Pr(\theta_i = 0,\ \forall i)(1-\Sp) + \Pr(\exists i\ \theta_i =1)\Se\Big] \\
+    %
+    %
+    %
+    &= 1 + N\Big [(1-q)^N(1-\Sp) + (1-(1-q)^N)\Se \Big ] \\
+    %
+    %
+    %
+    &= 1 + N\Big [(1-q)^N(1-\Sp - \Se) + \Se \Big ] \\
+    %
+    %
+    %
+    &= 1 + \Se N + N(1-q)^N(1-\Sp - \Se).
+  \end{split}
+\end{align}
+
+The difference per person is:
+\begin{equation}
+  1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se)
+\end{equation}
+
+\subsection{Scheme false-negative rate}
+We calculate the false-negative rate of Dorfman's pooling scheme.  The
+false-negative rate for a single \emph{infected} individual,
+henceforth referred to as "Donald" under our model and under the
+common assumption. We distinguish three types of false-negative events
+when performing pooling. A \emph{single test}'s false-negative is the
+event of a negative result upon testing Donald separately, i.e., in an
+RT-PCR test without pooling. A \emph{pooled} false-negative occurs
+when a pooled test containing Donald's sample (and other samples)
+yields a negative result, i.e., the pooling fails to detect at least
+one positive result. Lastly, a \emph{scheme} false-negative occurs
+when an entire pooling scheme fails to identify Donald as
+infected. Our first task is to calculate Dorfman's scheme
 false-negative rate. Rephrasing, we wish to answer the following
@@
 
-\subsubsection{Scheme false-negative rate}
 Denote the prevalence of infection in the (tested) population $q$. We
@@
 \begin{equation}\label{eq:erel}
-\rel{ours} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
+\err{exact} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
 \end{equation}
-$\rel{ours}$ is the percentage increase in the pooling scheme false-negative
+$\err{exact}$ is the percentage increase in the pooling scheme false-negative
 rate, relative to the single test false-negative rate.
@@
 the percentage increase in scheme false-negative rate is just
-$\rel{common} := 100\cdot \Se$.
+$\err{common} := 100\cdot \Se$.
 
 \section{Results}\label{section:results}
-We plot $\rel{ours}$ for varying prevalence $q$ and sensitivity $\Se$
-values, and make the comparison with $\rel{common}$. As recommended by
+We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
+values, and make the comparison with $\err{common}$. As recommended by
 \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
@@
   FalsePositiveEstimate}, an increase of at least $60\%$ in
-$\rel{ours}$ can be expected (Figure \ref{figy}).
+$\err{exact}$ can be expected (Figure \ref{figy}).
 
@@
 the single test false-negative rates increases the relative error
-$\rel{ours}$. These effects can be seen in Figure \ref{figy} (left
+$\err{exact}$. These effects can be seen in Figure \ref{figy} (left
 panel), upon conditioning on pool size. Extending the range for $\Sp$
-yields no qualitative differences. We further compare $\rel{ours}$ to
-$\rel{common}$, showing the discrepancy changes as a function of both
+yields no qualitative differences. We further compare $\err{exact}$ to
+$\err{common}$, showing the discrepancy changes as a function of both
 prevalence and the single test sensitivity (Figure \ref{figy}, right
@@
   \centering
-  \includegraphics[width=18cm]{heatmap.jpg}
+  \includegraphics[width=18cm]{heatmap_sfn.jpg}
   \caption{Relative increase in Dorfman pooling false-negative rates
-    $\rel{ours}$. Left: Colors represent $\rel{ours}$, the relative
+    $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
     percentage increase in the scheme false-negative rates relative
@@
     \eqref{eq:erel}). Right: colors represent the difference between
-    $\rel{common}$ and $\rel{ours}$. The disease prevalence $q$, is
+    $\err{common}$ and $\err{exact}$. The disease prevalence $q$, is
+    varied on the x-axis, while the test sensitivity is varied on
+    the y-axis. Pool size $N$, was chosen according to $q$ as in
+    \cite{DorfmanYuvalDor}.}\label{figy}
+\end{figure}
+
+The efficiency is words...
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=18cm]{heatmap_efficiency.jpg}
+  \caption{Relative increase in Dorfman pooling efficiency
+    $\eff{exact}$. Left: Colors represent $\eff{exact}$, the relative
+    percentage increase in the scheme false-negative rates relative
+    to the single test false-negative rates (eq
+    \eqref{eq:erel}). Right: colors represent the difference between
+    $\eff{common}$ and $\eff{exact}$. The disease prevalence $q$, is
     varied on the x-axis, while the test sensitivity is varied on
@@
 low values of single test false-negative rates, increase
-$\rel{ours}$. These results remain qualitatively similar under varying
+$\err{exact}$. These results remain qualitatively similar under varying
 parameter values, in the observed ranges
@@
 \end{document}
+
+%% Todos
+%% add refs to dilution and paragraph in start

### 2021-03-23 comment.tex
sga

@@
 \section{Methods}
-We present the \emph{common assumption} in Section~\ref{subsec:common}
+We define and present the \emph{common assumption} in Section~\ref{subsec:common}
 and refute it in Section~\ref{subsec:refute}. Our model is presented
@@
 \section{Discussion}
-This is my suggestion for a Discussion structure:
-short recap; implications for group testing in different prevalence settings; limitations; ways to reduce errors (*cough* DOPE); conclusion/take home message
 
-We have presented a novel model for pooled tests. It is hard to
-assess the validity of our model, since this would require
-considerably more data than available in \cite{Salazar}. However, our
-model is as simple as possible, given that the standard assumptions
-are refuted. We have no doubt that further study of the assumptions
-underlying pooling is due. Furthermore, adapting common group testing
-algorithms \cite{Kim} to novel error models should be further
-studied.
+\subsection{Short recap}
+We have presented a novel model for SARS-CoV-2 pooled tests. To the best of our knowledge, ours is the
+only one that is actually based on experimental data.
 
-Our results have practical implications for the use of
-pooling schemes.
-As the COVID-19 pandemic progresses, the infection prevalence in various tested populations undergoes frequent changes.
-Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low.
-Such mass-tested populations often include air travel passengers \cite{JTM}, potentially importing infections, or asymptomatic individuals, such as children, and can be crucial for controlling outbreaks \cite{MinaScience}.
-
-For example, although Dorfman pooling improves testing
-throughput, we have shown that low values of infection prevalence, or
+we have shown that low values of infection prevalence, or
 low values of single test false-negative rates, increase
@@
 \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
-  InterpretingCOVID19Test} (Figure \ref{figy}). Hence,
+  InterpretingCOVID19Test} (Figure \ref{figy}).
 
@@
 pooling is not implied by the common assumption
+
+\subsection{Imlications for group testing in different prevalence settings}
+Our results have practical implications for the use of
+pooling schemes.
+As the COVID-19 pandemic progresses, the infection prevalence in various tested populations undergoes frequent changes.
+Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low.
+Such mass-tested populations often include air travel passengers \cite{JTM,RobinHood}, potentially importing infections, or asymptomatic individuals, such as children, and can be crucial for controlling outbreaks \cite{MinaScience}.
+
+\subsection{Limitations}
+Lacking large amounts of experimental data, it is hard to
+assess the validity of our model. The data in \cite{Salazar} does allow us to reject the 
+common assumption but not much more. However, our model is as simple as possible, given that the standard assumptions are refuted.
+
+\subsection{Ways to reduce errors}
+
+\subsection{Conclusion}
+We have no doubt that further study of the assumptions
+underlying pooling is due. Furthermore, adapting common group testing
+algorithms \cite{Kim} to novel error models should be further
+studied.
+
+For example, although Dorfman pooling improves testing
+throughput,
 

### 2021-03-24 comment.tex
worsd

@@
 \section{Results}\label{section:results}
+\subsection{Scheme false-negative}
 We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$

### 2021-03-24 comment.tex
damn this paper

@@
 amplification of all samples occurs only once, we assume the opposite:
-amplification of viral RNA occurs for each sample independently.
-According to \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool},
-erroneous detection (a false-positive) does not depend on the number
-of negative samples in a pool. We incorporate this assumption in our
-model with a small modification: We assume that an erroneous
-amplification can occur in \emph{any} pool. Specifically, it is
-possible that correct amplifications fail and an erroneous one occurs
-simultaneously. This assumption is somewhat specific for the current
-application of screening for SARS-CoV-2 via RT-PCR. For example,
-cross-reactivity with other coronaviruses would have violated this
-assumption, but it was ruled out in \cite{KitComparison}.
+amplification of viral RNA occurs (or fails) for each sample
+independently.  According to \cite{Simplistic1, Simplistic2, Kim,
+  OptimalDorfmanPool}, erroneous detection (a false-positive) does not
+depend on the number of negative samples in a pool. We incorporate
+this assumption in our model with a small modification: We assume that
+an erroneous amplification can occur in \emph{any} pool. Specifically,
+it is possible that correct amplifications fails and an erroneous one
+occurs simultaneously. This assumption is somewhat specific for the
+current application of screening for SARS-CoV-2 via RT-PCR. For
+example, cross-reactivity with other coronaviruses would have violated
+this assumption. However, cross-reactivity was ruled out in
+\cite{KitComparison}.
 
@@
 
-The difference per person is:
+The difference is:
 \begin{equation}
-  1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se)
+  \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
 \end{equation}
@@
 \subsection{Scheme false-negative rate}
-We calculate the false-negative rate of Dorfman's pooling scheme. The
-false-negative rate for a single \emph{infected} individual,
-henceforth referred to as "Donald" under our model and under the
-common assumption. We distinguish three types of false-negative events
-when performing pooling. A \emph{single test}'s false-negative is the
-event of a negative result upon testing Donald separately, i.e., in an
-RT-PCR test without pooling. A \emph{pooled} false-negative occurs
-when a pooled test containing Donald's sample (and other samples)
-yields a negative result, i.e., the pooling fails to detect at least
-one positive result. Lastly, a \emph{scheme} false-negative occurs
-when an entire pooling scheme fails to identify Donald as
-infected. Our first task is to calculate Dorfman's scheme
+We calculate the false-negative rate for a single \emph{infected}
+individual, henceforth referred to as "Donald" under our model and
+under the common assumption. We distinguish three types of
+false-negative events when performing pooling. A \emph{single test}'s
+false-negative is the event of a negative result upon testing Donald
+separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
+false-negative occurs when a pooled test containing Donald's sample
+(and other samples) yields a negative result, i.e., the pooling fails
+to detect at least one positive result. Lastly, a \emph{scheme}
+false-negative occurs when an entire pooling scheme fails to identify
+Donald as infected. Our first task is to calculate Dorfman's scheme
 false-negative rate. Rephrasing, we wish to answer the following
@@
 
-The efficiency is words...
+\subsection{Scheme efficiency}
+Our results suggest that the efficiency under the common assumption is
+always greater than the efficiency under our exact model. This should
+not come as a surprise, as the common assumption implies larger
+false-negative rates and, consequently, a lower number of tests being
+utilized. As can be expected, the discrepancy is largest for lower
+test sensitivity rates $\Se$.
 
@@
 
-\subsection{Imlications for group testing in different prevalence settings}
-Our results have practical implications for the use of
-pooling schemes.
-As the COVID-19 pandemic progresses, the infection prevalence in various tested populations undergoes frequent changes.
-Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low.
-Such mass-tested populations often include air travel passengers \cite{JTM,RobinHood}, potentially importing infections, or asymptomatic individuals, such as children, and can be crucial for controlling outbreaks \cite{MinaScience}.
+\subsection{Implications for group testing in different prevalence settings}
+Our results have practical implications for the use of pooling
+schemes. As the COVID-19 pandemic progresses, the infection prevalence
+in various tested populations undergoes frequent changes. Hence, as
+our results suggest, pooling strategies for mass testing should be
+used with caution in populations where infection rates are low. Such
+mass-tested populations often include air travel passengers
+\cite{JTM,RobinHood}, potentially importing infections, or
+asymptomatic individuals, such as children, and can be crucial for
+controlling outbreaks \cite{MinaScience}.
+
+One important consequence of the efficiency charts is that
+practitioners order RT-PCR reagents, they should expect \emph{lower}
+scheme efficiency, and thus order more reagents.
 
 \subsection{Limitations}
-Lacking large amounts of experimental data, it is hard to
-assess the validity of our model. The data in \cite{Salazar} does allow us to reject the 
-common assumption but not much more. However, our model is as simple as possible, given that the standard assumptions are refuted.
+Lacking large amounts of experimental data, it is hard to assess the
+validity of our model. The data in \cite{Salazar} does allow us to
+reject the common assumption but not much more. Our model is as simple
+as possible, given that the standard assumptions are refuted and
+dilution effects are not taken into consideration.
 
 \subsection{Ways to reduce errors}
-
-\subsection{Conclusion}
-We have no doubt that further study of the assumptions
-underlying pooling is due. Furthermore, adapting common group testing
-algorithms \cite{Kim} to novel error models should be further
-studied.
-
-For example, although Dorfman pooling improves testing
-throughput,
-
-%% The common assumption implies that one assumes that the probability of
-%% a false-negative is identical for the pooled and single
-%% tests. Although simplistic in nature, the common assumption does capture
-%% the intuition behind our results: For Donald to be considered negative
-%% under Dorfman pooling, he has to test negative twice. In addition to
-%% this approximation only being an upper bound \cite{Simplistic2}, it
-%% does not account for the effect of infection prevalence on the
-%% false-negative rates under Dorfman pooling. Supporting the importance
-%% of this effect, such an association between prevalence and the scheme
-%% false-negative rate under Dorfman pooling has been empirically noted
-%% \cite{DorfmanYuvalDor}.
-
 Although our application focused on Dorfman pooling, our analysis
@@
 for it to fail.
+
+Thus, finding better pooling schemes is imperative. One scheme that
+holds great promise is DOPE \cite{DOPE}. This scheme utilizes
+considerably less RT-PCR tests and has lower error rates than many
+competing schemes.
+
+\subsection{Conclusion}
+We have no doubt that further study of the assumptions underlying
+pooling is due. Furthermore, adapting common group testing algorithms
+\cite{Kim} to novel error models should be further studied.
+
+For example, although Dorfman pooling improves testing throughput,
+
+%% The common assumption implies that one assumes that the probability of
+%% a false-negative is identical for the pooled and single
+%% tests. Although simplistic in nature, the common assumption does capture
+%% the intuition behind our results: For Donald to be considered negative
+%% under Dorfman pooling, he has to test negative twice. In addition to
+%% this approximation only being an upper bound \cite{Simplistic2}, it
+%% does not account for the effect of infection prevalence on the
+%% false-negative rates under Dorfman pooling. Supporting the importance
+%% of this effect, such an association between prevalence and the scheme
+%% false-negative rate under Dorfman pooling has been empirically noted
+%% \cite{DorfmanYuvalDor}.
 

### 2021-03-24 comment.tex
words and pix

@@
 sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
-concern for pools as large as 64 individuals \cite{Lion,
+concern for pools as large as 64 individuals \cite{PoolSize30, Lion,
   DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed that
@@
 
-\subsection{Our model}
+\subsection{Our model}\label{subsec:ours}
 Since the essence of the failed common assumption is that
@@
   FalsePositiveEstimate}, an increase of at least $60\%$ in
-$\err{exact}$ can be expected (Figure \ref{figy}).
+$\err{exact}$ can be expected (Figure \ref{fig1}).
 
@@
 the single test false-negative rates increases the relative error
-$\err{exact}$. These effects can be seen in Figure \ref{figy} (left
+$\err{exact}$. These effects can be seen in Figure \ref{fig1} (left
 panel), upon conditioning on pool size. Extending the range for $\Sp$
@@
 $\err{common}$, showing the discrepancy changes as a function of both
-prevalence and the single test sensitivity (Figure \ref{figy}, right
+prevalence and the single test sensitivity (Figure \ref{fig1}, right
 panel).
@@
     the y-axis. Pool size $N$, was chosen according to $q$ as in
-    \cite{DorfmanYuvalDor}.}\label{figy}
+    \cite{DorfmanYuvalDor}.}\label{fig1}
 \end{figure}
@@
     the y-axis. Pool size $N$, was chosen according to $q$ as in
-    \cite{DorfmanYuvalDor}.}\label{figy}
+    \cite{DorfmanYuvalDor}.}\label{fig2}
 \end{figure}
@@
 \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
-  InterpretingCOVID19Test} (Figure \ref{figy}).
+  InterpretingCOVID19Test} (Figure \ref{fig1}).
 
@@
 
-For example, although Dorfman pooling improves testing throughput,
+%% For example, although Dorfman pooling improves testing throughput,
 
@@
 To conclude, pooling is an important technique which can improve
-testing throughput in a cost-effective manner. Nevertheless, a
-substantial increase in pooling schemes' false-negative rates can be
-expected. Such an increase has crucial implications for controlling
-the spread of COVID-19.
+testing throughput in a cost-effective manner. Nevertheless, care must
+be given to pooling schemes' false-negative rates and
+efficiency. These are analyzed via a common assumption that we refute.
+The implications of refuting the common assumption were presented and
+should be considered by practitioners.
 
-\bibliographystyle{amsplain}
-%\bibliographystyle{ama}
+%\bibliographystyle{amsplain}
+\bibliographystyle{ama}
 \bibliography{refs}

### 2021-03-29 comment.tex
removed efficiency

@@
 \end{abstract}
+\newpage
 
-\section{Introduction}
+\section*{Introduction}
 Reverse transcription polymerase chain reaction (RT-PCR) testing is a
@@
 
-\subsection{RT-PCR}\label{subsec:rtpcr}
-We briefly explain the process of testing for SARS-CoV-2 by RT-PCR in
-Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
-details, and a comprehensive review of RT-PCR can be found in
-\cite{COVID-RTPCR, PCRBook}.
+% \subsection{RT-PCR}\label{subsec:rtpcr}
+% We briefly explain the process of testing for SARS-CoV-2 by RT-PCR in
+% Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
+% details, and a comprehensive review of RT-PCR can be found in
+% \cite{COVID-RTPCR, PCRBook}.
 
-PCR is a process in which a targeted DNA strand's frequency is
-amplified to create billions of copies in a reaction mixture. At the
-end of this \emph{amplification} process, the presence of targeted DNA
-molecules in a reaction mixture can be confidently
-determined. Arguably, the best way to understand PCR is as a chain
-reaction: Each targeted DNA molecule is copied into two, that are
-copied into four and so forth. This cascade gives PCR its name:
-polymerase chain reaction.
+% PCR is a process in which a targeted DNA strand's frequency is
+% amplified to create billions of copies in a reaction mixture. At the
+% end of this \emph{amplification} process, the presence of targeted DNA
+% molecules in a reaction mixture can be confidently
+% determined. Arguably, the best way to understand PCR is as a chain
+% reaction: Each targeted DNA molecule is copied into two, that are
+% copied into four and so forth. This cascade gives PCR its name:
+% polymerase chain reaction.
 
-Specifically, the PCR process is comprised of cycles, with each cycle
-doubling the abundance of the targeted DNA. In each cycle, each
-double-strand DNA molecule is broken into two separate strands. An
-enzyme called DNA polymerase generates a new double-strand DNA replica
-from each separate strand. Replication cannot start without a specific
-short DNA sequence, called a primer, that has to be introduced into
-the reaction mixture. Introducing the right primer into the reaction
-mixture ensures (almost) only the targeted DNA sequence is copied. A
-protein is added to the reaction mixture, which emits light upon
-successfully binding the targeted DNA. Once enough light is emitted
-the tested reaction mixture is declared positive and a
-\emph{detection} is said to occur. If, on the other hand, a predefined
-number of cycles pass without a detection event then the reaction
-mixture is declared negative.
+% Specifically, the PCR process is comprised of cycles, with each cycle
+% doubling the abundance of the targeted DNA. In each cycle, each
+% double-strand DNA molecule is broken into two separate strands. An
+% enzyme called DNA polymerase generates a new double-strand DNA replica
+% from each separate strand. Replication cannot start without a specific
+% short DNA sequence, called a primer, that has to be introduced into
+% the reaction mixture. Introducing the right primer into the reaction
+% mixture ensures (almost) only the targeted DNA sequence is copied. A
+% protein is added to the reaction mixture, which emits light upon
+% successfully binding the targeted DNA. Once enough light is emitted
+% the tested reaction mixture is declared positive and a
+% \emph{detection} is said to occur. If, on the other hand, a predefined
+% number of cycles pass without a detection event then the reaction
+% mixture is declared negative.
 
-Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
-extra preprocessing step is required. An enzyme called reverse
-transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
-reaction mixture into DNA. Once DNA replicas are made, the PCR process
-proceeds as described above, targeting SARS-CoV-2 DNA replicas.
+% Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
+% extra preprocessing step is required. An enzyme called reverse
+% transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
+% reaction mixture into DNA. Once DNA replicas are made, the PCR process
+% proceeds as described above, targeting SARS-CoV-2 DNA replicas.
 
-\subsection{Pooling}
-One possible source of failed detection of SARS-CoV-2 RNA in pooled
-RT-PCR testing (a \emph{false-negative}) is sample dilution. When
-pooling, several samples are mixed, so the concentration of viral RNA
-is reduced. This effect may cause a delay in amplification
-\cite{DorfmanYuvalDor}, no detection and, consequently, a
-false-negative. Another potential source for false-negatives is a
-failed sampling process.
+%\subsection{Pooling}
+% One possible source of failed detection of SARS-CoV-2 RNA in pooled
+% RT-PCR testing (a \emph{false-negative}) is sample dilution. When
+% pooling, several samples are mixed, so the concentration of viral RNA
+% is reduced. This effect may cause a delay in amplification
+% \cite{DorfmanYuvalDor}, no detection and, consequently, a
+% false-negative. Another potential source for false-negatives is a
+% failed sampling process.
 
@@
 
-\section{Methods}
-We define and present the \emph{common assumption} in Section~\ref{subsec:common}
-and refute it in Section~\ref{subsec:refute}. Our model is presented
-in Section~\ref{subsec:ours}.
+\section*{Methods}
+% We define and present the \emph{common assumption} in Section~\ref{subsec:common}
+% and refute it in Section~\ref{subsec:refute}. Our model is presented
+% in Section~\ref{subsec:ours}.
 
@@
 
-\subsection{The common assumption}\label{subsec:common}
+\subsection*{The common assumption}\label{subsec:common}
 Previous studies of group testing strategies assumed that the
@@
 Below, we refer to \eqref{eq:common} as the \emph{common
-  assumption}. The essence of the common assumption is that
-amplification takes place for all samples at once.
+  assumption}. 
 %% We refute the assumptions underlying \eqref{eq:common} in
@@
 
-\subsection{Refuting the common assumption}\label{subsec:refute}
+\subsection*{Refuting the common assumption}\label{subsec:refute}
 We refute the common assumption with experimental data summarized in
@@
 
-\subsection{Our model}\label{subsec:ours}
-Since the essence of the failed common assumption is that
+\subsection*{Our model}\label{subsec:ours}
+Since the essence of the refuted common assumption is that
 amplification of all samples occurs only once, we assume the opposite:
@@
 independently. According to \cite{Simplistic1, Simplistic2, Kim,
-  OptimalDorfmanPool}, erroneous detection (a false-positive) does not
+  OptimalDorfmanPool}, a false-positive does not
 depend on the number of negative samples in a pool. We incorporate
@@
 \cite{KitComparison}.
+% To summarize, we assume that for a single pool, a positive test result
+% is generated in one of two paths. Either, SARS-CoV-2 RNA from an
+% infected individual's sample is correctly amplified and detected, and
+% this can happen for each positive sample in a pool, with probability
+% $\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
+% RNA is introduced), an event that occurs at most once per pool, with
+% probability $1-\Sp$.
+These assumptions lead to our model, which is illustrated in 
+Figure~\ref{fig:likelihood} and summarized in
+%% A negative pooled test result occurs when all detection paths (both
+%% correct and erroneous) fail. The probability of no false-detection
+%% accounts for the $\spec$ term. The probability of no correct detection
+%% is $(1-\sens)$ per infected individual. The probability that all such
+%% paths fail is the product of the above mentioned terms, displayed in
+\eqref{eq:likelihood a}.%%  The probability of a positive result,
+%% presented in \eqref{eq:likelihood b}, is simply the complement.
+%% The probability of a negative pooled test result is presented in
+%% \eqref{eq:likelihood a} (along with its complement
+%% \eqref{eq:likelihood b}), and explained below.
+%\begin{subequations}
+\begin{align}%\label{eq:single likelihood}
+  %\begin{split}
+    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
+    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
+    %
+    %
+    %
+    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
+    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
+  %\end{split}
+\end{align}
+%\end{subequations}
+%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
+%% recalling that $\data \in \{0,1\}$ yields:
+
+%% \begin{align}\label{eq:combined likelihood}
+%%   \begin{split}
+%%     \Pr(\data| \theta) &= \left [\spec\prod_{i=1}^N
+%%       (1-\sens)^{\theta_i} \right ]^{1-\data} \left [1 - \spec\prod_{i=1}^N (1-\sens)^{\theta_i} \right ]^{\data}.
+%%   \end{split}
+%% \end{align}
 
@@
 
-To summarize, we assume that for a single pool, a positive test result
-is generated in one of two paths. Either, SARS-CoV-2 RNA from an
-infected individual's sample is correctly amplified and detected, and
-this can happen for each positive sample in a pool, with probability
-$\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
-RNA is introduced), an event that occurs at most once per pool, with
-probability $1-\Sp$. Our model is illustrated in
-Figure~\ref{fig:likelihood} and summarized in
-%% A negative pooled test result occurs when all detection paths (both
-%% correct and erroneous) fail. The probability of no false-detection
-%% accounts for the $\spec$ term. The probability of no correct detection
-%% is $(1-\sens)$ per infected individual. The probability that all such
-%% paths fail is the product of the above mentioned terms, displayed in
-\eqref{eq:likelihood a}.%%  The probability of a positive result,
-%% presented in \eqref{eq:likelihood b}, is simply the complement.
-%% The probability of a negative pooled test result is presented in
-%% \eqref{eq:likelihood a} (along with its complement
-%% \eqref{eq:likelihood b}), and explained below.
-%\begin{subequations}
-\begin{align}%\label{eq:single likelihood}
-  %\begin{split}
-    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
-    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
-    %
-    %
-    %
-    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
-    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
-  %\end{split}
-\end{align}
-%\end{subequations}
-%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
-%% recalling that $\data \in \{0,1\}$ yields:
-
-%% \begin{align}\label{eq:combined likelihood}
+%% \subsection{Scheme efficiency}
+%% The efficiency of Dorfman pooling scheme is the expected number of
+%% tests utilized. Let $T$ the number of tests. Then:
+%% \begin{align}
 %%   \begin{split}
-%%     \Pr(\data| \theta) &= \left [\spec\prod_{i=1}^N
-%%       (1-\sens)^{\theta_i} \right ]^{1-\data} \left [1 - \spec\prod_{i=1}^N (1-\sens)^{\theta_i} \right ]^{\data}.
+%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\sum_{k=0}^N \Pr(\sum \theta_i = k) \Pr(\data=1|\sum \theta_i = k) \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Sp(1-\Se)^k) \\
+%%     %
+%%     %
+%%     %
+%%     &= 1+N - N\Sp\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Se)^k \\
+%%     %
+%%     %
+%%     %
+%%     %&= 1+N - N\Sp \sum_{k=0}^N \binom{N}{k}(q-q\Se)^k(1-q)^k \\
+%%     %
+%%     %
+%%     %
+%%     &= 1+N - N\Sp (1-q\Se)^N\\    
 %%   \end{split}
@@
 
-\subsection{Scheme efficiency}
-The efficiency of Dorfman pooling scheme is the expected number of
-tests utilized. Let $T$ the number of tests. Then:
-\begin{align}
-  \begin{split}
-    \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
-    %
-    %
-    %
-    &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
-    %
-    %
-    %
-    &= 1 + N\sum_{k=0}^N \Pr(\sum \theta_i = k) \Pr(\data=1|\sum \theta_i = k) \\
-    %
-    %
-    %
-    &= 1 + N\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Sp(1-\Se)^k) \\
-    %
-    %
-    %
-    &= 1+N - N\Sp\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Se)^k \\
-    %
-    %
-    %
-    %&= 1+N - N\Sp \sum_{k=0}^N \binom{N}{k}(q-q\Se)^k(1-q)^k \\
-    %
-    %
-    %
-    &= 1+N - N\Sp (1-q\Se)^N\\    
-  \end{split}
-\end{align}
+%% Compare to the common assumption:
 
-Compare to the common assumption:
+%% \begin{align}
+%%   \begin{split}
+%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\Big[\Pr(\theta_i = 0,\ \forall i)(1-\Sp) + \Pr(\exists i\ \theta_i =1)\Se\Big] \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\Big [(1-q)^N(1-\Sp) + (1-(1-q)^N)\Se \Big ] \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + N\Big [(1-q)^N(1-\Sp - \Se) + \Se \Big ] \\
+%%     %
+%%     %
+%%     %
+%%     &= 1 + \Se N + N(1-q)^N(1-\Sp - \Se).
+%%   \end{split}
+%% \end{align}
 
-\begin{align}
-  \begin{split}
-    \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
-    %
-    %
-    %
-    &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
-    %
-    %
-    %
-    &= 1 + N\Big[\Pr(\theta_i = 0,\ \forall i)(1-\Sp) + \Pr(\exists i\ \theta_i =1)\Se\Big] \\
-    %
-    %
-    %
-    &= 1 + N\Big [(1-q)^N(1-\Sp) + (1-(1-q)^N)\Se \Big ] \\
-    %
-    %
-    %
-    &= 1 + N\Big [(1-q)^N(1-\Sp - \Se) + \Se \Big ] \\
-    %
-    %
-    %
-    &= 1 + \Se N + N(1-q)^N(1-\Sp - \Se).
-  \end{split}
-\end{align}
+%% The difference is:
+%% \begin{equation}
+%%   \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
+%% \end{equation}
 
-The difference is:
-\begin{equation}
-  \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
-\end{equation}
-
-\subsection{Scheme false-negative rate}
+\subsection*{Scheme false-negative rate}
 We calculate the false-negative rate for a single \emph{infected}
@@
 Donald as infected. Our first task is to calculate Dorfman's scheme
-false-negative rate. Rephrasing, we wish to answer the following
-question: what is the probability of not identifying Donald as
+false-negative rate. Equivalently, what is the probability of not identifying Donald as
 infected under a Dorfman pooling scheme?
@@
     %
-    &= \Sp(1-\Se)(1-q\Se)^{N-1}, \text{ by Newton's Binomial Theorem.} \\
+    &= \Sp(1-\Se)(1-q\Se)^{N-1}. \\
   \end{split}
@@
 
-\subsubsection{Comparison metric}
+\subsubsection*{Comparison metric}
 The single test false-negative rate $1-\Se$ and scheme false-negative
@@
 
-\section{Results}\label{section:results}
-\subsection{Scheme false-negative}
+\section*{Results}\label{section:results}
+\subsection*{Scheme false-negative}
 We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
@@
   \centering
-  \includegraphics[width=18cm]{heatmap_sfn.jpg}
+  \includegraphics[width=\textwidth]{heatmap_sfn.jpg}
   \caption{Relative increase in Dorfman pooling false-negative rates
@@
 
-\subsection{Scheme efficiency}
-Our results suggest that the efficiency under the common assumption is
-always greater than the efficiency under our exact model. This should
-not come as a surprise, as the common assumption implies larger
-false-negative rates and, consequently, a lower number of tests being
-utilized. As can be expected, the discrepancy is largest for lower
-test sensitivity rates $\Se$.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=18cm]{heatmap_efficiency.jpg}
-  \caption{Relative increase in Dorfman pooling efficiency
-    $\eff{exact}$. Left: Colors represent $\eff{exact}$, the relative
-    percentage increase in the scheme false-negative rates relative
-    to the single test false-negative rates (eq
-    \eqref{eq:erel}). Right: colors represent the difference between
-    $\eff{common}$ and $\eff{exact}$. The disease prevalence $q$, is
-    varied on the x-axis, while the test sensitivity is varied on
-    the y-axis. Pool size $N$, was chosen according to $q$ as in
-    \cite{DorfmanYuvalDor}.}\label{fig2}
-\end{figure}
-
-\section{Discussion}
-
-\subsection{Short recap}
+\section*{Discussion}
+%\subsection{Short recap}
 We have presented a novel model for SARS-CoV-2 pooled tests. To the best of our knowledge, ours is the
@@
 
-\subsection{Implications for group testing in different prevalence settings}
+%\subsection{Implications for group testing in different prevalence settings}
 Our results have practical implications for the use of pooling
@@
 
-\subsection{Limitations}
+%\subsection{Limitations}
 Lacking large amounts of experimental data, it is hard to assess the
@@
 
-\subsection{Ways to reduce errors}
+%\subsection{Ways to reduce errors}
 Although our application focused on Dorfman pooling, our analysis
@@
 
-\subsection{Conclusion}
+%\subsection{Conclusion}
 We have no doubt that further study of the assumptions underlying
@@
 
-%\bibliographystyle{amsplain}
-\bibliographystyle{ama}
+\bibliographystyle{amsplain}
 \bibliography{refs}

### 2021-03-30 comment.tex
figure 1 caption, a lot of changes in the discussion

@@
       \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
-      \node[shape=circle,draw=black] (B) at (2,0) {$\theta_2=0$};
-      \node[shape=circle,draw=black] (D) at (4,0) {$\theta_3=0$};
-      \node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};
+      \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
+      \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
+      %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};
 
-      \node[shape=ellipse,draw=black]
-      (F) at (9,0) {False-positive};
+      \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};
 
-      \node[shape=rectangle,draw=black] (G) at (4,7)
+      \node[shape=rectangle,draw=black] (G) at (1.5,5)
            {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
@@
         every node/.style={fill=white},
-        every edge/.style={draw=red,very thick}]
-      \path [->] (A) edge node {$1-\sens$} (4.5,6.6);
+        every edge/.style={draw=black,very thick}]
+      \path [->] (A) edge node {$1-\sens$} (2,4.6);
       %\path [->] (B) edge node {$0$} (G);
       %\path [->] (D) edge node {$0$} (G);
-      \path [->] (E) edge node {$1-\sens$} (4.5,6.6);
-      \path [->] (F) edge node {$\spec$} (5.5,6.6); 
+      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
+        \path [->] (F) edge node {$\spec$} (3.7,5); 
     \end{scope}
@@
   \caption{Illustration of our model. A pool contains individuals
-    $\{1,2,3,4\}$ with state $\theta=(1,0,0,1)$ (i.e. individuals 1
-    and 4 are infected). A negative pooled test ($\data=0$) implies
-    that three detection paths failed. A false-negative occurred for
-    $1$ and $4$, each with probability $(1-\sens)$ for
-    each. Additionally, no erroneous detection of SARS-CoV-2 occurred,
-    with probability $\spec$. Individuals $2$ and $3$ are not infected
-    and do not contribute to the probability of the pooled test
-    result. A positive pooled test arises if any one of the above
-    mentioned paths results in a detection.}\label{fig:likelihood}
+    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1
+    and 2 are infected). A negative pooled test ($\data=0$) occurs when
+    three detection paths fail: A false-negative occurs for individuals
+    $1$ and $2$, each with probability $(1-\sens)$.
+    Additionally, no false-positive detection of SARS-CoV-2 occurs,
+    with probability $\spec$. Individual $3$ is not infected
+    and does not contribute to the probability of the pooled test
+    result. The pooled test is positive if any path results in a detection.}\label{fig:likelihood}
 \end{figure}
@@
 %\subsection{Short recap}
-We have presented a novel model for SARS-CoV-2 pooled tests. To the best of our knowledge, ours is the
-only one that is actually based on experimental data.
-
-we have shown that low values of infection prevalence, or
-low values of single test false-negative rates, increase
+Here, we presented a novel probabilistic model for the outcomes of pooled PCR tests, parametrized for SARS-CoV-2.
+Contrary to previous models, we assume, based on data (see section XX), that multiple infected individuals increase the chance of a positive pool.
+A direct consequence of our model is that  false-negative rates depend on infection prevalence. Specifically, low values of infection prevalence increase
 $\err{exact}$. These results remain qualitatively similar under varying
@@
 
-Our results lead to a fact almost disregarded in
+Our results lead to a conflict: low infection prevalence leads to high efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also increasing false-negative rates.
+
+fact almost disregarded in
 \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
-when prevalence is low, such circumstances are exactly those leading
+when infection prevalence is low, such circumstances are exactly those leading
 to a substantial increase in false-negative rates. Such behavior of
@@
 %\subsection{Implications for group testing in different prevalence settings}
-Our results have practical implications for the use of pooling
-schemes. As the COVID-19 pandemic progresses, the infection prevalence
+As the COVID-19 pandemic progresses, the infection prevalence
 in various tested populations undergoes frequent changes. Hence, as
-our results suggest, pooling strategies for mass testing should be
+our results suggest, pooling strategies employed for mass testing should be
 used with caution in populations where infection rates are low. Such
@@
 \cite{JTM,RobinHood}, potentially importing infections, or
-asymptomatic individuals, such as children, and can be crucial for
+presymptomatic and asymptomatic individuals, and can be crucial for
 controlling outbreaks \cite{MinaScience}.
@@
 infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-If the first pool yields a positive result, it is split in two. Then
+If the first pool yields a positive result, it is split in two; otherwise, everyone are declared negative. Then
 the splitting is repeated until resulting pools are negative or
@@
 
-Thus, finding better pooling schemes is imperative. One scheme that
-holds great promise is DOPE \cite{DOPE}. This scheme utilizes
-considerably less RT-PCR tests and has lower error rates than many
-competing schemes.
+Thus, finding improved pooling schemes is imperative.
+An especially important consideration in designing such schemes is explicitly taking the intrinsic PCR error rates into account.
+We (ref), as well as others (ref), have developed such models, although their implementation in real-world settings has not yet commenced.
 
@@
 
-As mentioned in \cite{DorfmanYuvalDor}, introducing a positive
+Another important factor in pooling is dependence between samples. As mentioned in \cite{DorfmanYuvalDor}, introducing a positive
 dependence within a pool decreases the false-positive rate. In the
 extreme case, consider a fully connected pool, where one infection
-implies the entire pool is infected. In this case, a calculation
-analogous to the one conducted above recovers the initial
+determines the entire pool is infected. In this case, a calculation
+analogous to the one conducted above (section??) recovers the initial
 false-negative rate $1-\Se$. Interestingly, pooling was also noted to
@@
 another advantage to sampling dependent individuals in pooling
-schemes.
+schemes. Hence, deliberately pooling individuals with positive dependence of infection should be further studied and likely encouraged when performing pooling.
 
-To conclude, pooling is an important technique which can improve
+To conclude, pooling is an important technique that can improve
 testing throughput in a cost-effective manner. Nevertheless, care must
-be given to pooling schemes' false-negative rates and
-efficiency. These are analyzed via a common assumption that we refute.
-The implications of refuting the common assumption were presented and
-should be considered by practitioners.
+be given to pooling schemes' false-negative rates, especially under low infection prevalence settings.
 

### 2021-03-30 comment.tex
blocking arrow, remove paragraphs

@@
 
-\author[1,2]{Yair Daon, PhD}
-\author[2,3]{Amit Huppert, PhD}
-\author[1,2]{Uri Obolski, PhD}
+\author[1,2]{Yair Daon}
+\author[2,3,*]{Amit Huppert}
+\author[1,2,*]{Uri Obolski}
 
@@
   Research, Tel Hashomer, Israel}
+\affil[*]{Equal contribution}
 %\email{yair.daon@gmail.com}
@@
 Formally, we consider a pool containing $N$ individuals
-$\{1,\dots,N\}$. We denote the true infection states $\theta \in
+$\{1,\dots,N\}$. We denote the true infection state $\theta \in
 \{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
@@
 test's specificity (true-positive rate) is denoted $\spec$. Pooled
-test result (data) is denoted $\data$, where $\data=0$ iff the test
-returned a negative result.
+test result (data) is denoted $\data \in \{0,1\}$, where $\data=0$ iff
+the test returned a negative result.
 
@@
 \subsection*{Refuting the common assumption}\label{subsec:refute}
-We refute the common assumption with experimental data summarized in
-Table \ref{table}, collected from \cite{Salazar}. There, the authors
-investigate Dorfman pooling and, regardless of the pooled test result,
-follow up and test each pool member separately. We focus on 128 pools
-for which at least one subsequent separate test was positive --- of
-which 29 pooled tests were negative and 99 positive. In the data cited
-in \cite{Salazar}, of the 29 negative pools, subsequent separate
-testing yielded a single positive result in 24. In contrast, of the 99
-positive pools, 42 yielded a single positive test upon subsequent
-separate testing.
+We refute the common assumption with experimental data collected from
+\cite{Salazar}, and summarized in Table \ref{table}. There, the
+authors investigate Dorfman pooling and, regardless of the pooled test
+result, follow up and test each pool member separately. We focus on
+128 pools for which at least one subsequent separate test was positive
+--- of which 29 pooled tests were negative and 99 positive. In the
+data cited in \cite{Salazar}, of the 29 negative pools, subsequent
+separate testing yielded a single positive result in 24. In contrast,
+of the 99 positive pools, 42 yielded a single positive test upon
+subsequent separate testing.
 
@@
 for pools with one subsequent positively tested member and pools with
-two or more such members. $H_0$ is a direct consequence of the
-assumptions, and rejecting $H_0$ implies the common assumption are not
+two or more such members. $H_0$ is a direct consequence of the common
+assumption, and rejecting $H_0$ implies the common assumption is not
 realistic, at least for SARS-CoV-2.
@@
 amplification of all samples occurs only once, we assume the opposite:
-amplification of viral RNA occurs (or fails) for each sample
+amplification of viral RNA succeeds or fails for each sample
 independently. According to \cite{Simplistic1, Simplistic2, Kim,
@@
       \path [->] (F) edge node {$\spec$} (3.7,5);
+      \path [-Bar] (D) edge (4.5, 2);
     \end{scope}
@@
   \caption{Illustration of our model. A pool contains individuals
-    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1
-    and 2 are infected). A negative pooled test ($\data=0$) occurs when
-    three detection paths fail: A false-negative occurs for individuals
-    $1$ and $2$, each with probability $(1-\sens)$.
+    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
+    are infected). A negative pooled test ($\data=0$) occurs when
+    three detection paths fail: A false-negative occurs for
+    individuals $1$ and $2$, each with probability $(1-\sens)$.
     Additionally, no false-positive detection of SARS-CoV-2 occurs,
-    with probability $\spec$. Individual $3$ is not infected
-    and does not contribute to the probability of the pooled test
-    result. The pooled test is positive if any path results in a detection.}\label{fig:likelihood}
+    with probability $\spec$. Individual $3$ is not infected and does
+    not contribute to the probability of the pooled test result. The
+    pooled test is positive ($\data=1$) if any path results in a
+    detection.}\label{fig:likelihood}
 \end{figure}
@@
 \end{equation}
-$\err{exact}$ is the percentage increase in the pooling scheme false-negative
-rate, relative to the single test false-negative rate.
+$\err{exact}$ is the percentage increase in the pooling scheme
+false-negative rate, relative to the single test false-negative rate.
 
@@
 $1-\Se^2$. A short calculation shows that this approximation implies
-the percentage increase in scheme false-negative rate is just
-$\err{common} := 100\cdot \Se$.
+the percentage increase in scheme false-negative rate is $\err{common}
+:= 100\cdot \Se$.
 
@@
 %\subsection{Short recap}
-Here, we presented a novel probabilistic model for the outcomes of pooled PCR tests, parametrized for SARS-CoV-2.
-Contrary to previous models, we assume, based on data (see section XX), that multiple infected individuals increase the chance of a positive pool.
-A direct consequence of our model is that  false-negative rates depend on infection prevalence. Specifically, low values of infection prevalence increase
-$\err{exact}$. These results remain qualitatively similar under varying
-parameter values, in the observed ranges
+Here, we presented a novel probabilistic model for the outcomes of
+pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
+assumption, we assume, based on data (see section XX), that multiple
+infected individuals increase the likelihood of a positive pooled test
+result. A direct consequence of our model is that false-negative rates
+depend on infection prevalence. Specifically, low values of infection
+prevalence increase $\err{exact}$. These results remain qualitatively
+similar under varying parameter values, in the observed ranges
 \cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
@@
 
-Our results lead to a conflict: low infection prevalence leads to high efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also increasing false-negative rates.
-
-fact almost disregarded in
+Our results lead to a conflict: low infection prevalence leads to high
+efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
+increasing false-negative rates. This is a fact almost disregarded in
 \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
-when infection prevalence is low, such circumstances are exactly those leading
-to a substantial increase in false-negative rates. Such behavior of
-pooling is not implied by the common assumption
+when infection prevalence is low, such circumstances are exactly those
+leading to a substantial increase in false-negative rates. Such
+behavior of pooling is not implied by the common assumption.
 
@@
 
-One important consequence of the efficiency charts is that
-practitioners order RT-PCR reagents, they should expect \emph{lower}
-scheme efficiency, and thus order more reagents.
+%% One important consequence of the efficiency charts is that
+%% practitioners order RT-PCR reagents, they should expect \emph{lower}
+%% scheme efficiency, and thus order more reagents.
 
@@
 validity of our model. The data in \cite{Salazar} does allow us to
-reject the common assumption but not much more. Our model is as simple
-as possible, given that the standard assumptions are refuted and
-dilution effects are not taken into consideration.
+reject the common assumption but not much more. Our model assumes as
+little as as possible, given that the standard assumptions are refuted
+and dilution effects are not taken into consideration.
 
@@
 infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-If the first pool yields a positive result, it is split in two; otherwise, everyone are declared negative. Then
-the splitting is repeated until resulting pools are negative or
-individuals are tested separately. With an initial pool size of 32,
-Donald will necessarily have to test positive in pools of size 32, 16,
-8, 4 and 2, as well as in a single test, for the scheme to correctly
-identify him as infected. Compare this to the Dorfman scheme that
-requires a positive test in a pool of size $N=8$, and an additional
-single positive test to identify Donald as infected. The hierarchical
-pooling scheme of \cite{Lion, Kim} will necessarily yield more
-false-negatives than Dorfman pooling --- there are additional places
-for it to fail.
+If the first pool yields a positive result, it is split in two;
+otherwise, everyone are declared negative. Then the splitting is
+repeated until resulting pools are negative or individuals are tested
+separately. With an initial pool size of 32, Donald will necessarily
+have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
+in a single test, for the scheme to correctly identify him as
+infected. Compare this to the Dorfman scheme that requires a positive
+test in a pool of size $N=8$, and an additional single positive test
+to identify Donald as infected. The hierarchical pooling scheme of
+\cite{Lion, Kim} will necessarily yield more false-negatives than
+Dorfman pooling --- there are additional places for it to fail.
 
-Thus, finding improved pooling schemes is imperative.
-An especially important consideration in designing such schemes is explicitly taking the intrinsic PCR error rates into account.
-We (ref), as well as others (ref), have developed such models, although their implementation in real-world settings has not yet commenced.
+Thus, finding improved pooling schemes is imperative. An especially
+important consideration in designing such schemes is explicitly taking
+the intrinsic PCR error rates into account. We (ref), as well as
+others (ref), have developed such models, although their
+implementation in real-world settings has not yet commenced.
 
@@
 
-Another important factor in pooling is dependence between samples. As mentioned in \cite{DorfmanYuvalDor}, introducing a positive
-dependence within a pool decreases the false-positive rate. In the
-extreme case, consider a fully connected pool, where one infection
-determines the entire pool is infected. In this case, a calculation
-analogous to the one conducted above (section??) recovers the initial
-false-negative rate $1-\Se$. Interestingly, pooling was also noted to
-have increased throughput when infection probabilities are dependent
-between the pooled individuals \cite{DorfmanYuvalDor}, providing
-another advantage to sampling dependent individuals in pooling
-schemes. Hence, deliberately pooling individuals with positive dependence of infection should be further studied and likely encouraged when performing pooling.
+%% Another important factor in pooling is dependence between samples. As
+%% mentioned in \cite{DorfmanYuvalDor}, introducing a positive dependence
+%% within a pool decreases the false-positive rate. In the extreme case,
+%% consider a fully connected pool, where one infection determines the
+%% entire pool is infected. In this case, a calculation analogous to the
+%% one conducted above (section??) recovers the initial false-negative
+%% rate $1-\Se$. Interestingly, pooling was also noted to have increased
+%% throughput when infection probabilities are dependent between the
+%% pooled individuals \cite{DorfmanYuvalDor}, providing another advantage
+%% to sampling dependent individuals in pooling schemes. Hence,
+%% deliberately pooling individuals with positive dependence of infection
+%% should be further studied and likely encouraged when performing
+%% pooling.
 
-To conclude, pooling is an important technique that can improve
-testing throughput in a cost-effective manner. Nevertheless, care must
-be given to pooling schemes' false-negative rates, especially under low infection prevalence settings.
+%% To conclude, pooling is an important technique that can improve
+%% testing throughput in a cost-effective manner. Nevertheless, care must
+%% be given to pooling schemes' false-negative rates, especially under
+%% low infection prevalence settings.
 

### 2021-03-30 comment.tex
equal contibutions

@@
 \author[1,2]{Yair Daon}
-\author[2,3,*]{Amit Huppert}
+\author[1,3,*]{Amit Huppert}
 \author[1,2,*]{Uri Obolski}
 
-\affil[1]{Porter School of the Environment and Earth Sciences, The
+\affil[1]{School of Public Health, The Faculty of Medicine, Tel Aviv
+  University\\ Tel Aviv, Israel}
+
+\affil[2]{Porter School of the Environment and Earth Sciences, The
   Faculty of Exact Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-
-\affil[2]{School of Public Health, The Faculty of Medicine, Tel Aviv
-  University\\ Tel Aviv, Israel}
 

### 2021-03-31 comment.tex
Updates from Overleaf

--- 2021-03-30 comment.tex
@@
       \path [->] (F) edge node {$\spec$} (3.7,5);
-      \path [-Bar] (D) edge (4.5, 2);
+      %\path [-Bar] (D) edge (4.5, 2);
     \end{scope}
@@
     with probability $\spec$. Individual $3$ is not infected and does
-    not contribute to the probability of the pooled test result. The
-    pooled test is positive ($\data=1$) if any path results in a
-    detection.}\label{fig:likelihood}
+    not contribute to the probability of the pooled test result.}\label{fig:likelihood}
 \end{figure}
@@
 pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
-assumption, we assume, based on data (see section XX), that multiple
+assumption, we assume, based on data \cite{Salazar}, that multiple
 infected individuals increase the likelihood of a positive pooled test
@@
   InterpretingCOVID19Test} (Figure \ref{fig1}).
-
 Our results lead to a conflict: low infection prevalence leads to high
 efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
-increasing false-negative rates. This is a fact almost disregarded in
-\cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
-when infection prevalence is low, such circumstances are exactly those
-leading to a substantial increase in false-negative rates. Such
-behavior of pooling is not implied by the common assumption.
+increasing false-negative rates.
+
+% This is a fact almost disregarded in
+% \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
+% when infection prevalence is low, such circumstances are exactly those
+% leading to a substantial increase in false-negative rates. Such
+% behavior of pooling is not implied by the common assumption.
 
@@
 
-%% One important consequence of the efficiency charts is that
-%% practitioners order RT-PCR reagents, they should expect \emph{lower}
-%% scheme efficiency, and thus order more reagents.
-
 %\subsection{Limitations}
-Lacking large amounts of experimental data, it is hard to assess the
-validity of our model. The data in \cite{Salazar} does allow us to
-reject the common assumption but not much more. Our model assumes as
-little as as possible, given that the standard assumptions are refuted
-and dilution effects are not taken into consideration.
+%NONE
 
@@
 (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
-positive pooled results to correctly identify Donald as
-infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
+positive pooled results to correctly identify an infected individual as
+such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
 If the first pool yields a positive result, it is split in two;
@@
 important consideration in designing such schemes is explicitly taking
-the intrinsic PCR error rates into account. We (ref), as well as
-others (ref), have developed such models, although their
+the intrinsic PCR error rates into account. We \cite{DOPE}, as well as
+others \cite{BayesianDorfman}, have developed such models, although their
 implementation in real-world settings has not yet commenced.
@@
 %\subsection{Conclusion}
-We have no doubt that further study of the assumptions underlying
-pooling is due. Furthermore, adapting common group testing algorithms
-\cite{Kim} to novel error models should be further studied.
 
@@
 
-%% To conclude, pooling is an important technique that can improve
-%% testing throughput in a cost-effective manner. Nevertheless, care must
-%% be given to pooling schemes' false-negative rates, especially under
-%% low infection prevalence settings.
+%% Conclusion
+To conclude, pooling is an important technique that can improve
+testing throughput in a cost-effective manner. Nevertheless, care must
+be given to pooling schemes' false-negative rates, especially under
+low infection prevalence settings.
 

### 2021-03-31 comment.tex
commitz

@@
 \begin{abstract}
-\textbf{Background}: Pooling is a popular strategy for increasing
+pooling is a popular strategy for increasing
 SARS-CoV-2 testing throughput. One popular pooling scheme is Dorfman
-pooling: test $N$ individuals simultaneously. If the test is positive
---- retest each individual separately. However, requiring more than
-one positive test may lead to increased false-negative rates.
-
-\textbf{Methods}: We analyze the false-negative rate (i.e., the
-probability of a negative result for an infected individual) of
-Dorfman pooling via a new probabilistic model. We demonstrate that
-different, previously made probabilistic assumptions regarding pooling
-are unlikely in light of empiric data. Our model is conservative in
-that it ignores sample dilution effects, which can only worsen pooling
-performance.
-
-\textbf{Results}: We show that one can expect a 60-80\% increase in
+pooling: test $N$ individuals simultaneously. If the test is positive, each individual is  then tested separately. When error rates are taken into consideration,
+most analyses assume that including extra infected samples in a pooled test has no effect 
+on the pooled test outcome. We refute this assumption with experimental data
+and suggest a novel probabilistic model for the outcomes of pooled tests. Our model is
+parsimonous and assumes as little as possible. We analyze the false-negative
+rate (i.e., the probability of a negative result for an infected individual) of
+Dorfman pooling via a our model. We show that one can expect a 60-80\% increase in
 false-negative rates under Dorfman pooling, for reasonable parameter
@@
 pooling increase when the prevalence of infection decreases.
-
-\textbf{Discussion}: In most pooling schemes, identifying an infected
+In most pooling schemes, identifying an infected
 individual requires positive results in multiple tests and hence
@@
 not infected, henceforth referred to as the scheme's false-negative
-rate and denoted $\mi$. In order to correctly identify Donald
+rate and denoted $\mi$. In order to correctly identify an infected individual
 as infected, both pooled and separate tests have to yield a positive
@@
 the single test false-negative rates increases the relative error
-$\err{exact}$. These effects can be seen in Figure \ref{fig1} (left
+$\err{exact}$. These effects can be seen in Figure~\ref{fig1} (left
 panel), upon conditioning on pool size. Extending the range for $\Sp$
@@
 $\err{common}$, showing the discrepancy changes as a function of both
-prevalence and the single test sensitivity (Figure \ref{fig1}, right
+prevalence and the single test sensitivity (Figure~\ref{fig1}, right
 panel).
@@
 pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
-assumption, we assume, based on data \cite{Salazar}, that multiple
+assumption, we assume, based on data (Table~\ref{table}), that multiple
 infected individuals increase the likelihood of a positive pooled test
@@
 
-Thus, finding improved pooling schemes is imperative. An especially
+Thus, improving pooling schemes is imperative. An especially
 important consideration in designing such schemes is explicitly taking
 the intrinsic PCR error rates into account. We \cite{DOPE}, as well as
-others \cite{BayesianDorfman}, have developed such models, although their
-implementation in real-world settings has not yet commenced.
+others \cite{BayesianDorfman}, have developed such models, although they have not yet been implemented in real-world settings.
 

### 2021-03-31 comment.tex
add abstract

@@
 \begin{abstract}
-pooling is a popular strategy for increasing
-SARS-CoV-2 testing throughput. One popular pooling scheme is Dorfman
-pooling: test $N$ individuals simultaneously. If the test is positive, each individual is  then tested separately. When error rates are taken into consideration,
-most analyses assume that including extra infected samples in a pooled test has no effect 
-on the pooled test outcome. We refute this assumption with experimental data
-and suggest a novel probabilistic model for the outcomes of pooled tests. Our model is
-parsimonous and assumes as little as possible. We analyze the false-negative
-rate (i.e., the probability of a negative result for an infected individual) of
-Dorfman pooling via a our model. We show that one can expect a 60-80\% increase in
-false-negative rates under Dorfman pooling, for reasonable parameter
-values. Moreover, we show that the false-negative rates under Dorfman
-pooling increase when the prevalence of infection decreases.
-In most pooling schemes, identifying an infected
-individual requires positive results in multiple tests and hence
-substantially increases false-negative rates. Furthermore, this
-phenomenon is more pronounced when infection prevalence is low ---
-exactly when pooling is most efficient. Thus, pooling presents an
-inherent trade-off: it is most efficient when it is least
-accurate. The deterioration of false-negative rates and the
-aforementioned trade-off are inherent problems of pooling schemes and
-should be kept in mind by practitioners and policy makers.
+pooling is a popular strategy for increasing SARS-CoV-2 testing
+throughput. One popular pooling scheme is Dorfman pooling: test $N$
+individuals simultaneously. If the test is positive, each individual
+is then tested separately. Most analyses of pooling scheme error rates
+assume that including extra infected samples in a pooled test has no
+effect on the pooled test outcome. We refute this assumption with
+experimental data and suggest a novel and parsimonous probabilistic
+model for the outcomes of pooled tests. As an application, we analyze
+the false-negative rate (i.e., the probability of a negative result
+for an infected individual) of Dorfman pooling via a our model. We
+show that the false-negative rates under Dorfman pooling increase when
+the prevalence of infection decreases. Thus, pooling is riskiest
+exactly when it is most efficient. Conclusion/ summary.
+%% This is an inherent problem of Dorfman pooling and should be kept in
+%% mind by practitioners and policy makers alike.
 \end{abstract}

### 2021-03-31 comment.tex
abstract

@@
 \begin{abstract}
-pooling is a popular strategy for increasing SARS-CoV-2 testing
-throughput. One popular pooling scheme is Dorfman pooling: test $N$
+Pooling is a method of simultaneously testing multiple samples for the presence of pathogens. Pooling of RT-PCR tests of SARS-CoV-2 is increasing in popularity, due to its high testing
+throughput. A popular pooling scheme is Dorfman pooling: test $N$
 individuals simultaneously. If the test is positive, each individual
-is then tested separately. Most analyses of pooling scheme error rates
-assume that including extra infected samples in a pooled test has no
-effect on the pooled test outcome. We refute this assumption with
+is then tested separately; otherwise, all are declared negative. Most analyses of pooling scheme error rates
+assume that including more than a single infected sample in a pooled test does not increase the probability of a positive outcome. We challenge this assumption with
 experimental data and suggest a novel and parsimonous probabilistic
@@
 the false-negative rate (i.e., the probability of a negative result
-for an infected individual) of Dorfman pooling via a our model. We
+for an infected individual) of Dorfman pooling. We
 show that the false-negative rates under Dorfman pooling increase when
-the prevalence of infection decreases. Thus, pooling is riskiest
-exactly when it is most efficient. Conclusion/ summary.
+the prevalence of infection decreases.
+However, low infection prevalence is exactly the condition when Dorfman pooling achieves highest throughput efficiency.
+We therefore implore the cautious use of pooling and developement of pooling strategies that consider the tests' error rates.
 %% This is an inherent problem of Dorfman pooling and should be kept in

### 2021-03-31 comment.tex
uri commentz

@@
 assume that including more than a single infected sample in a pooled test does not increase the probability of a positive outcome. We challenge this assumption with
-experimental data and suggest a novel and parsimonous probabilistic
+experimental data and suggest a novel and parsimonious probabilistic
 model for the outcomes of pooled tests. As an application, we analyze
@@
 However, low infection prevalence is exactly the condition when Dorfman pooling achieves highest throughput efficiency.
-We therefore implore the cautious use of pooling and developement of pooling strategies that consider the tests' error rates.
+We therefore implore the cautious use of pooling and development of pooling strategies that consider the tests' error rates.
 %% This is an inherent problem of Dorfman pooling and should be kept in

### 2021-03-31 comment.tex
clean up, remove section on lion

@@
 \affil[*]{Equal contribution}
-%\email{yair.daon@gmail.com}
 \date{}
@@
 \begin{abstract}
-Pooling is a method of simultaneously testing multiple samples for the presence of pathogens. Pooling of RT-PCR tests of SARS-CoV-2 is increasing in popularity, due to its high testing
-throughput. A popular pooling scheme is Dorfman pooling: test $N$
-individuals simultaneously. If the test is positive, each individual
-is then tested separately; otherwise, all are declared negative. Most analyses of pooling scheme error rates
-assume that including more than a single infected sample in a pooled test does not increase the probability of a positive outcome. We challenge this assumption with
-experimental data and suggest a novel and parsimonious probabilistic
-model for the outcomes of pooled tests. As an application, we analyze
-the false-negative rate (i.e., the probability of a negative result
-for an infected individual) of Dorfman pooling. We
-show that the false-negative rates under Dorfman pooling increase when
-the prevalence of infection decreases.
-However, low infection prevalence is exactly the condition when Dorfman pooling achieves highest throughput efficiency.
-We therefore implore the cautious use of pooling and development of pooling strategies that consider the tests' error rates.
-%% This is an inherent problem of Dorfman pooling and should be kept in
-%% mind by practitioners and policy makers alike.
+Pooling is a method of simultaneously testing multiple samples for the
+presence of pathogens. Pooling of SARS-CoV-2 tests is increasing in
+popularity, due to its high testing throughput. A popular pooling
+scheme is Dorfman pooling: test $N$ individuals simultaneously. If the
+test is positive, each individual is then tested separately;
+otherwise, all are declared negative. Most analyses of pooling schemes
+error rates assume that including more than a single infected sample
+in a pooled test does not increase the probability of a positive
+outcome. We challenge this assumption with experimental data and
+suggest a novel and parsimonious probabilistic model for the outcomes
+of pooled tests. As an application, we analyze the false-negative rate
+(i.e., the probability of a negative result for an infected
+individual) of Dorfman pooling. We show that the false-negative rates
+under Dorfman pooling increase when the prevalence of infection
+decreases. However, low infection prevalence is exactly the condition
+when Dorfman pooling achieves highest throughput efficiency. We
+therefore implore the cautious use of pooling and development of
+pooling strategies that consider the tests' error rates.
 \end{abstract}
@@
 
-% \subsection{RT-PCR}\label{subsec:rtpcr}
-% We briefly explain the process of testing for SARS-CoV-2 by RT-PCR in
-% Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
-% details, and a comprehensive review of RT-PCR can be found in
-% \cite{COVID-RTPCR, PCRBook}.
-
-% PCR is a process in which a targeted DNA strand's frequency is
-% amplified to create billions of copies in a reaction mixture. At the
-% end of this \emph{amplification} process, the presence of targeted DNA
-% molecules in a reaction mixture can be confidently
-% determined. Arguably, the best way to understand PCR is as a chain
-% reaction: Each targeted DNA molecule is copied into two, that are
-% copied into four and so forth. This cascade gives PCR its name:
-% polymerase chain reaction.
-
-% Specifically, the PCR process is comprised of cycles, with each cycle
-% doubling the abundance of the targeted DNA. In each cycle, each
-% double-strand DNA molecule is broken into two separate strands. An
-% enzyme called DNA polymerase generates a new double-strand DNA replica
-% from each separate strand. Replication cannot start without a specific
-% short DNA sequence, called a primer, that has to be introduced into
-% the reaction mixture. Introducing the right primer into the reaction
-% mixture ensures (almost) only the targeted DNA sequence is copied. A
-% protein is added to the reaction mixture, which emits light upon
-% successfully binding the targeted DNA. Once enough light is emitted
-% the tested reaction mixture is declared positive and a
-% \emph{detection} is said to occur. If, on the other hand, a predefined
-% number of cycles pass without a detection event then the reaction
-% mixture is declared negative.
-
-% Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
-% extra preprocessing step is required. An enzyme called reverse
-% transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
-% reaction mixture into DNA. Once DNA replicas are made, the PCR process
-% proceeds as described above, targeting SARS-CoV-2 DNA replicas.
-
-%\subsection{Pooling}
-% One possible source of failed detection of SARS-CoV-2 RNA in pooled
-% RT-PCR testing (a \emph{false-negative}) is sample dilution. When
-% pooling, several samples are mixed, so the concentration of viral RNA
-% is reduced. This effect may cause a delay in amplification
-% \cite{DorfmanYuvalDor}, no detection and, consequently, a
-% false-negative. Another potential source for false-negatives is a
-% failed sampling process.
-
-%% However, \cite{Lion} showed that this effect can be
-%% safely ignored when mixing up to 32 samples, which we correspondingly
-%% assume.
-
 \section*{Methods}
-% We define and present the \emph{common assumption} in Section~\ref{subsec:common}
-% and refute it in Section~\ref{subsec:refute}. Our model is presented
-% in Section~\ref{subsec:ours}.
-
 Formally, we consider a pool containing $N$ individuals
@@
   assumption}.
-%% We refute the assumptions underlying \eqref{eq:common} in
-%% Section~\ref{subsec:ours} and suggest more plausible ones in
-%% Section~\ref{subsec:ours}.
 
@@
 this assumption. However, cross-reactivity was ruled out in
-\cite{KitComparison}.
-% To summarize, we assume that for a single pool, a positive test result
-% is generated in one of two paths. Either, SARS-CoV-2 RNA from an
-% infected individual's sample is correctly amplified and detected, and
-% this can happen for each positive sample in a pool, with probability
-% $\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
-% RNA is introduced), an event that occurs at most once per pool, with
-% probability $1-\Sp$.
-These assumptions lead to our model, which is illustrated in 
+\cite{KitComparison}. These assumptions lead to our model, which is illustrated in 
 Figure~\ref{fig:likelihood} and summarized in
-%% A negative pooled test result occurs when all detection paths (both
-%% correct and erroneous) fail. The probability of no false-detection
-%% accounts for the $\spec$ term. The probability of no correct detection
-%% is $(1-\sens)$ per infected individual. The probability that all such
-%% paths fail is the product of the above mentioned terms, displayed in
-\eqref{eq:likelihood a}.%%  The probability of a positive result,
-%% presented in \eqref{eq:likelihood b}, is simply the complement.
-%% The probability of a negative pooled test result is presented in
-%% \eqref{eq:likelihood a} (along with its complement
-%% \eqref{eq:likelihood b}), and explained below.
-%\begin{subequations}
-\begin{align}%\label{eq:single likelihood}
-  %\begin{split}
+\eqref{eq:likelihood a}.
+\begin{align}
     \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
     (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
-    %
-    %
-    %
-    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
-    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
-  %\end{split}
 \end{align}
-%\end{subequations}
-%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
-%% recalling that $\data \in \{0,1\}$ yields:
-
-%% \begin{align}\label{eq:combined likelihood}
-%%   \begin{split}
-%%     \Pr(\data| \theta) &= \left [\spec\prod_{i=1}^N
-%%       (1-\sens)^{\theta_i} \right ]^{1-\data} \left [1 - \spec\prod_{i=1}^N (1-\sens)^{\theta_i} \right ]^{\data}.
-%%   \end{split}
-%% \end{align}
 
@@
 
-%% \subsection{Scheme efficiency}
-%% The efficiency of Dorfman pooling scheme is the expected number of
-%% tests utilized. Let $T$ the number of tests. Then:
-%% \begin{align}
-%%   \begin{split}
-%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\sum_{k=0}^N \Pr(\sum \theta_i = k) \Pr(\data=1|\sum \theta_i = k) \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Sp(1-\Se)^k) \\
-%%     %
-%%     %
-%%     %
-%%     &= 1+N - N\Sp\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Se)^k \\
-%%     %
-%%     %
-%%     %
-%%     %&= 1+N - N\Sp \sum_{k=0}^N \binom{N}{k}(q-q\Se)^k(1-q)^k \\
-%%     %
-%%     %
-%%     %
-%%     &= 1+N - N\Sp (1-q\Se)^N\\    
-%%   \end{split}
-%% \end{align}
-
-%% Compare to the common assumption:
-
-%% \begin{align}
-%%   \begin{split}
-%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\Big[\Pr(\theta_i = 0,\ \forall i)(1-\Sp) + \Pr(\exists i\ \theta_i =1)\Se\Big] \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\Big [(1-q)^N(1-\Sp) + (1-(1-q)^N)\Se \Big ] \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + N\Big [(1-q)^N(1-\Sp - \Se) + \Se \Big ] \\
-%%     %
-%%     %
-%%     %
-%%     &= 1 + \Se N + N(1-q)^N(1-\Sp - \Se).
-%%   \end{split}
-%% \end{align}
-
-%% The difference is:
-%% \begin{equation}
-%%   \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
-%% \end{equation}
-
 \subsection*{Scheme false-negative rate}
@@
 to detect at least one positive result. Lastly, a \emph{scheme}
-false-negative occurs when an entire pooling scheme fails to identify
-Donald as infected. Our first task is to calculate Dorfman's scheme
-false-negative rate. Equivalently, what is the probability of not identifying Donald as
-infected under a Dorfman pooling scheme?
+false-negative occurs when an entire pooling scheme fails to
+identify Donald as infected. Our first task is to calculate Dorfman's
+scheme false-negative rate. Equivalently, what is the probability of
+not identifying Donald as infected under Dorfman pooling?
 
@@
 \section*{Discussion}
-%\subsection{Short recap}
 Here, we presented a novel probabilistic model for the outcomes of
@@
 
-% This is a fact almost disregarded in
-% \cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
-% when infection prevalence is low, such circumstances are exactly those
-% leading to a substantial increase in false-negative rates. Such
-% behavior of pooling is not implied by the common assumption.
-
-%\subsection{Implications for group testing in different prevalence settings}
 As the COVID-19 pandemic progresses, the infection prevalence
@@
 
-%\subsection{Limitations}
-%NONE
+%% Although our application focused on Dorfman pooling, our analysis
+%% applies to other pooling schemes. Pooling schemes
+%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
+%% positive pooled results to correctly identify an infected individual as
+%% such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
+%% If the first pool yields a positive result, it is split in two;
+%% otherwise, everyone are declared negative. Then the splitting is
+%% repeated until resulting pools are negative or individuals are tested
+%% separately. With an initial pool size of 32, Donald will necessarily
+%% have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
+%% in a single test, for the scheme to correctly identify him as
+%% infected. Compare this to the Dorfman scheme that requires a positive
+%% test in a pool of size $N=8$, and an additional single positive test
+%% to identify Donald as infected. The hierarchical pooling scheme of
+%% \cite{Lion, Kim} will necessarily yield more false-negatives than
+%% Dorfman pooling --- there are additional places for it to fail.
 
-%\subsection{Ways to reduce errors}
-Although our application focused on Dorfman pooling, our analysis
-applies to other pooling schemes. Pooling schemes
-(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
-positive pooled results to correctly identify an infected individual as
-such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-If the first pool yields a positive result, it is split in two;
-otherwise, everyone are declared negative. Then the splitting is
-repeated until resulting pools are negative or individuals are tested
-separately. With an initial pool size of 32, Donald will necessarily
-have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
-in a single test, for the scheme to correctly identify him as
-infected. Compare this to the Dorfman scheme that requires a positive
-test in a pool of size $N=8$, and an additional single positive test
-to identify Donald as infected. The hierarchical pooling scheme of
-\cite{Lion, Kim} will necessarily yield more false-negatives than
-Dorfman pooling --- there are additional places for it to fail.
+Thus, improving pooling schemes is imperative. An especially important
+consideration in designing such schemes is explicitly taking the
+intrinsic PCR error rates into account. We \cite{DOPE}, as well as
+others \cite{BayesianDorfman}, have developed such models, although
+they have not yet been implemented in real-world settings.
 
-Thus, improving pooling schemes is imperative. An especially
-important consideration in designing such schemes is explicitly taking
-the intrinsic PCR error rates into account. We \cite{DOPE}, as well as
-others \cite{BayesianDorfman}, have developed such models, although they have not yet been implemented in real-world settings.
-
-%\subsection{Conclusion}
-
-%% For example, although Dorfman pooling improves testing throughput,
-
-%% The common assumption implies that one assumes that the probability of
-%% a false-negative is identical for the pooled and single
-%% tests. Although simplistic in nature, the common assumption does capture
-%% the intuition behind our results: For Donald to be considered negative
-%% under Dorfman pooling, he has to test negative twice. In addition to
-%% this approximation only being an upper bound \cite{Simplistic2}, it
-%% does not account for the effect of infection prevalence on the
-%% false-negative rates under Dorfman pooling. Supporting the importance
-%% of this effect, such an association between prevalence and the scheme
-%% false-negative rate under Dorfman pooling has been empirically noted
-%% \cite{DorfmanYuvalDor}.
-
-%% Another important factor in pooling is dependence between samples. As
-%% mentioned in \cite{DorfmanYuvalDor}, introducing a positive dependence
-%% within a pool decreases the false-positive rate. In the extreme case,
-%% consider a fully connected pool, where one infection determines the
-%% entire pool is infected. In this case, a calculation analogous to the
-%% one conducted above (section??) recovers the initial false-negative
-%% rate $1-\Se$. Interestingly, pooling was also noted to have increased
-%% throughput when infection probabilities are dependent between the
-%% pooled individuals \cite{DorfmanYuvalDor}, providing another advantage
-%% to sampling dependent individuals in pooling schemes. Hence,
-%% deliberately pooling individuals with positive dependence of infection
-%% should be further studied and likely encouraged when performing
-%% pooling.
-
-%% Conclusion
 To conclude, pooling is an important technique that can improve

### 2021-03-31 comment.tex
close to final

@@
 therefore implore the cautious use of pooling and development of
-pooling strategies that consider the tests' error rates.
+pooling schemes that consider the correct tests' error rates.
 \end{abstract}
@@
 sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
-concern for pools as large as 64 individuals \cite{PoolSize30, Lion,
-  DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed that
-the probability of a true-positive (the test's \emph{sensitivity})
-does not depend on the number of infected samples in the pool, but
-rather on the existence of at least one such sample. Thus, the
-probability of a positive result in a pooled test has been (assumed)
-identical for a pool with one sample from an infected individual and,
-e.g., five such samples. This assumption is common in the group
-testing literature \cite{Kim, OptimalDorfmanPool}, as well as in more
-specific, COVID-19 focused studies \cite{Simplistic1, Simplistic2}. In
-this study we challenge these commonly made assumptions and show how
-using a more accurate probabilistic model affects estimation of
-false-negative rates for Dorfman pooling.
+concern, even for pools as large as 64 individuals \cite{PoolSize30,
+  Lion, DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed
+that the probability of a true-positive (the test's
+\emph{sensitivity}) does not depend on the number of infected samples
+in the pool, but rather on the existence of at least one such
+sample. Thus, the probability of a positive result in a pooled test
+has been (assumed) identical for a pool with one sample from an
+infected individual and, e.g., five such samples. This assumption is
+common in the group testing literature \cite{Kim, OptimalDorfmanPool},
+as well as in more specific, COVID-19 focused studies
+\cite{Simplistic1, Simplistic2}. In this study we challenge this
+commonly made assumption and show how using a more accurate
+probabilistic model affects estimation of false-negative rates for
+Dorfman pooling.
 
@@
 \subsection*{The common assumption}\label{subsec:common}
-Previous studies of group testing strategies assumed that the
-false-negative probability does not depend on the number of infected
-samples, but merely on the existence of at least one such sample in a
-pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
-context of SARS-CoV-2 also employ similar assumptions
-\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
-that the probability of a negative result is the same for a pool with
-a single sample coming from an infected individual and several
- samples from infected individuals. Explicitly:
+Previous studies of pooling schemes assumed that the false-negative
+probability does not depend on the number of infected samples, but
+merely on the existence of at least one such sample in a pool
+\cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
+context of SARS-CoV-2 also employ a similar assumption
+\cite{Simplistic1, Simplistic2}. %% Specifically, these studies assume
+%% that the probability of a negative result is the same for a pool with
+%% a single sample coming from an infected individual and several samples
+%% from infected individuals. 
+Explicitly, these studies assume:
 
@@
 amplification of viral RNA succeeds or fails for each sample
-independently. According to \cite{Simplistic1, Simplistic2, Kim,
-  OptimalDorfmanPool}, a false-positive does not
-depend on the number of negative samples in a pool. We incorporate
-this assumption in our model with a small modification: We assume that
-an erroneous amplification can occur in \emph{any} pool. Specifically,
-it is possible that correct amplifications fails and an erroneous one
-occurs simultaneously. This assumption is somewhat specific for the
-current application of screening for SARS-CoV-2 via RT-PCR. For
-example, cross-reactivity with other coronaviruses would have violated
-this assumption. However, cross-reactivity was ruled out in
-\cite{KitComparison}. These assumptions lead to our model, which is illustrated in 
-Figure~\ref{fig:likelihood} and summarized in
-\eqref{eq:likelihood a}.
-\begin{align}
-    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
-    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
+independently. Furthermore, according to \cite{Simplistic1,
+  Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
+depend on the number of negative samples in a pool either. Following
+our previous reasoning, we incorporate this assumption in our model
+with a small modification: We assume that a false-positive also occurs
+indpendently of other amplifications. This means a false-positive can
+occur in \emph{any} pool. Specifically, it is possible that every
+correct amplifications fails and an erroneous one occurs
+simultaneously. This assumption is somewhat specific for the current
+application of screening for SARS-CoV-2 via RT-PCR. For example,
+cross-reactivity with other coronaviruses would have violated this
+assumption. However, cross-reactivity was ruled out in
+\cite{KitComparison}. These assumptions lead to our model, which is
+illustrated in Figure~\ref{fig:likelihood} and summarized in
+\eqref{eq:likelihood}.
+\begin{align}\label{eq:likelihood}
+    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N (1-\sens)^{\theta_i} =
+    \spec(1-\Se)^{\sum_i\theta_i}.
 \end{align}
@@
     individuals $1$ and $2$, each with probability $(1-\sens)$.
-    Additionally, no false-positive detection of SARS-CoV-2 occurs,
-    with probability $\spec$. Individual $3$ is not infected and does
-    not contribute to the probability of the pooled test result.}\label{fig:likelihood}
+    Additionally, no false-positive detection occurs, with probability
+    $\spec$. Individual $3$ is not infected and does not contribute to
+    the probability of the pooled test result.}\label{fig:likelihood}
 \end{figure}
 
-\subsection*{Scheme false-negative rate}
+\subsection*{Application: scheme false-negative rate}
 We calculate the false-negative rate for a single \emph{infected}
-individual, henceforth referred to as "Donald" under our model and
+individual, henceforth referred to as "Donald", under our model and
 under the common assumption. We distinguish three types of
@@
 to detect at least one positive result. Lastly, a \emph{scheme}
-false-negative occurs when an entire pooling scheme fails to
-identify Donald as infected. Our first task is to calculate Dorfman's
-scheme false-negative rate. Equivalently, what is the probability of
+false-negative occurs when an entire pooling scheme fails to identify
+Donald as infected. Our first task is to calculate Dorfman's scheme
+false-negative rate. Equivalently, we ask: what is the probability of
 not identifying Donald as infected under Dorfman pooling?
@@
 Denote the prevalence of infection in the (tested) population $q$. We
-assign Donald as individual $1$, so that $\theta_1=1$. Then:
+denote Donald as individual $1$, so that $\theta_1=1$. Then:
 
@@
     \begin{split}
-        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
+        \mi :&= 1 - \Se\Pr(\data=1|\theta_1=1)\\
         %
@@
 According to the common assumption, the scheme false-negative rate is
-$1-\Se^2$. A short calculation shows that this approximation implies
-the percentage increase in scheme false-negative rate is $\err{common}
-:= 100\cdot \Se$.
+$1-\Se^2$. A short calculation shows that this implies the percentage
+increase in scheme false-negative rate is $\err{common} := 100\cdot
+\Se$.
 
@@
 in various tested populations undergoes frequent changes. Hence, as
-our results suggest, pooling strategies employed for mass testing should be
+our results suggest, pooling schemes employed for mass testing should be
 used with caution in populations where infection rates are low. Such

### 2021-04-01 comment.tex
before amit checks

@@
 \begin{document}
-\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors}
+\title{An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors}
 
@@
 popularity, due to its high testing throughput. A popular pooling
-scheme is Dorfman pooling: test $N$ individuals simultaneously. If the
+scheme is Dorfman pooling: test $N$ individuals simultaneously, if the
 test is positive, each individual is then tested separately;
-otherwise, all are declared negative. Most analyses of pooling schemes
-error rates assume that including more than a single infected sample
+otherwise, all are declared negative. Most analyses of the error rates of pooling schemes
+assume that including more than a single infected sample
 in a pooled test does not increase the probability of a positive
@@
 therefore implore the cautious use of pooling and development of
-pooling schemes that consider the correct tests' error rates.
+pooling schemes that consider correctly accounting for tests' error rates.
 \end{abstract}
@@
 \{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
-test's sensitivity (true-negative rate) is denoted $\sens$, and the
-test's specificity (true-positive rate) is denoted $\spec$. Pooled
+test's sensitivity (true-positive rate) is denoted $\sens$, and the
+test's specificity (true-negative rate) is denoted $\spec$. Pooled
 test result (data) is denoted $\data \in \{0,1\}$, where $\data=0$ iff
@@
 context of SARS-CoV-2 also employ a similar assumption
-\cite{Simplistic1, Simplistic2}. %% Specifically, these studies assume
-%% that the probability of a negative result is the same for a pool with
-%% a single sample coming from an infected individual and several samples
-%% from infected individuals. 
-Explicitly, these studies assume:
+\cite{Simplistic1, Simplistic2}. Explicitly, these studies assume:
 
@@
   Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
-depend on the number of negative samples in a pool either. Following
-our previous reasoning, we incorporate this assumption in our model
-with a small modification: We assume that a false-positive also occurs
-indpendently of other amplifications. This means a false-positive can
-occur in \emph{any} pool. Specifically, it is possible that every
-correct amplifications fails and an erroneous one occurs
+depend on the number of negative samples in a pool either. For lack of data pointing otherwise, we incorporate this assumption into our model
+with a small modification. We do assume that a false amplification can occur only once per pool. However, we also assume
+false amplification is independent of any other correct amplification. Specifically, it is possible that every
+correct amplification fails \emph{and} an erroneous one occurs
 simultaneously. This assumption is somewhat specific for the current
@@
 depend on infection prevalence. Specifically, low values of infection
-prevalence increase $\err{exact}$. These results remain qualitatively
+prevalence increase the false-negative rates of Dorfman pooling. These results remain qualitatively
 similar under varying parameter values, in the observed ranges
@@
   InterpretingCOVID19Test} (Figure \ref{fig1}).
-Our results lead to a conflict: low infection prevalence leads to high
+Our results give rise to a conflict: low infection prevalence leads to high
 efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
@@
 mass-tested populations often include air travel passengers
-\cite{JTM,RobinHood}, potentially importing infections, or
-presymptomatic and asymptomatic individuals, and can be crucial for
+\cite{JTM} or presymptomatic and asymptomatic individuals \cite{RobinHood}, and can be crucial for
 controlling outbreaks \cite{MinaScience}.
-
-%% Although our application focused on Dorfman pooling, our analysis
-%% applies to other pooling schemes. Pooling schemes
-%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
-%% positive pooled results to correctly identify an infected individual as
-%% such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
-%% If the first pool yields a positive result, it is split in two;
-%% otherwise, everyone are declared negative. Then the splitting is
-%% repeated until resulting pools are negative or individuals are tested
-%% separately. With an initial pool size of 32, Donald will necessarily
-%% have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
-%% in a single test, for the scheme to correctly identify him as
-%% infected. Compare this to the Dorfman scheme that requires a positive
-%% test in a pool of size $N=8$, and an additional single positive test
-%% to identify Donald as infected. The hierarchical pooling scheme of
-%% \cite{Lion, Kim} will necessarily yield more false-negatives than
-%% Dorfman pooling --- there are additional places for it to fail.
 

### 2021-04-21 paper.tex
commit of stuff

--- 2021-04-01 comment.tex
@@
 test is positive, each individual is then tested separately;
-otherwise, all are declared negative. Most analyses of the error rates of pooling schemes
-assume that including more than a single infected sample
-in a pooled test does not increase the probability of a positive
-outcome. We challenge this assumption with experimental data and
-suggest a novel and parsimonious probabilistic model for the outcomes
-of pooled tests. As an application, we analyze the false-negative rate
-(i.e., the probability of a negative result for an infected
-individual) of Dorfman pooling. We show that the false-negative rates
-under Dorfman pooling increase when the prevalence of infection
-decreases. However, low infection prevalence is exactly the condition
-when Dorfman pooling achieves highest throughput efficiency. We
-therefore implore the cautious use of pooling and development of
-pooling schemes that consider correctly accounting for tests' error rates.
+otherwise, all are declared negative. Most analyses of the error rates
+of pooling schemes assume that including more than a single infected
+sample in a pooled test does not increase the probability of a
+positive outcome. We challenge this assumption with experimental data
+and suggest a novel and parsimonious probabilistic model for the
+outcomes of pooled tests. As an application, we analyze the
+false-negative rate (i.e., the probability of a negative result for an
+infected individual) of Dorfman pooling. We show that the
+false-negative rates under Dorfman pooling increase when the
+prevalence of infection decreases. However, low infection prevalence
+is exactly the condition when Dorfman pooling achieves highest
+throughput efficiency. We therefore implore the cautious use of
+pooling and development of pooling schemes that consider correctly
+accounting for tests' error rates.
 \end{abstract}
@@
   Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
-depend on the number of negative samples in a pool either. For lack of data pointing otherwise, we incorporate this assumption into our model
-with a small modification. We do assume that a false amplification can occur only once per pool. However, we also assume
-false amplification is independent of any other correct amplification. Specifically, it is possible that every
-correct amplification fails \emph{and} an erroneous one occurs
-simultaneously. This assumption is somewhat specific for the current
-application of screening for SARS-CoV-2 via RT-PCR. For example,
-cross-reactivity with other coronaviruses would have violated this
-assumption. However, cross-reactivity was ruled out in
-\cite{KitComparison}. These assumptions lead to our model, which is
+depend on the number of negative samples in a pool either. For lack of
+data pointing otherwise, we incorporate this assumption into our model
+with a small modification. We do assume that a false amplification can
+occur only once per pool. However, we also assume false amplification
+is independent of any other correct amplification. Specifically, it is
+possible that every correct amplification fails \emph{and} an
+erroneous one occurs simultaneously. This assumption is somewhat
+specific for the current application of screening for SARS-CoV-2 via
+RT-PCR. For example, cross-reactivity with other coronaviruses would
+have violated this assumption. However, cross-reactivity was ruled out
+in \cite{KitComparison}. These assumptions lead to our model, which is
 illustrated in Figure~\ref{fig:likelihood} and summarized in
@@
 
+\subsection*{Code Availability}
+Code for this manuscript is available at
+\url{https://github.com/yairdaon/errorates}.
+
 \section*{Results}\label{section:results}
@@
 
+\section*{Acknowledgements}
+Yair Daon was supported by a post-doctoral fellowship from the Tel
+Aviv University Center for Combating Pandemics and the Raymond and Beverly
+Sackler dean's post-doctoral fellowship.
+
 \bibliographystyle{amsplain}

### 2021-09-09 comment.tex
Updates from Overleaf

@@
 
-\affil[3]{The Gertner Institute for Epidemiology and Health Policy
-  Research, Tel Hashomer, Israel}
+\affil[3]{The Biostatistics and Biomathematics Unit, The
+Gertner Institute for Epidemiology and Health Policy Research, Sheba Medical Center, Tel Hashomer, 52621 Ramat Gan, Israel}
 \affil[*]{Equal contribution}
@@
 \begin{abstract}
-Pooling is a method of simultaneously testing multiple samples for the
-presence of pathogens. Pooling of SARS-CoV-2 tests is increasing in
-popularity, due to its high testing throughput. A popular pooling
-scheme is Dorfman pooling: test $N$ individuals simultaneously, if the
-test is positive, each individual is then tested separately;
-otherwise, all are declared negative. Most analyses of the error rates of pooling schemes
+PCR testing is an important tool to mitigate outbreaks of infectious diseases.
+One way of increasing testing throughput is by simultaneously testing multiple samples for the presence of a pathogen, a technique known as \emph{pooling}.
+During the current COVID-19 pandemic, rapidly testing individuals for the presence of SARS-CoV-2 is conducted in large amounts.
+Since testing is often a bottleneck in mitigating the spread of SARS-CoV-2, pooling is increasing in popularity.
+Most analyses of the error rates of pooling schemes
 assume that including more than a single infected sample
@@
 outcome. We challenge this assumption with experimental data and
-suggest a novel and parsimonious probabilistic model for the outcomes
-of pooled tests. As an application, we analyze the false-negative rate
-(i.e., the probability of a negative result for an infected
-individual) of Dorfman pooling. We show that the false-negative rates
-under Dorfman pooling increase when the prevalence of infection
+suggest a novel probabilistic model for the outcomes
+of pooled tests. As an application, we analyze the false-negative rates
+ of one common pooling scheme known as Dorfman pooling. We show that the false-negative rates of Dorfman pooling
+increase when the prevalence of infection
 decreases. However, low infection prevalence is exactly the condition
-when Dorfman pooling achieves highest throughput efficiency. We
+under which Dorfman pooling achieves highest throughput. We
 therefore implore the cautious use of pooling and development of
@@
 Since the essence of the refuted common assumption is that
-amplification of all samples occurs only once, we assume the opposite:
+amplification of all samples occurs only once, we assume a more realistic model:
 amplification of viral RNA succeeds or fails for each sample
@@
   Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
-depend on the number of negative samples in a pool either. For lack of data pointing otherwise, we incorporate this assumption into our model
+depend on the number of negative samples in a pool. For lack of data pointing otherwise, we incorporate this assumption into our model
 with a small modification. We do assume that a false amplification can occur only once per pool. However, we also assume
@@
 assumption. However, cross-reactivity was ruled out in
-\cite{KitComparison}. These assumptions lead to our model, which is
+\cite{KitComparison}. These assumptions lead to the following  model, which is
 illustrated in Figure~\ref{fig:likelihood} and summarized in
@@
   \end{tikzpicture}
-  \caption{Illustration of our model. A pool contains individuals
+  \caption{Illustration of the revised probabilistic model. A pool contains individuals
     $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
@@
 If the pooled test yields a positive result, Donald is tested
-separately. Taking a conservative stand, we assume such a simple
+separately. Taking a conservative stand,  it is assumed that such a simple
 procedure poses no risk of introducing contaminant RNA. Therefore, the
@@
 According to the common assumption, the scheme false-negative rate is
-$1-\Se^2$. A short calculation shows that this implies the percentage
+$1-\Se^2$. A straight forward calculation shows that this implies the percentage
 increase in scheme false-negative rate is $\err{common} := 100\cdot
@@
     percentage increase in the scheme false-negative rates relative
-    to the single test false-negative rates (eq
-    \eqref{eq:erel}). Right: colors represent the difference between
+    to the single test false-negative rates \eqref{eq:erel}. Right: colors represent the difference between
     $\err{common}$ and $\err{exact}$. The disease prevalence $q$, is
@@
     the y-axis. Pool size $N$, was chosen according to $q$ as in
-    \cite{DorfmanYuvalDor}.}\label{fig1}
+    \cite{DorfmanYuvalDor}. The left panel shows that $\err{exact}$ is largest for low prevalence values $q$. The difference between $\err{common}$ and $\err{exact}$
+    can be as large as 30\%, as seen in the right panel. }\label{fig1}
 \end{figure}
@@
 \section*{Discussion}
-Here, we presented a novel probabilistic model for the outcomes of
+Here, we developed a more realistic and novel probabilistic model for the outcomes of
 pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
@@
 
-To conclude, pooling is an important technique that can improve
+To conclude, pooling is an important technique that can increase
 testing throughput in a cost-effective manner. Nevertheless, care must
@@
 low infection prevalence settings.
+
+\section*{Acknowledgements}
+Yair Daon was supported by a post-doctoral fellowship from the Tel Aviv University Center for Combating Pandemics and the Raymond and Beverly Sackler dean's
+post-doctoral fellowship.
 

### 2021-09-09 paper.tex

@@
 Since the essence of the refuted common assumption is that
-amplification of all samples occurs only once, we assume the opposite:
-amplification of viral RNA succeeds or fails for each sample
-independently. Furthermore, according to \cite{Simplistic1,
+amplification of all samples occurs only once, we assume a more
+realistic model: amplification of viral RNA succeeds or fails for each
+sample independently. Furthermore, according to \cite{Simplistic1,
   Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
@@
   \end{tikzpicture}
-  \caption{Illustration of our model. A pool contains individuals
-    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
-    are infected). A negative pooled test ($\data=0$) occurs when
-    three detection paths fail: A false-negative occurs for
-    individuals $1$ and $2$, each with probability $(1-\sens)$.
-    Additionally, no false-positive detection occurs, with probability
-    $\spec$. Individual $3$ is not infected and does not contribute to
-    the probability of the pooled test result.}\label{fig:likelihood}
+  \caption{Illustration of the revised probabilistic model. A pool
+    contains individuals $\{1,2,3\}$ with state $\theta=(1,1,0)$
+    (i.e., individuals 1 and 2 are infected). A negative pooled test
+    ($\data=0$) occurs when three detection paths fail: A
+    false-negative occurs for individuals $1$ and $2$, each with
+    probability $(1-\sens)$.  Additionally, no false-positive
+    detection occurs, with probability $\spec$. Individual $3$ is not
+    infected and does not contribute to the probability of the pooled
+    test result.}\label{fig:likelihood}
 \end{figure}
@@
 If the pooled test yields a positive result, Donald is tested
-separately. Taking a conservative stand, we assume such a simple
-procedure poses no risk of introducing contaminant RNA. Therefore, the
-separate test yields a positive result with probability $\Se$.
+separately. Taking a conservative stand, it is assumed that such a
+simple procedure poses no risk of introducing contaminant
+RNA. Therefore, the separate test yields a positive result with
+probability $\Se$.
 
@@
 not infected, henceforth referred to as the scheme's false-negative
-rate and denoted $\mi$. In order to correctly identify an infected individual
-as infected, both pooled and separate tests have to yield a positive
-result. Thus, the scheme's false-negative rate is:
+rate and denoted $\mi$. In order to correctly identify an infected
+individual as infected, both pooled and separate tests have to yield a
+positive result. Thus, the scheme's false-negative rate is:
 \begin{align}\label{eq:sfn}
@@
 According to the common assumption, the scheme false-negative rate is
-$1-\Se^2$. A short calculation shows that this implies the percentage
-increase in scheme false-negative rate is $\err{common} := 100\cdot
-\Se$.
+$1-\Se^2$. A straight forward calculation shows that this implies the
+percentage increase in scheme false-negative rate is $\err{common} :=
+100\cdot \Se$.
 
@@
     $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
-    percentage increase in the scheme false-negative rates relative
-    to the single test false-negative rates (eq
-    \eqref{eq:erel}). Right: colors represent the difference between
-    $\err{common}$ and $\err{exact}$. The disease prevalence $q$, is
-    varied on the x-axis, while the test sensitivity is varied on
-    the y-axis. Pool size $N$, was chosen according to $q$ as in
-    \cite{DorfmanYuvalDor}.}\label{fig1}
+    percentage increase in the scheme false-negative rates relative to
+    the single test false-negative rates \eqref{eq:erel}. Right:
+    colors represent the difference between $\err{common}$ and
+    $\err{exact}$. The disease prevalence $q$, is varied on the
+    x-axis, while the test sensitivity is varied on the y-axis. Pool
+    size $N$, was chosen according to $q$ as in
+    \cite{DorfmanYuvalDor}. The left panel shows that $\err{exact}$ is
+    largest for low prevalence values $q$. The difference between
+    $\err{common}$ and $\err{exact}$ can be as large as 30\%, as seen
+    in the right panel. }\label{fig1}
 \end{figure}
@@
 \section*{Discussion}
-Here, we presented a novel probabilistic model for the outcomes of
-pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
-assumption, we assume, based on data (Table~\ref{table}), that multiple
-infected individuals increase the likelihood of a positive pooled test
-result. A direct consequence of our model is that false-negative rates
-depend on infection prevalence. Specifically, low values of infection
-prevalence increase the false-negative rates of Dorfman pooling. These results remain qualitatively
-similar under varying parameter values, in the observed ranges
-\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
-  InterpretingCOVID19Test} (Figure \ref{fig1}).
-Our results give rise to a conflict: low infection prevalence leads to high
-efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
-increasing false-negative rates.
+Here, we developed a more realistic and novel probabilistic model for
+the outcomes of pooled PCR tests, parametrized for
+SARS-CoV-2. Contrary to the common assumption, we assume, based on
+data (Table~\ref{table}), that multiple infected individuals increase
+the likelihood of a positive pooled test result. A direct consequence
+of our model is that false-negative rates depend on infection
+prevalence. Specifically, low values of infection prevalence increase
+the false-negative rates of Dorfman pooling. These results remain
+qualitatively similar under varying parameter values, in the observed
+ranges \cite{KitComparison,EstimatingRatesKucrika,
+  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
+\ref{fig1}).  Our results give rise to a conflict: low infection
+prevalence leads to high efficiency of Dorfman pooling
+\cite{DorfmanYuvalDor}, while also increasing false-negative rates.
 
-As the COVID-19 pandemic progresses, the infection prevalence
-in various tested populations undergoes frequent changes. Hence, as
-our results suggest, pooling schemes employed for mass testing should be
+As the COVID-19 pandemic progresses, the infection prevalence in
+various tested populations undergoes frequent changes. Hence, as our
+results suggest, pooling schemes employed for mass testing should be
 used with caution in populations where infection rates are low. Such
-mass-tested populations often include air travel passengers
-\cite{JTM} or presymptomatic and asymptomatic individuals \cite{RobinHood}, and can be crucial for
-controlling outbreaks \cite{MinaScience}.
+mass-tested populations often include air travel passengers \cite{JTM}
+or presymptomatic and asymptomatic individuals \cite{RobinHood}, and
+can be crucial for controlling outbreaks \cite{MinaScience}.
 
@@
 Yair Daon was supported by a post-doctoral fellowship from the Tel
-Aviv University Center for Combating Pandemics and the Raymond and Beverly
-Sackler dean's post-doctoral fellowship.
+Aviv University Center for Combating Pandemics and the Raymond and
+Beverly Sackler dean's post-doctoral fellowship.
 

### 2021-09-09 paper.tex
with some review corrections

@@
 COVID-19 pandemic. As such, the need for large-scale testing has
-resulted in the development of techniques to increase the the
-throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
-  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques,
-often termed \emph{pooling}, or group testing, have originated in the
-seminal work of Dorfman and his pooling technique
-\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
-selects $N$ individuals and performs a single RT-PCR test on their
-combined (\emph{pooled}) samples. If the pooled test yields a positive
-result, then each individual is retested separately; otherwise,
-everyone is declared negative. The throughput efficiency of Dorfman
-pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
-its error rates thoroguhly investigated \cite{Kim, Simplistic1,
-  OptimalDorfmanPool}.
+resulted in the development of techniques to increase the throughput
+of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
+  MatrixPooling, LionDorfman}. These techniques, often termed
+\emph{pooling}, or group testing, have originated in the seminal work
+of Dorfman and his pooling technique \cite{DorfmanOriginal,
+  DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$ individuals
+and performs a single RT-PCR test on their combined (\emph{pooled})
+samples. If the pooled test yields a positive result, then each
+individual is retested separately; otherwise, everyone is declared
+negative. The throughput efficiency of Dorfman pooling has been
+demonstrated empirically \cite{DorfmanYuvalDor} and its error rates
+thoroughly investigated \cite{Kim, Simplistic1, OptimalDorfmanPool}.
 
@@
 
-Below, we refer to \eqref{eq:common} as the \emph{common
+Below, we refer to equation \eqref{eq:common} as the \emph{common
   assumption}.
@@
 illustrated in Figure~\ref{fig:likelihood} and summarized in
-\eqref{eq:likelihood}.
+equation \eqref{eq:likelihood}.
 \begin{align}\label{eq:likelihood}
@@
     %
-    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)
-    \Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
+    &= \sum_{k=0}^{N-1}\Pr\left(\sum_{i=2}^N\theta_i=k\right)
+    \Pr\left(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k\right)\\
     %
@@
 \begin{equation}\label{eq:erel}
-\err{exact} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
+\err{exact} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100\%.
 \end{equation}
@@
 percentage increase in scheme false-negative rate is $\err{common} :=
-100\cdot \Se$.
+100\%\cdot \Se$.
 
@@
   InterpretingCOVID19Test, EstimatingRatesLourenco,
-  FalsePositiveEstimate}, an increase of at least $60\%$ in
-$\err{exact}$ can be expected (Figure \ref{fig1}).
+  FalsePositiveEstimate}, the increase in scheme false-negative rate
+(compared to $1-\Se$), summarized in $\err{exact}$, is at leaset
+$60\%$ (Figure \ref{fig1}).
 
@@
 decreases the scheme false-negative rate, as can also be easily seen
-from \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
+from equation \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
 the single test false-negative rates increases the relative error
@@
     percentage increase in the scheme false-negative rates relative to
-    the single test false-negative rates \eqref{eq:erel}. Right:
+    the single test false-negative rates equation \eqref{eq:erel}. Right:
     colors represent the difference between $\err{common}$ and
@@
 consideration in designing such schemes is explicitly taking the
-intrinsic PCR error rates into account. We \cite{DOPE}, as well as
-others \cite{BayesianDorfman}, have developed such models, although
-they have not yet been implemented in real-world settings.
+intrinsic PCR error rates into account. As we have demonstrated above,
+analyzing expected error rates of a given scheme, given some error
+model is relatively straightforward. Designing a pooling scheme
+tailored for a specific error model is, however, far from trivial. In
+a recent study \cite{DOPE}, we have developed a novel pooling scheme
+that takes advantage of this specific error model. Our new pooling
+scheme shows great promise in terms of both accuracy, as well as
+efficiency, although it has not yet been implemented in real-world
+settings.
 

### 2021-09-09 paper.tex
added some stuff in the intro

--- 2021-09-09 paper.tex
@@
 COVID-19 pandemic. As such, the need for large-scale testing has
-resulted in the development of techniques to increase the the
+resulted in the application of techniques to increase the the
 throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
   BayesianDorfman, MatrixPooling, LionDorfman}. These techniques,
-often termed \emph{pooling}, or group testing, have originated in the
+often termed \emph{pooling}, or group testing, have originated in the 1940's in the
 seminal work of Dorfman and his pooling technique
@@
 its error rates thoroguhly investigated \cite{Kim, Simplistic1,
-  OptimalDorfmanPool}.
+  OptimalDorfmanPool}. Other commonly used techniques include binary pooling, wherein... Matrix pooling.. many other complex techniques..
 
@@
 commonly made assumption and show how using a more accurate
-probabilistic model affects estimation of false-negative rates for
+probabilistic model, which affects the estimation of false-negative rates for
 Dorfman pooling.

### 2021-09-09 paper.tex
discussion stuff

--- 2021-09-09 paper.tex
@@
 COVID-19 pandemic. As such, the need for large-scale testing has
-resulted in the development of techniques to increase the throughput
+resulted in the application of techniques to increase the throughput
 of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
   MatrixPooling, LionDorfman}. These techniques, often termed
-\emph{pooling}, or group testing, have originated in the seminal work
-of Dorfman and his pooling technique \cite{DorfmanOriginal,
-  DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$ individuals
-and performs a single RT-PCR test on their combined (\emph{pooled})
-samples. If the pooled test yields a positive result, then each
-individual is retested separately; otherwise, everyone is declared
-negative. The throughput efficiency of Dorfman pooling has been
-demonstrated empirically \cite{DorfmanYuvalDor} and its error rates
-thoroughly investigated \cite{Kim, Simplistic1, OptimalDorfmanPool}.
+\emph{pooling}, or group testing, have originated in the 1940's in the
+seminal work of Dorfman and his pooling technique
+\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
+selects $N$ individuals and performs a single RT-PCR test on their
+combined (\emph{pooled}) samples. If the pooled test yields a positive
+result, then each individual is retested separately; otherwise,
+everyone is declared negative. The throughput efficiency of Dorfman
+pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
+its error rates thoroguhly investigated \cite{Kim, Simplistic1,
+  OptimalDorfmanPool}. Other commonly used techniques include
+recursive pooling, wherein... Matrix pooling.. many other complex
+techniques..
 
@@
 commonly made assumption and show how using a more accurate
-probabilistic model affects estimation of false-negative rates for
-Dorfman pooling.
+probabilistic model affects the estimation of false-negative
+rates for Dorfman pooling.
+
+Several consequences arise from our analysis. Estimation of optimal
+pool size (for Dorfman pooling), which relies on previously made
+asumptions should be revisited \cite{OptimalDorfmanPool}. The error
+rates of any pooling scheme in use should be reesetimated
+\cite{BayesianDorfman, Kim}. And, finally, new methods should be
+developed in order to take advantage of our novel model \cite{DOPE}.
 
@@
 \section*{Discussion}
-Here, we developed a more realistic and novel probabilistic model for
-the outcomes of pooled PCR tests, parametrized for
+In this study, we developed a more realistic and novel probabilistic
+model for the outcomes of pooled PCR tests, parametrized for
 SARS-CoV-2. Contrary to the common assumption, we assume, based on
@@
 the likelihood of a positive pooled test result. A direct consequence
-of our model is that false-negative rates depend on infection
-prevalence. Specifically, low values of infection prevalence increase
-the false-negative rates of Dorfman pooling. These results remain
-qualitatively similar under varying parameter values, in the observed
-ranges \cite{KitComparison,EstimatingRatesKucrika,
+of our model is that low values of infection
+prevalence increase the false-negative rates of Dorfman pooling. Importantly, our results remain qualitatively similar under varying parameter values,
+in the observed ranges \cite{KitComparison,EstimatingRatesKucrika,
   EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
-\ref{fig1}). Our results give rise to a conflict: low infection
+\ref{fig1}). Therefore, our results give rise to a conflict: low infection
 prevalence leads to high efficiency of Dorfman pooling
@@
 
-Thus, improving pooling schemes is imperative. An especially important
-consideration in designing such schemes is explicitly taking the
-intrinsic PCR error rates into account. As we have demonstrated above,
-analyzing expected error rates of a given scheme, given some error
-model is relatively straightforward. Designing a pooling scheme
-tailored for a specific error model is, however, far from trivial. In
-a recent study \cite{DOPE}, we have developed a novel pooling scheme
-that takes advantage of this specific error model. Our new pooling
-scheme shows great promise in terms of both accuracy, as well as
-efficiency, although it has not yet been implemented in real-world
-settings.
+Furthermore, an especially important
+consideration in designing pooling schemes is explicitly taking the
+intrinsic PCR error rates into account.
+As we have demonstrated above, analyzing the expected error rates of pooling schemes under a realistic probabilistic error model is readily achievable.
+Designing pooling schemes which optimize testing while accounting for error rates is, however, far from trivial.
+Nonetheless, various algorithms explicitly incorporate error rates into their pooling optimization processes (REFS), including a recent study by our group which employed the herein described error model \cite{DOPE}.
 

### 2021-09-09 paper.tex
more discussion

@@
 recursive pooling, wherein... Matrix pooling.. many other complex
-techniques..
+techniques.
 
@@
 Nonetheless, various algorithms explicitly incorporate error rates into their pooling optimization processes (REFS), including a recent study by our group which employed the herein described error model \cite{DOPE}.
+In our proposed algorithm, we have used a Bayesian experimental design criterion to create an algorithm which iteratively performs pooling schemes, updating its information based on previous results.
+In doing so, our pooling
+scheme shows potential in terms of both reducing false-negative rates, as well as increasing
+efficiency, although it has not yet been implemented in real-world
+settings.
 

### 2021-09-09 paper.tex
before merging uris changes

--- 2021-09-09 paper.tex
@@
 
-\affil[3]{The Biostatistics and Biomathematics Unit, The
-Gertner Institute for Epidemiology and Health Policy Research, Sheba Medical Center, Tel Hashomer, 52621 Ramat Gan, Israel}
+\affil[3]{The Biostatistics and Biomathematics Unit, The Gertner
+  Institute for Epidemiology and Health Policy Research, Sheba Medical
+  Center, Tel Hashomer, 52621 Ramat Gan, Israel}
+
 \affil[*]{Equal contribution}
+
 \date{}
@@
 of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
-  MatrixPooling, LionDorfman}. These techniques, often termed
-\emph{pooling}, or group testing, have originated in the 1940's in the
-seminal work of Dorfman and his pooling technique
+  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
+termed \emph{pooling}, or group testing, have originated in the 1940's
+in the seminal work of Dorfman and his pooling technique
 \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
@@
   OptimalDorfmanPool}. Other commonly used techniques include
-recursive pooling, wherein... Matrix pooling.. many other complex
-techniques..
+\emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, wherein if the
+first pooled test is positive, the pool is split into two and the
+process repeats. Otherwise, all pool members are declared negative.
+In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
+$N=mn$ is arranged in a matrix of dimensions $m\times n$ matrix. Each
+row and column are then pooled and individuals in the intersection of
+positive rows and columns are tested separately.
 
@@
 concern, even for pools as large as 64 individuals \cite{PoolSize30,
-  Lion, DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed
-that the probability of a true-positive (the test's
+  Lion, DorfmanYuvalDor, DilutionCOVID, CherifReview}. Consequently,
+studies assumed that the probability of a true-positive (the test's
 \emph{sensitivity}) does not depend on the number of infected samples
@@
 infected individual and, e.g., five such samples. This assumption is
-common in the group testing literature \cite{Kim, OptimalDorfmanPool},
-as well as in more specific, COVID-19 focused studies
+common in the group testing literature \cite{Kim, OptimalDorfmanPool,
+  CherifReview}, as well as in more specific, COVID-19 focused studies
 \cite{Simplistic1, Simplistic2}. In this study we challenge this
 commonly made assumption and show how using a more accurate
-probabilistic model affects the estimation of false-negative
-rates for Dorfman pooling.
+probabilistic model affects the estimation of false-negative rates for
+Dorfman pooling.
 
@@
 the likelihood of a positive pooled test result. A direct consequence
-of our model is that low values of infection
-prevalence increase the false-negative rates of Dorfman pooling. Importantly, our results remain qualitatively similar under varying parameter values,
+of our model is that false-negative rates depend on infection
+prevalence. A further consequence is that low values of infection
+prevalence increase the false-negative rates of Dorfman pooling. These
+results remain qualitatively similar under varying parameter values,
 in the observed ranges \cite{KitComparison,EstimatingRatesKucrika,
   EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
-\ref{fig1}). Therefore, our results give rise to a conflict: low infection
+\ref{fig1}). Our results give rise to a conflict: low infection
 prevalence leads to high efficiency of Dorfman pooling
@@
 
-Furthermore, an especially important
-consideration in designing pooling schemes is explicitly taking the
-intrinsic PCR error rates into account.
-As we have demonstrated above, analyzing the expected error rates of pooling schemes under a realistic probabilistic error model is readily achievable.
-Designing pooling schemes which optimize testing while accounting for error rates is, however, far from trivial.
-Nonetheless, various algorithms explicitly incorporate error rates into their pooling optimization processes (REFS), including a recent study by our group which employed the herein described error model \cite{DOPE}.
+Thus, improving pooling schemes is imperative. An especially important
+consideration in designing such schemes is explicitly taking the
+intrinsic PCR error rates into account. As we have demonstrated above,
+analyzing expected error rates of a given scheme, given some error
+model is relatively straightforward. Designing a pooling scheme
+tailored for a specific error model is, however, far from trivial. In
+a recent study \cite{DOPE}, we have developed a novel pooling scheme
+that takes advantage of this specific error model. Our new pooling
+scheme shows great promise in terms of both accuracy, as well as
+efficiency, although it has not yet been implemented in real-world
+settings.
 

### 2021-09-09 paper.tex

--- 2021-09-09 paper.tex
@@
   OptimalDorfmanPool}. Other commonly used techniques include
-recursive pooling, wherein... Matrix pooling.. many other complex
-techniques.
+\emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, wherein if the
+first pooled test is positive, the pool is split into two and the
+process repeats. Otherwise, all pool members are declared negative.
+In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
+$N=mn$ is arranged in a matrix of dimensions $m\times n$ matrix. Each
+row and column are then pooled and individuals in the intersection of
+positive rows and columns are tested separately.
 
@@
 the likelihood of a positive pooled test result. A direct consequence
-of our model is that low values of infection
-prevalence increase the false-negative rates of Dorfman pooling. Importantly, our results remain qualitatively similar under varying parameter values,
-in the observed ranges \cite{KitComparison,EstimatingRatesKucrika,
+of our model is that low values of infection prevalence increase the
+false-negative rates of Dorfman pooling. Importantly, our results
+remain qualitatively similar under varying parameter values, in the
+observed ranges \cite{KitComparison,EstimatingRatesKucrika,
   EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
-\ref{fig1}). Therefore, our results give rise to a conflict: low infection
-prevalence leads to high efficiency of Dorfman pooling
+\ref{fig1}). Therefore, our results give rise to a conflict: low
+infection prevalence leads to high efficiency of Dorfman pooling
 \cite{DorfmanYuvalDor}, while also increasing false-negative rates.
@@
 
-Furthermore, an especially important
-consideration in designing pooling schemes is explicitly taking the
-intrinsic PCR error rates into account.
-As we have demonstrated above, analyzing the expected error rates of pooling schemes under a realistic probabilistic error model is readily achievable.
-Designing pooling schemes which optimize testing while accounting for error rates is, however, far from trivial.
-Nonetheless, various algorithms explicitly incorporate error rates into their pooling optimization processes (REFS), including a recent study by our group which employed the herein described error model \cite{DOPE}.
-In our proposed algorithm, we have used a Bayesian experimental design criterion to create an algorithm which iteratively performs pooling schemes, updating its information based on previous results.
-In doing so, our pooling
-scheme shows potential in terms of both reducing false-negative rates, as well as increasing
-efficiency, although it has not yet been implemented in real-world
-settings.
+Furthermore, an especially important consideration in designing
+pooling schemes is explicitly taking the intrinsic PCR error rates
+into account. As we have demonstrated above, analyzing the expected
+error rates of pooling schemes under a realistic probabilistic error
+model is readily achievable. Designing pooling schemes which optimize
+testing while accounting for error rates is, however, far from
+trivial. Nonetheless, various algorithms explicitly incorporate error
+rates into their pooling optimization processes (REFS), including a
+recent study by our group which employed the herein described error
+model \cite{DOPE}.  In our proposed algorithm, we have used a Bayesian
+experimental design criterion to create an algorithm which iteratively
+performs pooling schemes, updating its information based on previous
+results. In doing so, our pooling scheme shows potential in terms of
+both reducing false-negative rates, as well as increasing efficiency,
+although it has not yet been implemented in real-world settings.
 

### 2021-09-09 paper.tex
after uris commentz

@@
 Furthermore, an especially important consideration in designing
-pooling schemes is explicitly taking the intrinsic PCR error rates
+pooling schemes is explicitly taking the intrinsic RT-PCR error rates
 into account. As we have demonstrated above, analyzing the expected
-error rates of pooling schemes under a realistic probabilistic error
+error rates of pooling schemes under a realistic error
 model is readily achievable. Designing pooling schemes which optimize
-testing while accounting for error rates is, however, far from
-trivial. Nonetheless, various algorithms explicitly incorporate error
-rates into their pooling optimization processes (REFS), including a
+testing while accounting for error rates is, however, considerably more involved.
+Nonetheless, various algorithms explicitly incorporate error
+rates into their pooling optimization processes \cite{BayesianDorfman, Kim, OptimalDorfmanPool}, including a
 recent study by our group which employed the herein described error
@@
 experimental design criterion to create an algorithm which iteratively
-performs pooling schemes, updating its information based on previous
+performs pooling, updating its information based on previous
 results. In doing so, our pooling scheme shows potential in terms of
-both reducing false-negative rates, as well as increasing efficiency,
+both reducing false-negative and false-positive rates, as well as increasing efficiency,
 although it has not yet been implemented in real-world settings.

### 2021-09-10 paper.tex
merge

@@
 pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
-its error rates thoroguhly investigated \cite{Kim, Simplistic1,
-  OptimalDorfmanPool}. Other commonly used techniques include
-\emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, wherein if the
+its error rates thoroughly investigated \cite{Kim, Simplistic1,
+  OptimalDorfmanPool}.
+
+The study and practice of pooling has advanced considerably since Dorfman's original paper, with 
+several other pooling techniques being applied for SARS-CoV-2.
+For example, in \emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, if the
 first pooled test is positive, the pool is split into two and the
@@
 In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
-$N=mn$ is arranged in a matrix of dimensions $m\times n$ matrix. Each
+$N=mn$ is arranged in a matrix of dimensions $m\times n$. Each
 row and column are then pooled and individuals in the intersection of
-positive rows and columns are tested separately.
+positive rows and columns are tested separately. Other, more sophisticated
+pooling schemes exist \cite{CompressedPooling}, but studies investigating the 
+underlying assumptions of pooling are surprisingly lacking.
 
@@
 
-Several consequences arise from our analysis. Estimation of optimal
-pool size (for Dorfman pooling), which relies on previously made
-asumptions should be revisited \cite{OptimalDorfmanPool}. The error
-rates of any pooling scheme in use should be reesetimated
+Several consequences arise from our analysis: Previously estimated optimal Dorfman
+pool size should be revisited, as it is influenced by the prevalence \cite{OptimalDorfmanPool}. The error
+rates of many pooling schemes in use should be reesetimated
 \cite{BayesianDorfman, Kim}. And, finally, new methods should be
@@
 In this study, we developed a more realistic and novel probabilistic
-model for the outcomes of pooled PCR tests, parametrized for
+model for the outcomes of pooled RT-PCR tests, parametrized for
 SARS-CoV-2. Contrary to the common assumption, we assume, based on
@@
 Nonetheless, various algorithms explicitly incorporate error
-rates into their pooling optimization processes \cite{BayesianDorfman, Kim, OptimalDorfmanPool}, including a
+rates into their pooling optimization processes \cite{Kim, OptimalDorfmanPool}, including a
 recent study by our group which employed the herein described error
-model \cite{DOPE}.  In our proposed algorithm, we have used a Bayesian
+model \cite{DOPE}.  In our proposed algorithm, we have used an
 experimental design criterion to create an algorithm which iteratively

### 2021-09-12 paper.tex
intro and minor changes everywhere

@@
 Reverse transcription polymerase chain reaction (RT-PCR) testing is a
-key component in breaking transmission chains and mitigating the
-COVID-19 pandemic. As such, the need for large-scale testing has
+key component in the fight against the COVID-19 pandemic.
+PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains.
+Moreover, in some countries, PCR testing has become a routine procedure of surveillance, testing all incoming air passengers in airports (ref), schools students (ref), or even the entire adult population (ref).
+As such, the need for large-scale testing has
 resulted in the application of techniques to increase the throughput
@@
 For example, in \emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, if the
-first pooled test is positive, the pool is split into two and the
+first pooled test is positive, the pool is split into two and the splitting
 process repeats. Otherwise, all pool members are declared negative.
@@
 common in the group testing literature \cite{Kim, OptimalDorfmanPool,
-  CherifReview}, as well as in more specific, COVID-19 focused studies
+  CherifReview}, as well as in more specific, SARS-CoV-2 focused studies
 \cite{Simplistic1, Simplistic2}. In this study we challenge this
-commonly made assumption and show how using a more accurate
+commonly made assumption and demostrate how using a more accurate
 probabilistic model affects the estimation of false-negative rates for
@@
 Several consequences arise from our analysis: Previously estimated optimal Dorfman
-pool size should be revisited, as it is influenced by the prevalence \cite{OptimalDorfmanPool}. The error
-rates of many pooling schemes in use should be reesetimated
-\cite{BayesianDorfman, Kim}. And, finally, new methods should be
-developed in order to take advantage of our novel model \cite{DOPE}.
+pool size \cite{OptimalDorfmanPool} should be revisited in the case of SARS-CoV-2, as it is influenced by the prevalence of infection. The error
+rates of other pooling schemes in use should also be reesetimated
+\cite{BayesianDorfman, Kim}.
+Another consequence of our results is that studies that seek precise error models should be conducted for other pathogens as well, since the standard error model could prove flawed for other pathogens.
+Specifically, all assumptions should be scrutinized, including (e.g.) the independence of tests, an  extremely widespread assumption in the group testing literature \cite{Kim, OptimalDorfmanPool} which dates back to Dorfman's original paper \cite{DorfmanOriginal}.
+Finally, new methods should be developed in order to take advantage of our (and other) more realistic models.
 
@@
   InterpretingCOVID19Test, EstimatingRatesLourenco,
-  FalsePositiveEstimate}, the increase in scheme false-negative rate
-(compared to $1-\Se$), summarized in $\err{exact}$, is at leaset
+  FalsePositiveEstimate}, the increase in the scheme false-negative rate
+(compared to $1-\Se$), expressed by $\err{exact}$, is at leaset
 $60\%$ (Figure \ref{fig1}).
@@
     colors represent the difference between $\err{common}$ and
-    $\err{exact}$. The disease prevalence $q$, is varied on the
+    $\err{exact}$. The infection prevalence $q$, is varied on the
     x-axis, while the test sensitivity is varied on the y-axis. Pool
@@
 Furthermore, an especially important consideration in designing
-pooling schemes is explicitly taking the intrinsic RT-PCR error rates
-into account. As we have demonstrated above, analyzing the expected
+pooling schemes is explicitly taking into account the intrinsic RT-PCR error rates
+. As we have demonstrated above, analyzing the expected
 error rates of pooling schemes under a realistic error
-model is readily achievable. Designing pooling schemes which optimize
+model is readily achievable. Designing pooling schemes that optimize
 testing while accounting for error rates is, however, considerably more involved.

### 2021-09-12 paper.tex
almost done

@@
 key component in the fight against the COVID-19 pandemic.
-PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains.
-Moreover, in some countries, PCR testing has become a routine procedure of surveillance, testing all incoming air passengers in airports (ref), schools students (ref), or even the entire adult population (ref).
+RT-PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains.
+Moreover, in some countries, RT-PCR testing has become a routine procedure of surveillance, testing incoming air passengers in airports \cite{PoolingAirports, TestingAirportsJapan}, schools students \cite{PoolSchool}, or even the entire adult population \cite{MassPooling}.
+
 As such, the need for large-scale testing has
@@
 
-The study and practice of pooling has advanced considerably since Dorfman's original paper, with 
-several other pooling techniques being applied for SARS-CoV-2.
+The study and practice of pooling has advanced considerably since Dorfman's original paper, with several other pooling techniques being applied for SARS-CoV-2.
 For example, in \emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, if the
@@
 \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
-different prevalence values. We observe that for a false-positive rate
+different prevalence values. We observe that for a specificity of
 $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
@@
 infection prevalence leads to high efficiency of Dorfman pooling
-\cite{DorfmanYuvalDor}, while also increasing false-negative rates.
+\cite{DorfmanYuvalDor},but also increases false-negative rates.
 

### 2021-09-12 paper.tex
spellcheck and refactor

@@
 Reverse transcription polymerase chain reaction (RT-PCR) testing is a
-key component in the fight against the COVID-19 pandemic.
-RT-PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains.
-Moreover, in some countries, RT-PCR testing has become a routine procedure of surveillance, testing incoming air passengers in airports \cite{PoolingAirports, TestingAirportsJapan}, schools students \cite{PoolSchool}, or even the entire adult population \cite{MassPooling}.
+key component in the fight against the COVID-19 pandemic. RT-PCR
+testing can be performed on individuals suspected of infection and
+hence break potential transmission chains. Moreover, in some
+countries, RT-PCR testing has become a routine procedure of
+surveillance, testing incoming air passengers in airports
+\cite{PoolingAirports, TestingAirportsJapan}, schools students
+\cite{PoolSchool}, or even the entire adult population
+\cite{MassPooling}.
 
-As such, the need for large-scale testing has
-resulted in the application of techniques to increase the throughput
-of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
-  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
-termed \emph{pooling}, or group testing, have originated in the 1940's
-in the seminal work of Dorfman and his pooling technique
+As such, the need for large-scale testing has resulted in the
+application of techniques to increase the throughput of RT-PCR tests
+\cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman, MatrixPooling,
+  LionDorfman, CherifReview}. These techniques, often termed
+\emph{pooling}, or group testing, have originated in the 1940's in the
+seminal work of Dorfman and his pooling technique
 \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
@@
 
-The study and practice of pooling has advanced considerably since Dorfman's original paper, with several other pooling techniques being applied for SARS-CoV-2.
-For example, in \emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, if the
-first pooled test is positive, the pool is split into two and the splitting
-process repeats. Otherwise, all pool members are declared negative.
-In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
-$N=mn$ is arranged in a matrix of dimensions $m\times n$. Each
-row and column are then pooled and individuals in the intersection of
-positive rows and columns are tested separately. Other, more sophisticated
-pooling schemes exist \cite{CompressedPooling}, but studies investigating the 
+The study and practice of pooling has advanced considerably since
+Dorfman's original paper, with several other pooling techniques being
+applied for SARS-CoV-2.  For example, in \emph{recursive} pooling
+\cite{Kim,RecursiveSevenFold}, if the first pooled test is positive,
+the pool is split into two and the splitting process
+repeats. Otherwise, all pool members are declared negative.  In
+\emph{Matrix} pooling \cite{MatrixPooling} a population of size $N=mn$
+is arranged in a matrix of dimensions $m\times n$. Each row and column
+are then pooled and individuals in the intersection of positive rows
+and columns are tested separately. Other, more sophisticated pooling
+schemes exist \cite{CompressedPooling}, but studies investigating the
 underlying assumptions of pooling are surprisingly lacking.
@@
 common in the group testing literature \cite{Kim, OptimalDorfmanPool,
-  CherifReview}, as well as in more specific, SARS-CoV-2 focused studies
-\cite{Simplistic1, Simplistic2}. In this study we challenge this
-commonly made assumption and demostrate how using a more accurate
+  CherifReview}, as well as in more specific, SARS-CoV-2 focused
+studies \cite{Simplistic1, Simplistic2}. In this study we challenge
+this commonly made assumption and demonstrate how using a more accurate
 probabilistic model affects the estimation of false-negative rates for
@@
 
-Several consequences arise from our analysis: Previously estimated optimal Dorfman
-pool size \cite{OptimalDorfmanPool} should be revisited in the case of SARS-CoV-2, as it is influenced by the prevalence of infection. The error
-rates of other pooling schemes in use should also be reesetimated
-\cite{BayesianDorfman, Kim}.
-Another consequence of our results is that studies that seek precise error models should be conducted for other pathogens as well, since the standard error model could prove flawed for other pathogens.
-Specifically, all assumptions should be scrutinized, including (e.g.) the independence of tests, an extremely widespread assumption in the group testing literature \cite{Kim, OptimalDorfmanPool} which dates back to Dorfman's original paper \cite{DorfmanOriginal}.
-Finally, new methods should be developed in order to take advantage of our (and other) more realistic models.
+Several consequences arise from our analysis: Previously estimated
+optimal Dorfman pool size \cite{OptimalDorfmanPool} should be
+revisited in the case of SARS-CoV-2, as it is influenced by the
+prevalence of infection. The error rates of other pooling schemes in
+use should also be revisited \cite{BayesianDorfman, Kim}. Another
+consequence of our results is that studies that seek precise error
+models should be conducted for other pathogens as well, since the
+standard error model could prove flawed for other
+pathogens. Specifically, all assumptions should be scrutinized,
+including (e.g.) the independence of tests, an extremely widespread
+assumption in the group testing literature \cite{Kim,
+  OptimalDorfmanPool} which dates back to Dorfman's original paper
+\cite{DorfmanOriginal}. Finally, new methods should be developed in
+order to take advantage of our (and other) more realistic models.
 
@@
 in \cite{KitComparison}. These assumptions lead to our model, which is
-illustrated in Figure~\ref{fig:likelihood} and summarized in
-equation \eqref{eq:likelihood}.
+illustrated in Figure~\ref{fig:likelihood} and summarized in equation
+\eqref{eq:likelihood}.
 \begin{align}\label{eq:likelihood}
@@
 In this study, we developed a more realistic and novel probabilistic
-model for the outcomes of pooled RT-PCR tests, parametrized for
+model for the outcomes of pooled RT-PCR tests, parameterized for
 SARS-CoV-2. Contrary to the common assumption, we assume, based on
@@
 Furthermore, an especially important consideration in designing
-pooling schemes is explicitly taking into account the intrinsic RT-PCR error rates
-. As we have demonstrated above, analyzing the expected
-error rates of pooling schemes under a realistic error
-model is readily achievable. Designing pooling schemes that optimize
-testing while accounting for error rates is, however, considerably more involved.
-Nonetheless, various algorithms explicitly incorporate error
-rates into their pooling optimization processes \cite{Kim, OptimalDorfmanPool}, including a
-recent study by our group which employed the herein described error
-model \cite{DOPE}.  In our proposed algorithm, we have used an
-experimental design criterion to create an algorithm which iteratively
-performs pooling, updating its information based on previous
-results. In doing so, our pooling scheme shows potential in terms of
-both reducing false-negative and false-positive rates, as well as increasing efficiency,
-although it has not yet been implemented in real-world settings.
+pooling schemes is explicitly taking into account the intrinsic RT-PCR
+error rates . As we have demonstrated above, analyzing the expected
+error rates of pooling schemes under a realistic error model is
+readily achievable. Designing pooling schemes that optimize testing
+while accounting for error rates is, however, considerably more
+involved.  Nonetheless, various algorithms explicitly incorporate
+error rates into their pooling optimization processes \cite{Kim,
+  OptimalDorfmanPool}, including a recent study by our group which
+employed the herein described error model \cite{DOPE}.  In our
+proposed algorithm, we have used an experimental design criterion to
+create an algorithm which iteratively performs pooling, updating its
+information based on previous results. In doing so, our pooling scheme
+shows potential in terms of both reducing false-negative and
+false-positive rates, as well as increasing efficiency, although it
+has not yet been implemented in real-world settings.
 

### 2021-09-14 paper.tex
Amit's remarks

@@
 \section*{Introduction}
-Reverse transcription polymerase chain reaction (RT-PCR) testing is a
-key component in the fight against the COVID-19 pandemic. RT-PCR
-testing can be performed on individuals suspected of infection and
-hence break potential transmission chains. Moreover, in some
+Reverse transcription polymerase chain reaction (RT-PCR) testing is the backbone and gold standard of infection surveillance across the globe, and a key component in the fight against the COVID-19 pandemic.
+One major application of this technology is the rapid testing of contacts of positive individuals in order to cut transmission chains.
+Moreover, in some
 countries, RT-PCR testing has become a routine procedure of
@@
 \emph{pooling}, or group testing, have originated in the 1940's in the
-seminal work of Dorfman and his pooling technique
+seminal work of Dorfman 
 \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
@@
 sample. Thus, the probability of a positive result in a pooled test
-has been (assumed) identical for a pool with one sample from an
+has been assumed identical for a pool with one sample from an
 infected individual and, e.g., five such samples. This assumption is
@@
 
-Several consequences arise from our analysis: Previously estimated
+Our analysis has several implications: Previously estimated
 optimal Dorfman pool size \cite{OptimalDorfmanPool} should be
-revisited in the case of SARS-CoV-2, as it is influenced by the
+revisited in general and specifically for the case of SARS-CoV-2, as it is influenced by the
 prevalence of infection. The error rates of other pooling schemes in
-use should also be revisited \cite{BayesianDorfman, Kim}. Another
-consequence of our results is that studies that seek precise error
+use should also be revisited \cite{BayesianDorfman, Kim}. Moreover, studies that seek precise error
 models should be conducted for other pathogens as well, since the
-standard error model could prove flawed for other
-pathogens. Specifically, all assumptions should be scrutinized,
+standard error model could prove flawed. Specifically, all assumptions should be scrutinized,
 including (e.g.) the independence of tests, an extremely widespread
@@
 \cite{DorfmanOriginal}. Finally, new methods should be developed in
-order to take advantage of our (and other) more realistic models.
+order to take advantage of more realistic models.
 
@@
 have violated this assumption. However, cross-reactivity was ruled out
-in \cite{KitComparison}. These assumptions lead to our model, which is
+in \cite{KitComparison}. These assumptions lead to the following model,
 illustrated in Figure~\ref{fig:likelihood} and summarized in equation
@@
 We calculate the false-negative rate for a single \emph{infected}
-individual, henceforth referred to as "Donald", under our model and
+individual, henceforth referred to as ''Donald'', under our model and
 under the common assumption. We distinguish three types of
@@
 \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
-different prevalence values. We observe that for a specificity of
+different prevalence values. We consider a specificity of
 $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
@@
   InterpretingCOVID19Test, EstimatingRatesLourenco,
-  FalsePositiveEstimate}, the increase in the scheme false-negative rate
-(compared to $1-\Se$), expressed by $\err{exact}$, is at leaset
+  FalsePositiveEstimate}. We observe that the increase in the scheme false-negative rate
+(compared to $1-\Se$), expressed by $\err{exact}$, is at least
 $60\%$ (Figure \ref{fig1}).
@@
 \section*{Discussion}
-In this study, we developed a more realistic and novel probabilistic
-model for the outcomes of pooled RT-PCR tests, parameterized for
+In this study a novel probabilistic model is developed to capture realistic features of the outcomes of pooled RT-PCR tests, parameterized for
 SARS-CoV-2. Contrary to the common assumption, we assume, based on
@@
 infection prevalence leads to high efficiency of Dorfman pooling
-\cite{DorfmanYuvalDor},but also increases false-negative rates.
+\cite{DorfmanYuvalDor}, but also increases false-negative rates.
 
@@
 used with caution in populations where infection rates are low. Such
-mass-tested populations often include air travel passengers \cite{JTM}
+mass-tested populations often include air travel passengers \cite{JTM},
+nursing homes \cite{DorfmanYuvalDor},
 or presymptomatic and asymptomatic individuals \cite{RobinHood}, and
@@
 pooling schemes is explicitly taking into account the intrinsic RT-PCR
-error rates . As we have demonstrated above, analyzing the expected
+error rates. As illustrated above, analyzing the expected
 error rates of pooling schemes under a realistic error model is

### 2021-09-14 paper.tex
because FUCK ScholarOne Manuscriptsgit add model_pic.jpg git add model_pic.jpg !

@@
 \section*{Introduction}
-Reverse transcription polymerase chain reaction (RT-PCR) testing is the backbone and gold standard of infection surveillance across the globe, and a key component in the fight against the COVID-19 pandemic.
-One major application of this technology is the rapid testing of contacts of positive individuals in order to cut transmission chains.
-Moreover, in some
-countries, RT-PCR testing has become a routine procedure of
-surveillance, testing incoming air passengers in airports
-\cite{PoolingAirports, TestingAirportsJapan}, schools students
-\cite{PoolSchool}, or even the entire adult population
-\cite{MassPooling}.
+Reverse transcription polymerase chain reaction (RT-PCR) testing is
+the backbone and gold standard of infection surveillance across the
+globe, and a key component in the fight against the COVID-19
+pandemic. One major application of this technology is the rapid
+testing of contacts of positive individuals in order to cut
+transmission chains. Moreover, in some countries, RT-PCR testing has
+become a routine procedure of surveillance, testing incoming air
+passengers in airports \cite{PoolingAirports, TestingAirportsJapan},
+schools students \cite{PoolSchool}, or even the entire adult
+population \cite{MassPooling}.
 
@@
 \emph{pooling}, or group testing, have originated in the 1940's in the
-seminal work of Dorfman 
-\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
-selects $N$ individuals and performs a single RT-PCR test on their
-combined (\emph{pooled}) samples. If the pooled test yields a positive
-result, then each individual is retested separately; otherwise,
-everyone is declared negative. The throughput efficiency of Dorfman
-pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
-its error rates thoroughly investigated \cite{Kim, Simplistic1,
-  OptimalDorfmanPool}.
+seminal work of Dorfman \cite{DorfmanOriginal, DorfmanYuvalDor}. Under
+Dorfman pooling, one selects $N$ individuals and performs a single
+RT-PCR test on their combined (\emph{pooled}) samples. If the pooled
+test yields a positive result, then each individual is retested
+separately; otherwise, everyone is declared negative. The throughput
+efficiency of Dorfman pooling has been demonstrated empirically
+\cite{DorfmanYuvalDor} and its error rates thoroughly investigated
+\cite{Kim, Simplistic1, OptimalDorfmanPool}.
 
@@
 sample. Thus, the probability of a positive result in a pooled test
-has been assumed identical for a pool with one sample from an
-infected individual and, e.g., five such samples. This assumption is
-common in the group testing literature \cite{Kim, OptimalDorfmanPool,
+has been assumed identical for a pool with one sample from an infected
+individual and, e.g., five such samples. This assumption is common in
+the group testing literature \cite{Kim, OptimalDorfmanPool,
   CherifReview}, as well as in more specific, SARS-CoV-2 focused
 studies \cite{Simplistic1, Simplistic2}. In this study we challenge
-this commonly made assumption and demonstrate how using a more accurate
-probabilistic model affects the estimation of false-negative rates for
-Dorfman pooling.
+this commonly made assumption and demonstrate how using a more
+accurate probabilistic model affects the estimation of false-negative
+rates for Dorfman pooling.
 
-Our analysis has several implications: Previously estimated
-optimal Dorfman pool size \cite{OptimalDorfmanPool} should be
-revisited in general and specifically for the case of SARS-CoV-2, as it is influenced by the
-prevalence of infection. The error rates of other pooling schemes in
-use should also be revisited \cite{BayesianDorfman, Kim}. Moreover, studies that seek precise error
-models should be conducted for other pathogens as well, since the
-standard error model could prove flawed. Specifically, all assumptions should be scrutinized,
-including (e.g.) the independence of tests, an extremely widespread
-assumption in the group testing literature \cite{Kim,
+Our analysis has several implications: Previously estimated optimal
+Dorfman pool size \cite{OptimalDorfmanPool} should be revisited in
+general and specifically for the case of SARS-CoV-2, as it is
+influenced by the prevalence of infection. The error rates of other
+pooling schemes in use should also be revisited \cite{BayesianDorfman,
+  Kim}. Moreover, studies that seek precise error models should be
+conducted for other pathogens as well, since the standard error model
+could prove flawed. Specifically, all assumptions should be
+scrutinized, including (e.g.) the independence of tests, an extremely
+widespread assumption in the group testing literature \cite{Kim,
   OptimalDorfmanPool} which dates back to Dorfman's original paper
@@
 have violated this assumption. However, cross-reactivity was ruled out
-in \cite{KitComparison}. These assumptions lead to the following model,
-illustrated in Figure~\ref{fig:likelihood} and summarized in equation
-\eqref{eq:likelihood}.
+in \cite{KitComparison}. These assumptions lead to the following
+model, illustrated in Figure~\ref{fig:likelihood} and summarized in
+equation \eqref{eq:likelihood}.
 \begin{align}\label{eq:likelihood}
@@
   \centering
-  \begin{tikzpicture}
-    \begin{scope}[every node/.style={circle,thick,draw}]
-      \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
-      \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
-      \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
-      %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};
+  \includegraphics[width=\textwidth]{model_pic.jpg}
+  %% \begin{tikzpicture}
+  %%   \begin{scope}[every node/.style={circle,thick,draw}]
+  %%     \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
+  %%     \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
+  %%     \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
+  %%     %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};
 
-      \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};
+  %%     \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};
 
-      \node[shape=rectangle,draw=black] (G) at (1.5,5)
-           {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
-    \end{scope}
+  %%     \node[shape=rectangle,draw=black] (G) at (1.5,5)
+  %%          {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
+  %%   \end{scope}
 
-    \begin{scope}[>={Stealth[black]},
-        every node/.style={fill=white},
-        every edge/.style={draw=black,very thick}]
-      \path [->] (A) edge node {$1-\sens$} (2,4.6);
-      %\path [->] (B) edge node {$0$} (G);
-      %\path [->] (D) edge node {$0$} (G);
-      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
-      \path [->] (F) edge node {$\spec$} (3.7,5);
-      %\path [-Bar] (D) edge (4.5, 2);
-    \end{scope}
-  \end{tikzpicture}
+  %%   \begin{scope}[>={Stealth[black]},
+  %%       every node/.style={fill=white},
+  %%       every edge/.style={draw=black,very thick}]
+  %%     \path [->] (A) edge node {$1-\sens$} (2,4.6);
+  %%     %\path [->] (B) edge node {$0$} (G);
+  %%     %\path [->] (D) edge node {$0$} (G);
+  %%     \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
+  %%     \path [->] (F) edge node {$\spec$} (3.7,5);
+  %%     %\path [-Bar] (D) edge (4.5, 2);
+  %%   \end{scope}
+  %% \end{tikzpicture}
   \caption{Illustration of the revised probabilistic model. A pool
@@
 \cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
-different prevalence values. We consider a specificity of
-$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
-sensitivity and prevalence values \cite{KitComparison,
-  InterpretingCOVID19Test, EstimatingRatesLourenco,
-  FalsePositiveEstimate}. We observe that the increase in the scheme false-negative rate
-(compared to $1-\Se$), expressed by $\err{exact}$, is at least
-$60\%$ (Figure \ref{fig1}).
+different prevalence values. We consider a specificity of $\Sp=0.95$
+\cite{DorfmanYuvalDor} and a range of reasonable sensitivity and
+prevalence values \cite{KitComparison, InterpretingCOVID19Test,
+  EstimatingRatesLourenco, FalsePositiveEstimate}. We observe that the
+increase in the scheme false-negative rate (compared to $1-\Se$),
+expressed by $\err{exact}$, is at least $60\%$ (Figure \ref{fig1}).
 
@@
 decreases the scheme false-negative rate, as can also be easily seen
-from equation \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
-the single test false-negative rates increases the relative error
-$\err{exact}$. These effects can be seen in Figure~\ref{fig1} (left
-panel), upon conditioning on pool size. Extending the range for $\Sp$
-yields no qualitative differences. We further compare $\err{exact}$ to
-$\err{common}$, showing the discrepancy changes as a function of both
-prevalence and the single test sensitivity (Figure~\ref{fig1}, right
-panel).
+from equation \eqref{eq:sfn}. For the chosen parameter ranges, the
+increase in the single test false-negative rates increases the
+relative error $\err{exact}$. These effects can be seen in
+Figure~\ref{fig1} (left panel), upon conditioning on pool
+size. Extending the range for $\Sp$ yields no qualitative
+differences. We further compare $\err{exact}$ to $\err{common}$,
+showing the discrepancy changes as a function of both prevalence and
+the single test sensitivity (Figure~\ref{fig1}, right panel).
 \begin{figure}[H]
@@
     $\err{common}$ and $\err{exact}$ can be as large as 30\%, as seen
-    in the right panel. }\label{fig1}
+    in the right panel.}\label{fig1}
 \end{figure}
@@
 \section*{Discussion}
-In this study a novel probabilistic model is developed to capture realistic features of the outcomes of pooled RT-PCR tests, parameterized for
-SARS-CoV-2. Contrary to the common assumption, we assume, based on
-data (Table~\ref{table}), that multiple infected individuals increase
-the likelihood of a positive pooled test result. A direct consequence
-of our model is that low values of infection prevalence increase the
-false-negative rates of Dorfman pooling. Importantly, our results
-remain qualitatively similar under varying parameter values, in the
-observed ranges \cite{KitComparison,EstimatingRatesKucrika,
-  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
-\ref{fig1}). Therefore, our results give rise to a conflict: low
-infection prevalence leads to high efficiency of Dorfman pooling
-\cite{DorfmanYuvalDor}, but also increases false-negative rates.
+In this study a novel probabilistic model is developed to capture
+realistic features of the outcomes of pooled RT-PCR tests,
+parameterized for SARS-CoV-2. Contrary to the common assumption, we
+assume, based on data (Table~\ref{table}), that multiple infected
+individuals increase the likelihood of a positive pooled test
+result. A direct consequence of our model is that low values of
+infection prevalence increase the false-negative rates of Dorfman
+pooling. Importantly, our results remain qualitatively similar under
+varying parameter values, in the observed ranges
+\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
+  InterpretingCOVID19Test} (Figure \ref{fig1}). Therefore, our results
+give rise to a conflict: low infection prevalence leads to high
+efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, but also
+increases false-negative rates.
 
@@
 used with caution in populations where infection rates are low. Such
-mass-tested populations often include air travel passengers \cite{JTM},
-nursing homes \cite{DorfmanYuvalDor},
-or presymptomatic and asymptomatic individuals \cite{RobinHood}, and
-can be crucial for controlling outbreaks \cite{MinaScience}.
+mass-tested populations often include air travel passengers
+\cite{JTM}, nursing homes \cite{DorfmanYuvalDor}, or presymptomatic
+and asymptomatic individuals \cite{RobinHood}, and can be crucial for
+controlling outbreaks \cite{MinaScience}.
 
@@
 pooling schemes is explicitly taking into account the intrinsic RT-PCR
-error rates. As illustrated above, analyzing the expected
-error rates of pooling schemes under a realistic error model is
-readily achievable. Designing pooling schemes that optimize testing
-while accounting for error rates is, however, considerably more
-involved.  Nonetheless, various algorithms explicitly incorporate
-error rates into their pooling optimization processes \cite{Kim,
+error rates. As illustrated above, analyzing the expected error rates
+of pooling schemes under a realistic error model is readily
+achievable. Designing pooling schemes that optimize testing while
+accounting for error rates is, however, considerably more involved.
+Nonetheless, various algorithms explicitly incorporate error rates
+into their pooling optimization processes \cite{Kim,
   OptimalDorfmanPool}, including a recent study by our group which
@@
 %% add refs to dilution and paragraph in start
+PCR pooling , or simultaneously testing multiple samples, for the
+presence of SARS-CoV-2 is increasing in popularity. It is commonly
+assumed that including more than a single infected sample in a pooled
+test does not increase the probability of a positive outcome. We
+challenge this assumption with experimental data. We then analyze the
+probability of misidentifying an infected individual and show that it
+increases when the prevalence of infection decreases. However, low
+infection prevalence is exactly the condition when pooling is most
+efficient. We therefore implore the cautious use of pooling.
+
+Relative increase in Dorfman pooling false-negative rates
+$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
+percentage increase in the scheme false-negative rates relative to the
+single test false-negative rates equation \eqref{eq:erel}. Right:
+colors represent the difference between $\err{common}$ and
+$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
+while the test sensitivity is varied on the y-axis. Pool size $N$, was
+chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
+shows that $\err{exact}$ is largest for low prevalence values $q$. The
+difference between $\err{common}$ and $\err{exact}$ can be as large as
+30\%, as seen in the right panel.
+
+Relative increase in Dorfman pooling false-negative rates
+$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
+percentage increase in the scheme false-negative rates relative to the
+single test false-negative rates equation \eqref{eq:erel}. Right:
+colors represent the difference between $\err{common}$ and
+$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
+while the test sensitivity is varied on the y-axis. Pool size $N$, was
+chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
+shows that $\err{exact}$ is largest for low prevalence values $q$. The
+difference between $\err{common}$ and $\err{exact}$ can be as large as
+30\%, as seen in the right panel.

### 2021-09-15 paper.tex
all in 1

@@
 \documentclass{article}
-\input{definitions}
+\usepackage{authblk}
+\usepackage{amsmath}
+\usepackage{todonotes}
+\usepackage{amssymb}
+\usepackage{subcaption}
+\usepackage[normalem]{ulem}
+\usepackage{graphicx}
+\usepackage{cite}
+\usepackage{marginnote}
+\usepackage[colorlinks=true,urlcolor=blue,citecolor=red,linkcolor=red,bookmarks=true]{hyperref}
+\usepackage{soul}
+%% \usepackage{pgfplots}
+%% \pgfplotsset{compat=1.14}
+%% \usepackage{tikz}
+%% \usetikzlibrary{arrows.meta}
+%% \usetikzlibrary{shapes.geometric}
+%% \usetikzlibrary{bayesnet} 
+\usepackage{float}
+\usepackage{geometry}
+\hypersetup{linkcolor=red,citecolor=green,urlcolor=blue}
+
+\newcommand{\fp}{P_{\text{fp}}}
+\newcommand{\fn}{P_{\text{fn}}}
+\newcommand{\Se}{S_\text{e}}
+\newcommand{\Sp}{S_\text{p}}
+\newcommand{\mi}{S_{\text{fn}}}
+\newcommand{\err}[1]{E_{\text{{#1}}}}%_{\text{rel}}}
+\newcommand{\eff}[1]{E_{\text{{#1}}}}%_{\text{rel}}}
+\renewcommand\Affilfont{\itshape\small}
+\renewcommand{\Pr}{\mathbb{P}}
+\newcommand{\data}{\mathbf{d}}
+\newcommand{\sens}{S_{\text{e}}}
+\newcommand{\spec}{S_{\text{p}}}
+% \newcommand{\tn}{P_{\text{tn}}}
 
@@
 \bibliographystyle{amsplain}
-\bibliography{refs}
+\providecommand{\bysame}{\leavevmode\hbox to3em{\hrulefill}\thinspace}
+\providecommand{\MR}{\relax\ifhmode\unskip\space\fi MR }
+% \MRhref is called by the amsart/book/proc definition of \MR.
+\providecommand{\MRhref}[2]{%
+  \href{http://www.ams.org/mathscinet-getitem?mr=#1}{#2}
+}
+\providecommand{\href}[2]{#2}
+\begin{thebibliography}{10}
+
+\bibitem{OptimalDorfmanPool}
+Hrayer Aprahamian, Douglas~R Bish, and Ebru~K Bish, \emph{Optimal group
+  testing: Structural properties and robust solutions, with application to
+  public health screening}, INFORMS Journal on Computing (2020).
+
+\bibitem{DorfmanYuvalDor}
+Netta Barak, Roni Ben-Ami, Tal Sido, Amir Perri, Aviad Shtoyer, Mila Rivkin,
+  Tamar Licht, Ayelet Peretz, Judith Magenheim, Irit Fogel, et~al.,
+  \emph{Lessons from applied large-scale pooling of 133,816 {SARS-CoV}-2
+  {RT-PCR} tests}, Science Translational Medicine (2021).
+
+\bibitem{Simplistic2}
+Alhaji Cherif, Nadja Grobe, Xiaoling Wang, and Peter Kotanko, \emph{Simulation
+  of pool testing to identify patients with coronavirus disease 2019 under
+  conditions of limited test availability}, JAMA network open \textbf{3}
+  (2020), no.~6, e2013075--e2013075.
+
+\bibitem{DOPE}
+Yair Daon, Amit Huppert, and Uri Obolsik, \emph{{DOPE}: {D}-{O}ptimal {P}ooling
+  {E}xperimental design with application to {SARS-CoV}-2 screening}, {arXiv}
+  (2021).
+
+\bibitem{JTM}
+Yair Daon, Robin~N Thompson, and Uri Obolski, \emph{Estimating covid-19
+  outbreak risk through air travel}, Journal of Travel Medicine \textbf{27}
+  (2020), no.~5, taaa093.
+
+\bibitem{Salazar}
+Adolfo de~Salazar, Antonio Aguilera, Rocio Trastoy, Ana Fuentes, Juan~Carlos
+  Alados, Manuel Causse, Juan~Carlos Gal{\'a}n, Antonio Moreno, Matilde Trigo,
+  Mercedes P{\'e}rez-Ruiz, et~al., \emph{Sample pooling for {SARS-CoV-2 RT-PCR}
+  screening}, Clinical Microbiology and Infection \textbf{26} (2020), no.~12,
+  1687.e1--1687.e5.
+
+\bibitem{PoolingAirports}
+Philip Dollard, Isabel Griffin, Andre Berro, Nicole~J Cohen, Kimberly Singler,
+  Yoni Haber, Chris de~la Motte~Hurst, Amber Stolp, Sukhshant Atti, Leslie
+  Hausman, et~al., \emph{Risk assessment and management of covid-19 among
+  travelers arriving at designated us airports, january 17--september 13,
+  2020}, Morbidity and Mortality Weekly Report \textbf{69} (2020), no.~45,
+  1681.
+
+\bibitem{DorfmanOriginal}
+Robert Dorfman, \emph{The detection of defective members of large populations},
+  The Annals of Mathematical Statistics \textbf{14} (1943), no.~4, 436--440.
+
+\bibitem{RecursiveSevenFold}
+Jens~Niklas Eberhardt, Nikolas~Peter Breuckmann, and Christiane~Sigrid
+  Eberhardt, \emph{Multi-stage group testing improves efficiency of large-scale
+  {COVID}-19 screening}, Journal of Clinical Virology \textbf{128} (2020),
+  104382.
+
+\bibitem{CherifReview}
+Nadja Grobe, Alhaji Cherif, Xiaoling Wang, Zijun Dong, and Peter Kotanko,
+  \emph{Sample pooling: burden or solution?}, Clinical Microbiology and
+  Infection (2021).
+
+\bibitem{LionDorfman}
+Rudolf Hanel and Stefan Thurner, \emph{Boosting test-efficiency by pooled
+  testing for {SARS-CoV}-2---{F}ormula for optimal pool size}, PLoS One
+  \textbf{15} (2020), no.~11, e0240652.
+
+\bibitem{DilutionCOVID}
+Yosuke Hirotsu, Makoto Maejima, Masahiro Shibusawa, Yuki Nagakubo, Kazuhiro
+  Hosaka, Kenji Amemiya, Hitomi Sueki, Miyoko Hayakawa, Hitoshi Mochizuki,
+  Toshiharu Tsutsui, et~al., \emph{Pooling {RT-qPCR} testing for {SARS-CoV-2}
+  in 1000 individuals of healthy and infection-suspected patients}, Scientific
+  reports \textbf{10} (2020), no.~1, 1--8.
+
+\bibitem{MassPooling}
+Ed~Holt, \emph{Slovakia to test all adults for sars-cov-2}, The Lancet
+  \textbf{396} (2020), no.~10260, 1386--1387.
+
+\bibitem{GroupDilution}
+Frank~K Hwang, \emph{Group testing with a dilution effect}, Biometrika
+  \textbf{63} (1976), no.~3, 671--680.
+
+\bibitem{Kim}
+Hae-Young Kim, Michael~G Hudgens, Jonathan~M Dreyfuss, Daniel~J Westreich, and
+  Christopher~D Pilcher, \emph{Comparison of group testing algorithms for case
+  identification in the presence of test errors}, Biometrics \textbf{63}
+  (2007), no.~4, 1152--1163.
+
+\bibitem{EstimatingRatesKucrika}
+Lauren~M Kucirka, Stephen~A Lauer, Oliver Laeyendecker, Denali Boon, and Justin
+  Lessler, \emph{Variation in false-negative rate of reverse transcriptase
+  polymerase chain reaction---based {SARS-CoV}-2 tests by time since exposure},
+  Annals of Internal Medicine (2020).
+
+\bibitem{MatrixPooling}
+Thomas~J Kwiatkowski~Jr, Huda~Y Zoghbi, Susan~A Ledbetter, Kimberly~A Ellison,
+  and A~Craig Chinault, \emph{Rapid identification of yeast artificial
+  chromosome clones by matrix pooling and crude lysate {PCR}.}, Nucleic acids
+  research \textbf{18} (1990), no.~23, 7191.
+
+\bibitem{PoolSize30}
+Stefan Lohse, Thorsten Pfuhl, Barbara Berk{\'o}-G{\"o}ttel, J{\"u}rgen
+  Rissland, Tobias Gei{\ss}ler, Barbara G{\"a}rtner, S{\"o}ren~L Becker, Sophie
+  Schneitler, and Sigrun Smola, \emph{Pooling of samples for testing for
+  {SARS}-{C}o{V}-2 in asymptomatic people}, The Lancet Infectious Diseases
+  (2020).
+
+\bibitem{MinaScience}
+Michael~J Mina and Kristian~G Andersen, \emph{{COVID}-19 testing: {O}ne size
+  does not fit all}, Science \textbf{371} (2021), no.~6525, 126--127.
+
+\bibitem{BayesianDorfman}
+Rodrigo Noriega and Matthew Samore, \emph{Increasing testing throughput and
+  case detection with a pooled-sample {B}ayesian approach in the context of
+  {COVID}-19}, bioRxiv (2020).
+
+\bibitem{TestingAirportsJapan}
+Masataro Norizuki, Masahiko Hachiya, Ayano Motohashi, Ataru Moriya, Kazuhisa
+  Mezaki, Moto Kimura, Wataru Sugiura, Hidechika Akashi, and Tamami Umeda,
+  \emph{Effective screening strategies for detection of asymptomatic covid-19
+  travelers at airport quarantine stations: Exploratory findings in japan},
+  Global Health \& Medicine (2021), 2020--01109.
+
+\bibitem{Simplistic1}
+Alexander Pikovski and Kajetan Bentele, \emph{Pooling of coronavirus tests
+  under unknown prevalence}, Epidemiology \& Infection \textbf{148} (2020).
+
+\bibitem{CompressedPooling}
+Noam Shental, Shlomia Levy, Vered Wuvshet, Shosh Skorniakov, Bar Shalem, Aner
+  Ottolenghi, Yariv Greenshpan, Rachel Steinberg, Avishay Edri, Roni Gillis,
+  et~al., \emph{Efficient high-throughput {SARS-CoV}-2 testing to detect
+  asymptomatic carriers}, Science advances \textbf{6} (2020), no.~37, eabc5961.
+
+\bibitem{PoolSchool}
+Catherine~M Sweeney-Reed, Doreen Wolff, Jakob Niggel, Michael Kabesch, and
+  Christian Apfelbacher, \emph{Pool testing as a strategy for prevention of
+  sars-cov-2 outbreaks in schools: Protocol for a feasibility study}, JMIR
+  Research Protocols \textbf{10} (2021), no.~5, e28673.
+
+\bibitem{RobinHood}
+Robin~N Thompson, Francesca~A Lovell-Read, and Uri Obolski, \emph{Time from
+  symptom onset to hospitalisation of coronavirus disease 2019 ({COVID-19})
+  cases: implications for the proportion of transmissions from infectors with
+  few symptoms}, Journal of clinical medicine \textbf{9} (2020), no.~5, 1297.
+
+\bibitem{KitComparison}
+Puck~B van Kasteren, Bas van Der~Veer, Sharon van~den Brink, Lisa Wijsman,
+  J{\o}rgen de~Jonge, Annemarie van~den Brandt, Richard Molenkamp, Chantal~BEM
+  Reusken, and Adam Meijer, \emph{Comparison of seven commercial {RT-PCR}
+  diagnostic kits for {COVID}-19}, Journal of Clinical Virology \textbf{128}
+  (2020), 104412.
+
+\bibitem{InterpretingCOVID19Test}
+Jessica Watson, Penny~F Whiting, and John~E Brush, \emph{Interpreting a
+  {COVID}-19 test result}, {BMJ} \textbf{369} (2020).
+
+\bibitem{DilutionHIV}
+Lawrence~M Wein and Stefanos~A Zenios, \emph{Pooled testing for {HIV}
+  screening: capturing the dilution effect}, Operations Research \textbf{44}
+  (1996), no.~4, 543--569.
+
+\bibitem{EstimatingRatesLourenco}
+Paul~S Wikramaratna, Robert~S Paton, Mahan Ghafari, and Jos{\'e}
+  Louren{\c{c}}o, \emph{Estimating the false-negative test probability of
+  {SARS-CoV}-2 by {RT-PCR}}, Eurosurveillance \textbf{25} (2020), no.~50,
+  2000568.
+
+\bibitem{FalsePositiveEstimate}
+Steven Woloshin, Neeraj Patel, and Aaron~S Kesselheim, \emph{False negative
+  tests for {SARS}-{C}o{V}-2 infection-challenges and implications}, New
+  England Journal of Medicine (2020).
+
+\bibitem{Lion}
+Idan Yelin, Noga Aharony, Einat~Shaer Tamar, Amir Argoetti, Esther Messer, Dina
+  Berenbaum, Einat Shafran, Areen Kuzli, Nagham Gandali, Omer Shkedi, et~al.,
+  \emph{Evaluation of {COVID-19 RT-qPCR} test in multi sample pools}, Clinical
+  Infectious Diseases \textbf{71} (2020), no.~16, 2073--2078.
+
+\end{thebibliography}
+
+% \bibliography{refs}
 
@@
 30\%, as seen in the right panel.
+
+Illustration of the revised probabilistic model. A pool contains
+individuals $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals
+1 and 2 are infected). A negative pooled test ($\data=0$) occurs when
+three detection paths fail: A false-negative occurs for individuals
+$1$ and $2$, each with probability $(1-\sens)$.  Additionally, no
+false-positive detection occurs, with probability $\spec$. Individual
+$3$ is not infected and does not contribute to the probability of the
+pooled test result.

### 2021-09-15 paper.tex
Hi Royal Society Open whatever. Screw you for wasting my evening just for creating a useless DOI for my code. Sincerely, Yair

@@
 \subsection*{Code Availability}
-Code for this manuscript is available at
-\url{https://github.com/yairdaon/errorates}.
+Data and relevant code for this research work are stored in GitHub:
+\url{https://github.com/yairdaon/errorates} and have been archived
+within the Zenodo repository:
+\url{https://doi.org/10.5281/zenodo.5510563} \cite{MyCode}.
 
@@
   (2020), no.~6, e2013075--e2013075.
+
+\bibitem{MyCode}
+Yair Daon, \emph{yairdaon/errorates: Release for archiving}, 2021.
 
@@
 \end{document}
-
-%% Todos
-%% add refs to dilution and paragraph in start
-PCR pooling , or simultaneously testing multiple samples, for the
-presence of SARS-CoV-2 is increasing in popularity. It is commonly
-assumed that including more than a single infected sample in a pooled
-test does not increase the probability of a positive outcome. We
-challenge this assumption with experimental data. We then analyze the
-probability of misidentifying an infected individual and show that it
-increases when the prevalence of infection decreases. However, low
-infection prevalence is exactly the condition when pooling is most
-efficient. We therefore implore the cautious use of pooling.
-
-Relative increase in Dorfman pooling false-negative rates
-$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
-percentage increase in the scheme false-negative rates relative to the
-single test false-negative rates equation \eqref{eq:erel}. Right:
-colors represent the difference between $\err{common}$ and
-$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
-while the test sensitivity is varied on the y-axis. Pool size $N$, was
-chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
-shows that $\err{exact}$ is largest for low prevalence values $q$. The
-difference between $\err{common}$ and $\err{exact}$ can be as large as
-30\%, as seen in the right panel.
-
-Relative increase in Dorfman pooling false-negative rates
-$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
-percentage increase in the scheme false-negative rates relative to the
-single test false-negative rates equation \eqref{eq:erel}. Right:
-colors represent the difference between $\err{common}$ and
-$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
-while the test sensitivity is varied on the y-axis. Pool size $N$, was
-chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
-shows that $\err{exact}$ is largest for low prevalence values $q$. The
-difference between $\err{common}$ and $\err{exact}$ can be as large as
-30\%, as seen in the right panel.
-
-Illustration of the revised probabilistic model. A pool contains
-individuals $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals
-1 and 2 are infected). A negative pooled test ($\data=0$) occurs when
-three detection paths fail: A false-negative occurs for individuals
-$1$ and $2$, each with probability $(1-\sens)$.  Additionally, no
-false-positive detection occurs, with probability $\spec$. Individual
-$3$ is not infected and does not contribute to the probability of the
-pooled test result.
